---
title: "Immune_Cvi_Sca"
author: "Peri Bolton"
date: "2024-02-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This outlines the process of identifying the Crotalus viridis orthologs in Sistrurus catenatus and lining these up with the DoS values from Mathur et al. 2023 PNAS, SFD associated DEGs from Mathur et al 2024 in Crotalus viridis and the immune genes I identified in 'candidate_genes_download' file. 

# Orthologs 

Using total set of Cviridis/Scatenatus/Stergiminus orthologs.

```{r read and summarize ortho info}
library(stringr)
library(ggvenn)
library(tidyr)
library(ggplot2)
library(ggpubr)
setwd("~/Library/CloudStorage/OneDrive-TheOhioStateUniversity/Rattlesnakes/Scate_SFD/analysis")
ortho<- read.table("../data/Mason unpubl/Orthogroups.tsv", fill=TRUE, sep="\t", header=TRUE)
colnames(ortho)<- c("Orthogroup","geneID","Scate","Sterg")
#head(ortho)
#ortho[5,]
ortho$geneID<- gsub("_cds","",ortho$geneID)


ortho_dispersed<- ortho %>% separate_longer_delim(geneID, delim = ",") 
ortho_dispersed<- ortho_dispersed %>% separate_longer_delim(Scate, delim=",")
ortho_dispersed$geneID<- trimws(ortho_dispersed$geneID)
ortho_dispersed$Scate<- trimws(ortho_dispersed$Scate)



ortho_single_copy<- ortho[!grepl(",",ortho$geneID),]
ortho_single_copy<- ortho_single_copy[!grepl(",",ortho_single_copy$Scate),]
ortho_multi_copy<- ortho[grepl(",",ortho$geneID) | grepl(",",ortho$Scate) ,]


```

# Cvi annotation

```{r read Cvi annotation}
library(rtracklayer)

annot<- readGFF("../data/Mathur_2024/cvir_genes.gff")
annot<- as.data.frame(annot)
annot$Anolis_Genename<- word(annot$Anolis_Homolog,4,4, sep="_")
annot$Anolis_Genename<- toupper(annot$Anolis_Genename)
annot$Anolis_entrez<- word(annot$Anolis_Homolog,3,3, sep="_")
annot$Anolis_Protein<- word(annot$Anolis_Homolog,1,2, sep="_")
annot$Anolis_Protein<- gsub("\\.[0-9]","",annot$Anolis_Protein)
annot$Python_Genename<- word(annot$Python_Homolog,4,4, sep="_")
annot$Python_entrez<-word(annot$Python_Homolog,3,3, sep="_")
annot$Thamnophis_Genename<- word(annot$Thamnophis_Homolog,4,4,sep="_")
annot$Thamnophis_entrez<- word(annot$Thamnophis_Homolog,3,3, sep="_")


## remove duplicates
#annot$geneID<- gsub("\\:exon\\:[0-9]+","",annot$ID)
annot$geneID<- word(annot$ID,1,1, sep="\\:")
annot<- annot[!duplicated(annot$geneID),]
#keep only mRNA annotations that were possible for calculating orthology.
annot<- annot[grepl("-mRNA-1",annot$geneID),]

#merge orthologs with Cvi annotation 
ortho_dispersed_annot<- merge(ortho_dispersed, annot, by="geneID", all=TRUE)
ortho_dispersed_annot$geneID<- gsub("-mRNA-[0-9]","",ortho_dispersed_annot$geneID)
#nrows is larger because of some number of genes had no orthologs. 
summary(is.na(ortho_dispersed_annot$Orthogroup))
```




# Cvi DEGs
```{r read DEGs}
liverTreat<- read.csv("https://raw.githubusercontent.com/samarth8392/SFDTranscriptomics/main/DataFiles/Liver/genelist/Liver_DGE_TreatOnly_withAnn.csv")
liverTreat$GeneID<- trimws(liverTreat$GeneID) #making sure mismatches don't occure due to random white-space at ends
liverTemp<- read.csv("https://raw.githubusercontent.com/samarth8392/SFDTranscriptomics/main/DataFiles/Liver/genelist/Liver_DGE_TempOnly_withAnn.csv")
liverTemp$GeneID<- trimws(liverTemp$GeneID)
liverInter<- read.csv("https://raw.githubusercontent.com/samarth8392/SFDTranscriptomics/main/DataFiles/Liver/genelist/Liver_DGE_Inter_withAnn.csv")
liverInter$GeneID<- trimws(liverInter$GeneID)
liverDEGs<- list(liverTemp=liverTemp$GeneID, liverTreat=liverTreat$GeneID, liverInter=liverInter$GeneID)
ggvenn(liverDEGs)

kidneyInter<- read.csv("https://raw.githubusercontent.com/samarth8392/SFDTranscriptomics/main/DataFiles/Kidney/genelist/Kidney_DGE_Inter_withAnn.csv")
kidneyInter$GeneID<- trimws(kidneyInter$GeneID)
kidneyTreat<- read.csv("https://raw.githubusercontent.com/samarth8392/SFDTranscriptomics/main/DataFiles/Kidney/genelist/Kidney_DGE_TreatOnly_withAnn.csv")
kidneyTreat$GeneID<- trimws(kidneyTreat$GeneID)
skinTreat<- read.csv("https://raw.githubusercontent.com/samarth8392/SFDTranscriptomics/main/DataFiles/Skin/genelist/Skin_DGE_TreatOnly_withAnn.csv")
skinTreat$GeneID<- trimws(skinTreat$GeneID)

#create a list of unique gene IDs for possible disease genes
total<- rbind(liverTreat, liverInter, kidneyTreat, kidneyInter, skinTreat)
length(unique(total$GeneID)) #how many unique geneIDs are here.
```



# Sca annotation

Read in the S. catenatus annotation

```{r Read Sca gff}
Sca_annot<- readGFF("../../reference_genome/Inhouse_assembly/Scate_HiC_rnd4.all.putative.function.gff")
nrow(Sca_annot)
Sca_annot<- as.data.frame(Sca_annot)
Sca_annot$bed<- paste0(Sca_annot$seqid,":", Sca_annot$start,"-", Sca_annot$end)

#columns ID and Name contain the IDs that correspond to the Ortholog file. Change the name for merging.
colnames(Sca_annot)[colnames(Sca_annot)=="Name"]<- "Scate"
summary(grepl("-RA",Sca_annot$Scate)) #17061 true
summary(grepl("-RA|MYO|NGF",Sca_annot$Scate))
summary(grepl("mRNA",Sca_annot$type))
unique(Sca_annot$type[grepl("-RA|MYO|NGF",Sca_annot$Scate)])
unique(Sca_annot$type[grepl("MYO|NGF",Sca_annot$Scate)])


Sca_annot<- Sca_annot[grepl("-RA|MYO|NGF",Sca_annot$Scate),]
Sca_annot<- droplevels(Sca_annot)
Sca_annot[grepl("MYO|NGF",Sca_annot$Scate),]


Sca_annot$Scate_annot_GeneName<- word(Sca_annot$Note,3,3,sep=" ") 
Sca_annot$Scate_annot_GeneName<- gsub("\\:","",Sca_annot$Scate_annot_GeneName)
Sca_annot$Scate_annot_Taxon<- str_extract(Sca_annot$Note,"Anolis carolinensis|Crotalus horridus|Ophiophagus hannah|Crotalus adamanteus|Crotalus atrox|Crotalus scutulatus|Sistrurus miliarius|Sistrurus tergeminus|Crotalus ruber|Crotalus cerastes|Crotalus lepidus|Crotalus durissus|Crotalus molossus|Crotalus tigris")

#these don't seem to correspond to the MYO and NGF IDs from the orthology file. But they will come out later. 

#pull the Scate equivalents of the SFD DEGs, where #total file is the list of SFD associated genes in ALL tissues. 
#ortho_annot_DEGs<- ortho_dispersed_annot[ortho_dispersed_annot$geneID %in% total$GeneID,]
```


Merge with the Cvi annotation that contains orthology results. Merge based on S. catenatus geneID.
So for each S.catenatus gene ID we have the Orthogroup, and the Cvi gene ID and its annotation with it. 
```{r}
#"Scate_annot_GeneName","Scate_annot_Taxon"

Sca_Cvi<- merge(Sca_annot[,c("Scate","bed","Note", "Scate_annot_GeneName","Scate_annot_Taxon")], ortho_dispersed_annot[,c("Scate","geneID","Orthogroup", "seqid","Anolis_Genename","Anolis_entrez","Anolis_Protein","Python_Genename","Python_entrez","Thamnophis_Genename","Thamnophis_entrez")], by="Scate", all.x=TRUE)
Sca_Cvi$Sca_Cvi<- paste0(Sca_Cvi$Scate, Sca_Cvi$geneID)
summary(is.na(Sca_Cvi$Sca_Cvi))
summary(duplicated(Sca_Cvi$Sca_Cvi))
#dups<-Sca_Cvi$Sca_Cvi[duplicated(Sca_Cvi$Sca_Cvi)]
#Sca_Cvi[Sca_Cvi$Sca_Cvi %in% dups,]
#Sca_Cvi<- Sca_Cvi[!duplicated(Sca_Cvi$Sca_Cvi),]
#summary(is.na(Sca_Cvi$Scate))

#pull the Scate equivalents of the SFD DEGs, where #total file is the list of SFD associated genes in ALL tissues. 
#Sca_DEGs<- as.data.frame(Sca_Cvi[which(Sca_Cvi$Scate %in% ortho_annot_DEGs$Scate),])
Sca_DEGs<- as.data.frame(Sca_Cvi[which(Sca_Cvi$geneID %in% total$GeneID),])
colnames(total)[colnames(total)=="GeneID"]<- "geneID"
Sca_DEGs2<- merge(Sca_Cvi, total, by="geneID")
Sca_DEGs2<- Sca_DEGs2[!is.na(Sca_DEGs2$Scate),]
length(unique(Sca_DEGs$Scate))
length(unique(Sca_DEGs$geneID))
#Sca_DEGs<- Sca_DEGs[!duplicated(Sca_DEGs$Scate),]
Sca_DEGs$Note<- unlist(Sca_DEGs$Note)
write.table(Sca_DEGs, file="../data/lists/Scate_DEGs_annot.tsv", row.names=FALSE, quote=TRUE, sep="\t")
```


# PNAS DoS

```{r}

#read in the bed file for the lower 10% of DoS values. 
lower<- read.csv("https://raw.githubusercontent.com/samarth8392/FGDvNGD/main/data/lower10_genes.bed.csv")
lower$bed<- paste0(lower$Chr,":", lower$Start,"-",lower$End)


#from the annotation above that includes Sca + Cvi annotation, pull out the genes based on the coordinates from the lower 10% file. 
Sca_annot_lower<- as.data.frame(Sca_Cvi[which(Sca_Cvi$bed %in% lower$bed),])


#lower2<- read.csv("../data/Mathur 2023 PNAS/pnas.2303043120.sd02.lower_correctedUniProtID.csv")
#Sca_annot_lower_notin2<- Sca_annot_lower[!Sca_annot_lower$bed %in% lower2$Coordinates,]
#Sca_annot_lower_notin2$Scate<- gsub("-RA","", Sca_annot_lower_notin2$Scate)
#not2_annot<- as.data.frame(Sca_annot[ grepl(paste0(Sca_annot_lower_notin2$Scate,collapse="|"),Sca_annot$ID),])

#keep only the geneIDs that have "-RA" at the end as these were the ones put through orthofinder.
Sca_annot_lower<- Sca_annot_lower[grepl("-RA",Sca_annot_lower$Scate),]
length(unique(Sca_annot_lower$bed)) #292 - this equals the input file length the actual file is longer because of multiple orthologs within Crotalus viridis



#read in the bed file for the upper 10% of DoS
upper<- read.csv("https://raw.githubusercontent.com/samarth8392/FGDvNGD/main/data/upper10_genes.bed.csv")
upper$bed<- paste0(upper$Chr,":", upper$Start,"-",upper$End)

#from the annotation above that includes Sca + Cvi annotation, pull out the genes based on the coordinates from the lower 10% file. 
Sca_annot_upper<- Sca_Cvi[which(Sca_Cvi$bed %in% upper$bed),]
Sca_annot_upper<- Sca_annot_upper[grepl("-RA",Sca_annot_upper$Scate),]
length(unique(Sca_annot_upper$bed)) #249 - this equals the input file length the actual file is longer because of paralogy 
```



```{r}
Cvi_immune_annot<- read.table("../data/candidate genes/v2/Cvi_immune_annot.txt", header=TRUE, sep="\t", quote="\"")
#merge with Sca annot. 
length(unique(Cvi_immune_annot$Scate))

Cvi_immune_annot2<- merge(Sca_annot[,c("Scate","bed","Scate_annot_GeneName","Scate_annot_Taxon")], Cvi_immune_annot, by="Scate")



#this is slightly shorter because some of these were annotated by andrew (Venoms Ficolins and VEGFa) and used in the orthogroup analysis. 
#length(unique(Cvi_immune_annot2$Scate))
#Cvi_immune_annot$Scate[!Cvi_immune_annot$Scate %in% Sca_annot$Scate] 

#Cvi_immune_annot2$Sca_Cvi<- paste0(Cvi_immune_annot2$Scate, Cvi_immune_annot2$geneID)
#summary(duplicated(Cvi_immune_annot2$Sca_Cvi))
#Cvi_immune_annot2<- Cvi_immune_annot2[!duplicated(Cvi_immune_annot2$Sca_Cvi),]
#nrow(Cvi_immune_annot2[which(Cvi_immune_annot2$Scate=="Scate00198-RA"),])
#nrow(Sca_Cvi[which(Sca_Cvi$Scate=="Scate00198-RA"),])

SFD_DoS_overlap<- list(lower_DoS=Sca_annot_lower$Scate, upper_DoS=Sca_annot_upper$Scate, Sca_immune=unique(Cvi_immune_annot2$Scate), DEG_ortho=unique(Sca_DEGs$Scate))
ggvenn(SFD_DoS_overlap) + labs(title="Overlap between PNAS DoS, Sca Immune Genes and Cvi DEG orthologs")

DEG_adaptive_immune<- Sca_DEGs2[Sca_DEGs2$Scate %in% Sca_annot_upper$Scate & Sca_DEGs2$Scate %in% Cvi_immune_annot2$Scate,]
DEG_adaptive_immune$category<- "Adaptive + Immune"
#this method pulls the full set of Crotalus annotations for each unique Scate gene ID.
#Immune_adaptive_DEG<- Cvi_immune_annot2[Cvi_immune_annot2$Scate %in% Sca_annot_upper$Scate & Cvi_immune_annot2$Scate %in% Sca_DEGs$Scate,]

#Immune_adaptive_DEG

#adaptive_DEG<- Sca_DEGs[Sca_DEGs$Scate %in% Sca_annot_upper$Scate & !Sca_DEGs$Scate %in% Cvi_immune_annot$Scate,]
DEG_adaptive<- Sca_DEGs2[Sca_DEGs2$Scate %in% Sca_annot_upper$Scate & !Sca_DEGs2$Scate %in% Cvi_immune_annot2$Scate,]
DEG_adaptive$category<- "Adaptive"
DEG_delet_immune<- Sca_DEGs2[Sca_DEGs2$Scate %in% Sca_annot_lower$Scate & Sca_DEGs2$Scate %in% Cvi_immune_annot2$Scate,]
DEG_delet_immune$category<- "Deleterious + Immune"
DEG_delet<- Sca_DEGs2[Sca_DEGs2$Scate %in% Sca_annot_lower$Scate & !Sca_DEGs2$Scate %in% Cvi_immune_annot2$Scate,]
DEG_delet$category<- "Deleterious"

#Cvi_immune_annot[Cvi_immune_annot$Scate=="Scate05874-RA",]
#Scate05874-RA
#augustus_masked-scaffold-un935-processed-gene-0.0
```

```{r}
source("config.R")
DEG_fitness<- rbind(DEG_adaptive_immune, DEG_adaptive, DEG_delet, DEG_delet_immune)
DEG_fitness$combo<- paste0(DEG_fitness$Scate,"_",DEG_fitness$Scate_annot_GeneName," (",DEG_fitness$geneID,"_",DEG_fitness$Anolis_Genename,")")
DEG_fitness_unique<- DEG_fitness[!duplicated(DEG_fitness$Scate),]
DEG_fitness$Note<- unlist(DEG_fitness$Note)
#write.csv(DEG_fitness, file="../data/lists/fitness_DEGs_annot.csv", row.names=FALSE, quote=FALSE)
g<- ggplot(DEG_fitness, aes(x=category, y=combo, fill=log2FoldChange)) + geom_tile() + scale_fill_viridis_c() + theme(axis.text.x = element_text(angle=90, hjust=1, vjust=0.5)) + peri_tile + labs(y="",x="") + coord_fixed()
g
outdir="../results/"
ggsave(filename=paste0(outdir,"tempfigure_selectionDEGs.png"), plot=g, width=200, height=150, units="mm", device="png", bg="white", dpi="print")
```



# create list files for all interest genes: 

```{r genelists, eval=FALSE}
#OG0000003 is an MHC annotation - how many genes total are in this orthogroup
nrow(ortho_dispersed_annot[which(ortho_dispersed_annot$Orthogroup=="OG0000003"),])
## 222 rows. 
#how many Cortalus IDs
length(unique(ortho_dispersed_annot$geneID[which(ortho_dispersed_annot$Orthogroup=="OG0000003")]))
length(unique(ortho_dispersed_annot$Scate[which(ortho_dispersed_annot$Orthogroup=="OG0000003")]))

#Snpsift takes gene IDs...
#output.file <- file(paste0("../data/lists/genelist_lower10.txt"), "wb")
#write.table(unique(Sca_annot_lower$Scate),file=output.file,row.names = FALSE,
#            col.names = FALSE,
#            quote = FALSE,
#            append = FALSE,
#            sep = "")
#close(output.file)
#
#output.file <- file(paste0("../data/lists/genelist_upper10.txt"), "wb")
#write.table(unique(Sca_annot_upper$Scate),file=output.file,row.names = FALSE,
#            col.names = FALSE,
#            quote = FALSE,
#            append = FALSE,
#            sep = "")
#close(output.file)

output.file <- file(paste0("../data/lists/genelist_SFD_DEGs.txt"), "wb")
write.table(unique(Sca_DEGs$Scate),file=output.file,row.names = FALSE,
            col.names = FALSE,
            quote = FALSE,
            append = FALSE,
            sep = "")
close(output.file)


output.file <- file(paste0("../data/lists/genelist_immune2.txt"), "wb")
write.table(unique(Cvi_immune_annot2$Scate),file=output.file,row.names = FALSE,
            col.names = FALSE,
            quote = FALSE,
            append = FALSE,
            sep = "")
close(output.file)


genes_of_interest<- unique(c(Cvi_immune_annot2$Scate, Sca_DEGs$Scate,Sca_annot_upper$Scate,Sca_annot_lower$Scate))

output.file <- file(paste0("../data/lists/genelist_therest.txt"), "wb")
write.table(unique(Sca_annot$Scate)[!Sca_annot$Scate %in% genes_of_interest],file=output.file,row.names = FALSE,
            col.names = FALSE,
            quote = FALSE,
            append = FALSE,
            sep = "")
close(output.file)

output.file <- file(paste0("../data/lists/genelist_dos.txt"), "wb")
write.table(unique(DoS$geneID),file=output.file,row.names = FALSE,
            col.names = FALSE,
            quote = FALSE,
            append = FALSE,
            sep = "")
close(output.file)
```

```{r bed files, eval=FALSE}
#read in a fresh gff.
Sca_annot<- readGFF("../../reference_genome/Scate_HiC_rnd4.all.putative.function.gff")
nrow(Sca_annot)

Sca_annot<- as.data.frame(Sca_annot)
Sca_annot$bed<- paste0(Sca_annot$seqid,":", Sca_annot$start,"-", Sca_annot$end)
Sca_annot_CDS<- Sca_annot[grepl("cds",Sca_annot$ID),]
Sca_annot_CDS$geneID<- gsub("-RA:cds","",Sca_annot_CDS$ID)
Sca_annot_gene<- Sca_annot[Sca_annot$type=="gene",]

## DoS ##
DoS<- read.csv("../data/Mathur_2023_PNAS/DoS.KLDRig.csv")
DoS$Scate<- paste0(DoS$geneID,"-RA")
#DoS<- DoS[order(DoS$dos1, decreasing=TRUE),]
#DoSupper10<- DoS[which(DoS$dos1>quantile(DoS$dos1, 0.9)),]
#extract CDS for genes with DoS values.
DoS_bed_CDS<- Sca_annot_CDS[Sca_annot_CDS$geneID %in% DoS$geneID,]
DoS_bed_CDS<- DoS_bed_CDS[,c("seqid","start","end","geneID")]
#convert to 0 based coordinates system.
DoS_bed_CDS$start<- DoS_bed_CDS$start-1

output.file <- file(paste0("../data/beds/DoS_CDS_geneIDs.bed"), "wb")
write.table(DoS_bed_CDS,file=output.file,row.names = FALSE,
            col.names = FALSE,
            quote = FALSE,
            append = FALSE,
            sep = "\t")
close(output.file)

# upper 10
upper10_CDS_bed<- Sca_annot_CDS[Sca_annot_CDS$geneID %in% gsub("-RA","",Sca_annot_upper$Scate),]

upper10_CDS_bed<- upper10_CDS_bed[,c("seqid","start","end","geneID")]
upper10_CDS_bed$start<- upper10_CDS_bed$start-1 
  
output.file <- file(paste0("../data/beds/upper10_CDS_geneIDs.bed"), "wb")
write.table(upper10_CDS_bed,file=output.file,row.names = FALSE,
            col.names = FALSE,
            quote = FALSE,
            append = FALSE,
            sep = "\t")
close(output.file)

upper10_gene_bed<- Sca_annot_upper
upper10_gene_bed$chr<- word(upper10_gene_bed$bed,1,1, sep=":")
upper10_gene_bed$c<- word(upper10_gene_bed$bed,2,2, sep=":")
upper10_gene_bed$start<- as.numeric(word(upper10_gene_bed$c,1,1, sep="-"))-1
upper10_gene_bed$end<- as.numeric(word(upper10_gene_bed$c,2,2, sep="-"))
upper10_gene_bed<- upper10_gene_bed[,c("chr","start","end","Scate")]
upper10_gene_bed$Scate<- gsub("-RA","",upper10_gene_bed$Scate)

output.file <- file(paste0("../data/beds/upper10_genes_geneIDs.bed"), "wb")
write.table(upper10_gene_bed,file=output.file,row.names = FALSE,
            col.names = FALSE,
            quote = FALSE,
            append = FALSE,
            sep = "\t")
close(output.file)

# lower 10
lower10_CDS_bed<- Sca_annot_CDS[Sca_annot_CDS$geneID %in% gsub("-RA","",Sca_annot_lower$Scate),]

lower10_CDS_bed<- lower10_CDS_bed[,c("seqid","start","end","geneID")]
lower10_CDS_bed$start<- lower10_CDS_bed$start-1

output.file <- file(paste0("../data/beds/lower10_CDS_geneIDs.bed"), "wb")
write.table(lower10_CDS_bed,file=output.file,row.names = FALSE,
            col.names = FALSE,
            quote = FALSE,
            append = FALSE,
            sep = "\t")
close(output.file)



lower10_gene_bed<- Sca_annot_lower
lower10_gene_bed$chr<- word(lower10_gene_bed$bed,1,1, sep=":")
lower10_gene_bed$c<- word(lower10_gene_bed$bed,2,2, sep=":")
lower10_gene_bed$start<- as.numeric(word(lower10_gene_bed$c,1,1, sep="-"))-1
lower10_gene_bed$end<- as.numeric(word(lower10_gene_bed$c,2,2, sep="-"))
lower10_gene_bed<- lower10_gene_bed[,c("chr","start","end","Scate")]
lower10_gene_bed$Scate<- gsub("-RA","",lower10_gene_bed$Scate)


output.file <- file(paste0("../data/beds/lower10_genes_geneIDs.bed"), "wb")
write.table(lower10_gene_bed,file=output.file,row.names = FALSE,
            col.names = FALSE,
            quote = FALSE,
            append = FALSE,
            sep = "\t")
close(output.file)

### Remaining DEGs ###
not_fitness_DEG<- readLines("../data/lists/genelist_SFD_DEGs.txt")

DEG_CDS_bed<- Sca_annot_CDS[Sca_annot_CDS$geneID %in% not_fitness_DEG,]

DEG_CDS_bed<- DEG_CDS_bed[,c("seqid","start","end","geneID")]
DEG_CDS_bed$start<- DEG_CDS_bed$start-1 
  
output.file <- file(paste0("../data/beds/DEG_CDS_geneIDs.bed"), "wb")
write.table(DEG_CDS_bed,file=output.file,row.names = FALSE,
            col.names = FALSE,
            quote = FALSE,
            append = FALSE,
            sep = "\t")
close(output.file)


out<- list()
for(g in not_fitness_DEG){
  sub<- Sca_annot_gene[which(Sca_annot_gene$Name==g),]
  out[[g]]<- data.frame(chr=sub$seqid, start=sub$start, end=sub$end, Scate=sub$Name)
}
DEG_gene_bed<- do.call("rbind",out)
DEG_gene_bed$start<- DEG_gene_bed$start-1

output.file <- file(paste0("../data/beds/DEG_genes_geneIDs.bed"), "wb")
write.table(DEG_gene_bed,file=output.file,row.names = FALSE,
            col.names = FALSE,
            quote = FALSE,
            append = FALSE,
            sep = "\t")
close(output.file)
#####
upper<- readLines("../data/lists/genelist_upper10.txt")
upper<- gsub("-RA","",upper)
lower<- readLines("../data/lists/genelist_lower10.txt")
lower<- gsub("-RA","",lower)
dos<- c(upper,lower)
immune<- readLines("../data/lists/genelist_immune.txt")
immune<- gsub("-RA","",immune)
DEG<- readLines("../data/lists/genelist_SFD_DEGs.txt")
DEG<- gsub("-RA","",DEG)


genelist<- c(DEG,dos,immune)

#pick a random 100 genes.
Sca_annot_gene2<- Sca_annot_gene[which(!Sca_annot_gene$ID %in% genelist),]
Sca_random<- Sca_annot_gene2[sample(rownames(Sca_annot_gene2),100, replace=F),]

Sca_random<- Sca_random[,c("seqid","start","end","Name")]
Sca_random$start<- Sca_random$start-1
Sca_random<- Sca_random[order(Sca_random$seqid, Sca_random$start),]
output.file <- file(paste0("../data/beds/random_genes_geneIDs.bed"), "wb")
write.table(Sca_random,file=output.file,row.names = FALSE,
            col.names = FALSE,
            quote = FALSE,
            append = FALSE,
            sep = "\t")
close(output.file)

Sca_random_CDS<- Sca_annot_CDS[which(Sca_annot_CDS$geneID %in% Sca_random$Name),]
Sca_random_CDS<- Sca_random_CDS[,c("seqid","start","end","geneID")]
Sca_random_CDS$start<- Sca_random_CDS$start-1
Sca_random_CDS<- Sca_random_CDS[order(Sca_random_CDS$seqid, Sca_random_CDS$start),]
output.file <- file(paste0("../data/beds/random_CDS_geneIDs.bed"), "wb")
write.table(Sca_random_CDS,file=output.file,row.names = FALSE,
            col.names = FALSE,
            quote = FALSE,
            append = FALSE,
            sep = "\t")
close(output.file)

output.file <- file(paste0("../data/lists/genelist_random_geneIDs.bed"), "wb")
write.table(Sca_random$Name,file=output.file,row.names = FALSE,
            col.names = FALSE,
            quote = FALSE,
            append = FALSE,
            sep = "\t")
close(output.file)

```


# Single Copy Orthologs only. 

```{r}
ortho_dispersed_sca<- ortho_dispersed[which(ortho_dispersed$Scate!=""),]
ortho_summary<- ortho_dispersed_sca %>% group_by(Orthogroup) %>% 
summarise(n_Cviri=length(unique(geneID)),n_Scate=length(unique(Scate)))
ortho_summary<- ortho_summary[which(ortho_summary$n_Scate>0),]

multis<- ortho_summary[ortho_summary$n_Cviri>1 | ortho_summary$n_Scate>1,]

Sca_annot_lower_single<- Sca_annot_lower[!Sca_annot_lower$Orthogroup %in% multis$Orthogroup,]
Sca_annot_upper_single<- Sca_annot_upper[!Sca_annot_upper$Orthogroup %in% multis$Orthogroup,]
Cvi_immune_annot2_single<- Cvi_immune_annot2[!Cvi_immune_annot2$Orthogroup %in% multis$Orthogroup,]
Sca_DEGs_single<- Sca_DEGs[!Sca_DEGs$Orthogroup %in% multis$Orthogroup,]

SFD_DoS_overlap<- list(lower_DoS=Sca_annot_lower_single$Scate, upper_DoS=Sca_annot_upper_single$Scate, Sca_immune=unique(Cvi_immune_annot2_single$Scate), DEG_ortho=Sca_DEGs_single$Scate)
ggvenn(SFD_DoS_overlap) + labs(title="Overlap between PNAS DoS, Sca Immune Genes and Cvi DEGs single copy orthologs ONLY")
```


