---
title: "What do we know about genes underlying resistance or susceptibility to snake fungal disease"
title-block-banner: true
author:
  - name: Peri E. Bolton
    orcid: 0000-0002-2057-1973
    email: bolton.161@osu.edu
    affiliations: Department of Evolution, Ecology and Organismal Biology, The Ohio State University
  - name: Andrew Mason (OrthoFinder)
    email: mason.501@osu.edu
    affiliations: Department of Evolution, Ecology and Organismal Biology, The Ohio State University
  - name: Samarth Mathur (FST)
    email: mathur.112@osu.edu
    affiliations: Department of Evolution, Ecology and Organismal Biology, The Ohio State University
  - name: H. Lisle Gibbs
    email: gibbs.128@osu.edu
    affiliations: Department of Evolution, Ecology and Organismal Biology, The Ohio State University
date: '`r lubridate::today()`'
engine: knitr
format:
  html:
    code-fold: true
    code-tools: true
    code-line-numbers: true
    embed-resources: true
    html-math-method: mathml
    toc: true
    toc-depth: 3
    toc-expand: 2
    toc-location: left
    toc-title: "Contents"
    number-sections: true
    number-depth: 3
    theme: 
      - minty
    page-layout: full
    grid: 
      margin-width: 250px
output:
  html_document:
      df_print: paged
params:
  workdir: "/Users/bolton.161/OneDrive - The Ohio State University/Rattlesnakes/Scate_SFD/analysis"
  datadir: "/Users/bolton.161/OneDrive - The Ohio State University/Rattlesnakes/Scate_SFD/data"
  outdir: "/Users/bolton.161/OneDrive - The Ohio State University/Rattlesnakes/Scate_SFD/results"
bibliography: metadata/references.bib
csl: metadata/nature.csl
knitr:
  opts_chunk:
    fig.path: /Users/bolton.161/OneDrive - The Ohio State University/Rattlesnakes/Scate_SFD/results/report/figures/
---

\flushleft

```{r setup, message = FALSE, warning = FALSE, echo=FALSE}
require("knitr")
knitr::opts_chunk$set(message = FALSE, warning = FALSE, fig.pos="H", fig.width=12, fig.height=7)
path = dirname(params$workdir)
knitr::opts_knit$set(root.dir=path)
```

```{r}
#| label: config
#| echo: false
#setwd("G:/My Drive/Rattlesnakes/Scate_SFD/analysis")
#setwd("~/OneDrive - The Ohio State University/Rattlesnakes/Scate_SFD/analysis")
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggpubr)
library(ggrepel)
library(viridis)
#library(ggh4x)
library(ggvenn)
library(openxlsx)
library(rstatix)
library(gprofiler2)
library(reshape2)
library(scales)
library(rrvgo)
library(clusterProfiler)
library(openxlsx)
workdir<- "~/OneDrive - The Ohio State University/Rattlesnakes/Scate_SFD/analysis"
datadir<- "~/OneDrive - The Ohio State University/Rattlesnakes/Scate_SFD/data"
outdir<- "~/OneDrive - The Ohio State University/Rattlesnakes/Scate_SFD/results"

source(paste0(workdir,"/config.R"))
source(paste0(workdir,"/functions.R"))


MHC_paralogs<- c("Scate05828","Scate05830","Scate05831","Scate05832","Scate05833","Scate05835","Scate05836","Scate05837","Scate05838","Scate05839","Scate05840","Scate05841","Scate05844","Scate05845","Scate05854","Scate05855","Scate05856","Scate05857","Scate05858","Scate05859","Scate05860","Scate05861","Scate05862","Scate05863","Scate05865","Scate05866","Scate05867","Scate05868","Scate05869","Scate05871","Scate05872","Scate05873","Scate05874","Scate05875","Scate05876","Scate05877","Scate05878")
upper<- readLines(paste0(datadir,"/lists/genelist_upper10.txt"))
upper<- gsub("-RA","",upper)
lower<- readLines(paste0(datadir,"/lists/genelist_lower10.txt"))
lower<- gsub("-RA","",lower)
dos<- c(upper,lower)
immune_original<- readLines(paste0(datadir,"/lists/genelist_immune.txt"))
immune_original<- gsub("-RA","",immune_original)
immune<- readLines(paste0(datadir,"/lists/genelist_immune2.txt"))
immune<- gsub("-RA","",immune)
immune<- c(immune,	"Scate15657") #for some reason TLR5 didn't make it into V2... 
DEG<- readLines(paste0(datadir,"/lists/genelist_SFD_DEGs.txt"))
DEG<- gsub("-RA","",DEG)
random<- readLines(paste0(datadir,"/lists/genelist_random_geneIDs.txt"))



mass<- c("KLDR","BERG","BPNP","CCRO","CEBO","JENN","KBPP","MOSQ","PRDF","ROME","SPVY","SSSP","WLRD")
pop3<- c("STER","KLDR","CEBO")
pops<- c("STER","KLDR")


viridispairs<- c("#440154","#63017b","#8102a0","#125750","#19786e","#1f968a","#2a3f68","#355084","#5779bb","#3bae4c","#40bd52","#69cc78")

keep<- c("keep","pops","mass","outdir","datadir","DEG","immune","lower","upper","hwecols","picols", "basesize", "baselwd","pop3", "gene_categories")

```

# Abstract

Genes can contribute to whether a host is susceptible or recovers from a disease. Variation in these genes can be important for how a population responds to current and novel pathogens. However, experiments to characterize these genes are unethical to conduct in endangered species. Moreover, pathogens with sub-lethal effects, difficult diagnosis and/or hosts that are difficult to follow in the wild means that natural experimental data is scarce or absent. This is the case for snake fungal disease (caused by *Ophidiomyces ophidiicola*) in endangered Eastern massasauga rattlesnakes (*Sistrurus catenatus*). Here, we suggest that genes experimentally shown to be inolved in disease process in *Crotalus viridis* may provide a useful backdoor to narrowing down the set of SFD relevant genes, particularly those with signatures of selection within exposed species. We conduct diversity and SFS-based neutrality tests on a set of 1048 (actually a subset for this report) orthologs in two threatened *S. catenatus* populations in Ohio, and compare with outbred sister species *S. tergeminus*. We show that diversity in these genes is reduced in *S. catenatus*, particularly in populations with small N\[e\] (CEBO). However, MHC Class I paralogs are substantially more diverse than than others. Further, we identify a set of genes that appear to have non-synonymous sites under divergent selection in across the sampled *S. catenatus* populations.

# Methods

## Candidate gene selection

We reasoned that genes that were functionally relevant to snake fungal disease resiliance/susceptibility in *Sistrurus catenatus* were from three main categories (@fig-venn)

-   Genes with known immune system function

-   Those genes previously identified as involved in disease process in *Crotalus viridis* by differential expression analysis [@mathur2024]

-   Genes under selection in *S. catenatus.* These genes were identified using the upper and lower 10% direction of selection (DoS) statistics on divergence and polymorphism data in synonymous and non-synonymous sites [@mathur2023]

We reasoned that genes in multiple functional categories are more likely to be relevant to disease susceptibility/resistance than those involved in disease process alone. And that genes which show some evidence of on-going selection might be particularly important.

We used `OrthoFinder` to identify orthologs between *C. viridis* and *S. catenatus* genome annotations. Then we linked the differentially expressed genes from *C. viridis* to its ortholog in *S. catenatus*, as a putative set of disease process genes in *S. catenatus*. The 855 unique DEGs in *C. viridis* corresponded to 1048 unique genes in *S. catenatus*.

Our candidate set of immune genes were broadly based on immune and inflammatory processes. Immune genes were identified in *C. viridis* based on GO and KEGG terms for the three species used for functional annotation: *Anolis carolinensis*, *Python vittatus* and *Thamnophis sirtalis*. These genes were based on a set of GO terms derived from a GOSlim query of GO:0002376 (Immune system process), GO:0006954 (Inflammation) and the set of immune GO terms used in [@silver2022], and KEGG annotations in these species. The specific query terms are listed in [Table XX]{style="color:red"}. The genes for these species associated with these terms were downloaded using R packages `protti::fetch_quickgo` and `KEGGREST::keggGet`. Then, any immune genes from these species were considered a potential immune gene in *C. viridis*, and by orthology *S. catenatus*.

```{r}
#| label: fig-venn
#| fig.width: 7
#| fig.height: 4
#| fig-cap: "Eastern Massasauga Rattlesnake genes in overlapping and putatively functional categories. It is these functional categories that we will apply throughout"

genes<- list("upper DoS"=upper, "lower DoS"=lower, "immune genes"=immune, "DEG ortholog"=DEG)
#genes<- list("immune genes"=immune, "immune genes2"=immune2,"DEG ortholog"=DEG)
ggvenn(genes, set_name_size = 6, text_size=5, stroke_size=0.4) + coord_cartesian(clip="off")

genes<- list("immune genes"=immune, "SFD DEG"=DEG)
g<- ggvenn(genes, set_name_size = 6, text_size=5, stroke_size=0.4, show_percentage = FALSE) + coord_cartesian(clip="off")
ggsave(filename=paste0(outdir,"/poster/genes_venn.pdf"), plot=g, device="pdf", height=, width=)
```

## Selection in disease process genes.

For this preliminary analysis we analysed all 31 DEGs that were identified as under selection in *S. catenatus* [@mathur2023], a subset of 100 DEGs, and a random set of 100 genes from the genome. We compared the largest *S. catenatus* population at Killdeer Plains Wildlife Area, OH (KLDR), to a small population at Cedar Bog, OH (CEBO). We further compared patterns against outbred sister species *S. tergeminus*, mapped to the same reference *S. catenatus* reference genome. We wanted to explore whether selection was ongoing in an outbred species, to compare the the effects in an outbred species with the bottlenecked *S. catenatus*.

To explore whether there was evidence of selection, we used `angsd v0.940` to calculate nucleotide diversity (π), heterozygosity and Tajima's D separately for synonymous and non-synonymous sites per focal gene. We labelled degenerate sites using python script `degenotate v1.2.4`[@sackton2024]. 2- and 3-fold degenerate sites were labelled as non-synonymous if a variant was considered as such by `SnpEff v5.2` in a genotyped file for another project (Mathur et al Unpubl, [@cingolani2012program]).

```{bash}
#| label: degenotate.slurm
#| eval: false
#!/bin/bash
#SBATCH --nodes=1
#SBATCH --cpus-per-task=4
#SBATCH --time=12:00:00
#SBATCH --account=
#SBATCH --job-name=job_degenotate
#SBATCH --output=job_degenotate-%j
#SBATCH --mail-type=begin        # send email when job begins
#SBATCH --mail-type=end          # send email when job ends
#SBATCH --mail-user=bolton.161@osu.edu


source ~/miniconda3/bin/activate ~/miniconda3/envs/degenotate/
INDIR=/fs/ess/scratch/PAS1533/peri/Scate_SFD
DEGENOTATE=/users/PAS1533/boltonp88/local/src/degenotate

python $DEGENOTATE/degenotate.py -a $INDIR/ref/Scate_HiC_rnd4.all.putative.function.gff -g $INDIR/ref/Scatenatus_HiC_v1.1.fasta -o $INDIR/ref/degenotate/Scate.degeneracy

```

```{bash}
#| label: job_fgd_degen
#| eval: false

#!/bin/bash
#SBATCH --nodes=1
#SBATCH --ntasks-per-node=1
#SBATCH --time=24:00:00
#SBATCH --cpus-per-task=1		# cpu-cores per task (>1 if multi-threaded tasks)
#SBATCH --account=PAS1533
#SBATCH --array=1-14
#SBATCH --job-name=job_fgd_div_DEG
#SBATCH --output=job_fgd_div_DEG-%A.%a.out
#SBATCH --mail-type=begin		# send email when job begins
#SBATCH --mail-type=end		  # send email when job ends
#SBATCH --mail-user=bolton.161@osu.edu

## DESCRIPTION ##
# This code is after trying JUST the VCFtools variants using the method from 
# Mathur et al 2023 PNAS. I wasn't happy with using the same denominator (CDS) 
# for ns and s sites, so I have incorporated this into this code
# Further, Lisle wanted to see number of "derived" alleles 
# Thus, I have calculated these with respect to the S. terg pseudoreference
# which incorporated major non-ref allelic variants from Sterg into the Scate genome.
# Counting genotypes is done with angsd parameters -doMajorMinor 4 -doHwe 1
# ALSO only want to include genes that are selected DEGs OR selected immune DEGs (30 genes).

#### NOTE: 20240618 - updated job submission format to work with 1003 genes remaining in DEGs. 

ANGSD=/users/PAS1533/boltonp88/local/src/angsd
INDIR=/fs/ess/scratch/PAS1533/peri/Scate_SFD


config=$INDIR/lists/final_172_pops.noabc.list
#config=$INDIR/test_pops.txt #2 pops

POP=$(awk -v num="$SLURM_ARRAY_TASK_ID" 'FNR==num {print $1}' ${config})

##PATHS

#make a dir for the angsd outputs
mkdir $INDIR/VOI/pi_DEG/$POP
OUTDIR=$INDIR/VOI/pi_DEG/$POP

PSREF=$INDIR/ref/Scatenatus_V1.1_STERpseudoref.fasta
BEDDIR=$INDIR/beds
BAMLISTDIR=$INDIR/lists/bamlists
BAMDIR=$INDIR/bams

##LISTS

#list of names in the population
poplist=$INDIR/lists/final_172_${POP}.txt

rm $BAMLISTDIR/$POP.bamlist

#make a bamlist for the population
cat $poplist | while read line; do echo ${BAMDIR}"/"${line}"_mark_dups.sorted.bam" >> $BAMLISTDIR/$POP.bamlist ; done

bamlist=$BAMLISTDIR/${POP}.bamlist

genelist=$INDIR/lists/genelist_other_DEG1.txt

#genes_bed=$BEDDIR/test_genes.bed
genes_bed=$BEDDIR/DEG_genes_geneIDs.bed
cds_bed=$BEDDIR/DEG_CDS_geneIDs.bed

chunksize=20
end=4
prev_job_ids=""
cpus=8

for ((i=0; i<=end; i++))
do
	TIME=$(date '+%H:%M:%S')
	batchname=$((i + 1))
	echo starting batch $batchname at $TIME
	start_line=$((i * chunksize + 1))
	head -n $((start_line + chunksize - 1)) "$genelist" | tail -n $chunksize > "${OUTDIR}/genelist_batch_${batchname}.txt"
	
	infile="${OUTDIR}/genelist_batch_${batchname}.txt"
	job_ids=""

	while read g; do
	
		TIME=$(date '+%H:%M:%S')
		
		script_name=$INDIR/VOI/pi_DEG/genejobs/${POP}.${batchname}.${g}.sh
		
		echo building ${POP}.${batchname}.${g}.sh script at $TIME
		echo
		
		  echo "#!/bin/sh -l 
#SBATCH --nodes=1
#SBATCH --ntasks-per-node=1
#SBATCH --time=3:00:00
#SBATCH --cpus-per-task=8		# cpu-cores per task (>1 if multi-threaded tasks)
#SBATCH --mem-per-cpu=8G		 # memory per cpu-core
#SBATCH --account=PAS1533
#SBATCH --job-name=${POP}.${batchname}.${g}.fgd.angsd.DEG
#SBATCH --output=${POP}.${batchname}.${g}.fgd.angsd-%A.%a.out
		
		
		source ~/miniconda3/bin/activate ~/miniconda3/envs/angsd/
		
		BAMLISTDIR=$INDIR/lists/bamlists
		
		bamlist=$BAMLISTDIR/${POP}.bamlist
		
		
		#get the first line of the genes names file
		
		echo "this is ${g}"
		echo
		echo "finding bedfiles for $g"
		#subset the bed files for that gene
		grep $g $genes_bed > $BEDDIR/${POP}.${g}_gene.bed
		grep $g $cds_bed > $BEDDIR/${POP}.${g}_cds.bed
		
		echo
		echo "now lets cut these bams into smaller bams $g"
		
		while read s; do
			
			sample=\"\$(basename \"\$s\")\"
			sample=\"\$(echo \"\$sample\" | sed 's/_mark_dups.sorted.bam//')\"
			
			
			samtools view \$s \
			-b -h -o $BAMDIR/ingene/${POP}.\${sample}.in.${g}.genes.bam \
			-L $BEDDIR/${POP}.${g}_gene.bed
			
			samtools sort -@ ${cpus} -T $BAMDIR/ingene/${POP}.\${sample}.in.${g}.genes.bam.temp \
			-O BAM -o $BAMDIR/ingene/${POP}.\${sample}.in.${g}.genes.sorted.bam \
			$BAMDIR/ingene/${POP}.\${sample}.in.${g}.genes.bam 
			
			samtools index -@ ${cpus} -b $BAMDIR/ingene/${POP}.\${sample}.in.${g}.genes.sorted.bam
			
			#make a list with all those files for angsd
			echo $BAMDIR/ingene/${POP}.\${sample}.in.${g}.genes.sorted.bam \
			>> $BAMLISTDIR/${POP}.${g}.list
			
			
			echo "processed \$sample for ${g} and put into file ${BAMLISTDIR}/${POP}.${g}.list"
			
		done < $bamlist
		
		bamlist_gene=$BAMLISTDIR/${POP}.${g}.list
		
		echo \$bamlist_gene
		
		for e in 0 2 3 4; do
			
			echo "doing $g \$e sites"
			echo 
			echo "intersecting \${e}th positions $g"
			
			#bedtools intersect \
			#-b $BEDDIR/${POP}.${g}_cds.bed \
			#-a $BEDDIR/degen_sites_\${e}.bed \
			#-wa > $BEDDIR/${POP}.${g}.degen\${e}.positions.bed
			
			grep $g $BEDDIR/degen_sites_\${e}.bed > $BEDDIR/${POP}.${g}.degen\${e}.positions.bed
			
			echo "making pos file \${e}th $g"
			echo
			
			if [[ \$e -eq 2 ]] || [[ \$e -eq 3 ]]; then 
		
				echo "\${e}th degen sites $g overlap with known missense sites class as nonsyn"
				echo
				bedtools intersect \
				-a $BEDDIR/${POP}.${g}.degen\${e}.positions.bed \
				-b $BEDDIR/final172.allchr.finalSNPs.missense.positions.bed \
				-wa >> $BEDDIR/${POP}.${g}.ns.positions.bed
				
				#-v Only report those entries in A that have no overlap in B. 
				bedtools intersect \
				-a $BEDDIR/${POP}.${g}.degen\${e}.positions.bed \
				-b $BEDDIR/final172.allchr.finalSNPs.missense.positions.bed \
				-v >> $BEDDIR/${POP}.${g}.syn.positions.bed
		
			
			elif [[ \$e -eq 0 ]]; then
				
				echo "\$e $g all possible changes nonsynonymous"
				cat $BEDDIR/${POP}.${g}.degen\${e}.positions.bed >> $BEDDIR/${POP}.${g}.ns.positions.bed
				
				
			elif [[ \$e -eq 4 ]]; then
				echo "\$e $g all possible changes synonymous"
				cat $BEDDIR/${POP}.${g}.degen\${e}.positions.bed >> $BEDDIR/${POP}.${g}.syn.positions.bed
			
			fi
				
			echo "\$e $g remove degenerate sites bed"
			echo
			rm $BEDDIR/${POP}.${g}.degen\${e}.positions.bed
				
		done
			
		for e in ns syn; do
		
			echo "\$e $g sort bed"
			sort -k 1,1 -k2,2n $BEDDIR/${POP}.${g}.\${e}.positions.bed \
			> ${POP}.${g}.\${e}.tmp && \
			mv ${POP}.${g}.\${e}.tmp $BEDDIR/${POP}.${g}.\${e}.positions.bed
			
			echo
			echo "\$e $g make pos file"
			awk '{print \$1\"\\t\"\$3}' \
			$BEDDIR/${POP}.${g}.\${e}.positions.bed \
			> $BEDDIR/${POP}.${g}.\${e}.positions.txt
			
			echo
			echo "\$e $g angsd index"
			$ANGSD/angsd sites index $BEDDIR/${POP}.${g}.\${e}.positions.txt
			
			echo
			echo "angsd calc $g \$e"
			
			#applying some extra quality filters here
			
			$ANGSD/angsd -P ${cpus} -dosaf 1 -GL 1 \
			-anc $INDIR/ref/Scatenatus_HiC_v1.1.fasta \
			-ref $INDIR/ref/Scatenatus_HiC_v1.1.fasta \
			-doMajorMinor 4 -doHWE 1 -minHetFreq -1 \
			-remove_bads 1 -uniqueOnly 1 \
			-minMapQ 20 -minQ 20 \
			-bam \$bamlist_gene \
			-sites $BEDDIR/${POP}.${g}.\${e}.positions.txt \
			-out $OUTDIR/${POP}.${g}.\${e}
			
			echo "calc SFS $g \$e"
			
			$ANGSD/misc/realSFS -P ${cpus} -fold 1 \
			$OUTDIR/${POP}.${g}.\${e}.saf.idx -maxIter 100 \
			> $OUTDIR/${POP}.${g}.\${e}.sfs
			
			#thetas
			echo "calc thetas $g \$e"
			
			$ANGSD/misc/realSFS saf2theta \
			$OUTDIR/${POP}.${g}.\${e}.saf.idx \
			-sfs $OUTDIR/${POP}.${g}.\${e}.sfs -fold 1 \
			-outname $OUTDIR/${POP}.${g}.\${e}
			
			$ANGSD/misc/thetaStat do_stat \
			$OUTDIR/${POP}.${g}.\${e}.thetas.idx \
			-outnames $OUTDIR/${POP}.${g}.\${e}
		done

		
			echo
			echo all done 
			" \
			> $script_name
			
			cd $INDIR/VOI/pi_DEG/genejobs/
		
		
			TIME=$(date '+%H:%M:%S')
			
      if [[ -n $(ls -l $INDIR/VOI/pi_DEG/${POP}/${POP}.${g}.*pestPG) ]]; then
        echo ${POP}.${g} exists not submitting
      else
        job_id=$(sbatch $script_name)
        
        
        job_id=$(echo $job_id | awk '{print $4}')
        
        echo submitting $script_name at $TIME as job_id $job_id
        echo
        
        job_ids="${job_ids}:${job_id}"
      fi
			  
	done < $infile

	prev_job_ids="${job_ids#:}"
	echo previous job_IDs $prev_job_ids
	echo
	
	while [[ -n $(squeue -u boltonp88 -j $(echo $prev_job_ids | tr ':' ',') | awk '{print $1}' | tail -n +2 ) ]]; do
		#while squeue -u boltonp88 -j $(echo $prev_batch_job_ids | tr ':' ',') > /dev/null; do	
		echo "Waiting for batch $batchname jobs to complete..."
		echo 
        #echo $(squeue -j $(echo $prev_batch_job_ids | tr ':' ','))
		echo 
		echo checking grep $(squeue -u boltonp88 -j $(echo $prev_job_ids | tr ':' ',') | awk '{print $1}' | tail -n +2)
		sleep 15m
	done

	TIME=$(date '+%H:%M:%S')
	echo done batch $batchname at $TIME
	
done

#clean up
rm ${BEDDIR}/${POP}*
rm ${BAMDIR}/ingene/${POP}*
rm ${BAMLISTDIR}/${POP}*
```

We compared non-synonymous site diversity to synonymous site diversity, and to diversity in 1000 gene-desert regions on assembled chromosomes with a non-overlapping window-size of 25,000 bp. Gene-desert regions were at least 500 kb up- and down-stream from coding sequences [@mathur2023].

```{bash}
#| label: job_genedesert_div.slurm
#| eval: FALSE

#!/bin/bash
#SBATCH --nodes=1
#SBATCH --ntasks-per-node=1
#SBATCH --time=48:00:00
#SBATCH --cpus-per-task=8       # cpu-cores per task (>1 if multi-threaded tasks)
#SBATCH --mem-per-cpu=12G         # memory per cpu-core
#SBATCH --account=PAS1533
#SBATCH --array=1-13
#SBATCH --job-name=job_genedesert_div
#SBATCH --output=job_genedesert_div-%A.%a.out
#SBATCH --mail-type=begin        # send email when job begins
#SBATCH --mail-type=end          # send email when job ends
#SBATCH --mail-user=bolton.161@osu.edu


source ~/miniconda3/bin/activate ~/miniconda3/envs/angsd/
ANGSD=/users/PAS1533/boltonp88/local/src/angsd
module load samtools

INDIR=/fs/ess/scratch/PAS1533/peri/Scate_SFD

config=$INDIR/lists/final_172_pops.noabc.list
#config=$INDIR/test_pops.txt #2 pops

POP=$(awk -v num="$SLURM_ARRAY_TASK_ID" 'FNR==num {print $1}' ${config})

##PATHS

#make a dir for the angsd outputs


BAMINDIR=$INDIR/bams
BAMOUTDIR=$INDIR/geneDesert/bams
OUTDIR=$INDIR/geneDesert/angsd_out
BEDDIR=$INDIR/beds
BAMLISTDIR=$INDIR/lists/bamlists
PSREF=$INDIR/ref/Scatenatus_V1.1_STERpseudoref.fasta

##LISTS

#list of names in the population
poplist=$INDIR/lists/final_172_${POP}.txt

#make a bamlist for the population #presumably this is already done from the fgd one.
cat $poplist | while read line; do echo ${BAMINDIR}"/"${line}"_mark_dups.sorted.bam" >> $BAMLISTDIR/${POP}.gd.bamlist ; done

bamlist=$BAMLISTDIR/${POP}.gd.bamlist


while read s; do

  sample=$(basename $s)
  sample=$(echo $sample | sed 's/_mark_dups.sorted.bam//')
  
  #this removes the genes from the dataset. 
  samtools view $s -@ 8 \
  -b -h -o $BAMOUTDIR/exclude/${sample}.inGene.bam \
  -U $BAMOUTDIR/include/${sample}.outGene.bam \
  -L $BEDDIR/Scate_HiC.sortedgff_justgenes.bed

  samtools view  -@ 4 -b -h \
  -o $BAMOUTDIR/exclude/${sample}.inGene500.bam \
  -U $BAMOUTDIR/include/${sample}.outGene500.bam \
  -L $BEDDIR/genedesert.500kb.updown.tab.bed \
  $BAMOUTDIR/include/${sample}.outGene.bam
  
   
  samtools sort -@ 8 -T $BAMOUTDIR/include/${sample}.outGene500.bam.temp \
  -O BAM -o $BAMOUTDIR/include/${sample}.outGene500.sorted.bam \
  $BAMOUTDIR/include/${sample}.outGene500.bam
      
  samtools index -@ 8 -b $BAMOUTDIR/include/${sample}.outGene500.sorted.bam

  #make a list with all those files for angsd
  echo $BAMOUTDIR/include/${sample}.outGene500.sorted.bam \
  >> $BAMLISTDIR/${POP}.genedesert.list

done < $bamlist

bamlist=$BAMLISTDIR/${POP}.genedesert.list

$ANGSD/angsd -P 8 -dosaf 1 -GL 1 \
-doMajorMinor 4 -doHWE 1 -minHetFreq -1 \
-remove_bads 1 -uniqueOnly 1 \
-minMapQ 20 -minQ 20 \
-anc $PSREF \
-ref $PSREF \
-bam $bamlist 
-out $OUTDIR/${POP}.genedesert.sites

echo calculate SFS

$ANGSD/misc/realSFS -P 8 \
$OUTDIR/${POP}.genedesert.sites.saf.idx -maxIter 100 \
> $OUTDIR/${POP}.genedesert.sites.sfs

echo calc thetas

$ANGSD/misc/realSFS saf2theta \
$OUTDIR/${POP}.genedesert.sites.saf.idx \
-sfs $OUTDIR/${POP}.genedesert.sites.sfs \
-outname $OUTDIR/${POP}.genedesert.sites

echo do_stat

$ANGSD/misc/thetaStat do_stat \
$OUTDIR/${POP}.genedesert.sites.thetas.idx \
-outnames $OUTDIR/${POP}.genedesert.sites

echo persite theta

$ANGSD/misc/thetaStat print \
$OUTDIR/${POP}.genedesert.sites.thetas.idx 
> $OUTDIR/${POP}.persite.sites.theta

echo windowed theta

$ANGSD/misc/thetaStat do_stat \
$OUTDIR/${POP}.genedesert.sites.thetas.idx \
-win 25000 -step 25000 \
-outnames $OUTDIR/${POP}.genedesert.sites.thetasWindow.gz

#clean up
rm $BAMOUTDIR/exclude/*.bam
rm $BAMLISTDIR/${POP}.gd.bamlist
rm $BAMLISTDIR/${POP}.genedesert.list
```

Then we used `R v.4.4.0` to further analyse the diversity data [@rcoreteam2024]. We compared category means using non-parametric Wilcoxon rank-sum tests and Holm-Bonferroni adjustment for multiple tests. Genes that were outside the 5th and 95th percentile were considered outliers and likely candidates for ongoing selection.

## Local Adaptation - F~ST~

```{r}
#| label: calc-fst
#| eval: false
#| echo: false


library(vcfR)
library(hierfstat)
library(dplyr)
library(adegenet)

### making nonSyn.snps dataframe ####
deg_genes<- read.table("../beds/DEG_genes_geneIDs.bed")
colnames(deg_genes)<- c("CHR","start","end","geneID")
vcf_pos<- read.table("final152.frqfilt.DEG_CDS.ns.pos.txt")
colnames(vcf_pos)<- c("CHR","position")
nonSyn.snps<- list()
for(g in deg_genes$geneID){
  #extent<- deg_genes[which(deg_genes$geneID==g),]
  extent<- deg_genes %>%
          filter(geneID==g)
  #positions<- vcf_pos[which(vcf_pos$CHR==extent$CHR),]
  positions<- vcf_pos %>%
              filter(CHR=extent$CHR & position>=extent$start & position<=extent$end)
  #positions<- positions[which(positions$position>=extent$start & positions$position<= extent$end),]
  if(nrow(positions)==0){
  }else{
  positions$geneID<- g
  nonSyn.snps[[g]]<- positions
  }
}
#nonSyn.snps<- do.call("rbind",nonSyn.snps)
nonSyn.snps<- bind_rows(nonSyn.snps)
nonSyn.snps$snpID<- paste(nonSyn.snps$CHR,nonSyn.snps$position, sep="_")
nonSyn.snps$snpID<- gsub("-",".",nonSyn.snps$snpID)
write.csv(nonSyn.snps,"final152.frqfilt.DEG_CDS.ns.geneIDs.csv", row.names=FALSE,quote=F)


### making sampleINFO dataframe ####
mass<- c("KLDR","BERG","BPNP","CCRO","CEBO","JENN","KBPP","MOSQ","PRDF","GRLL","ROME","SPVY","SSSP","WLRD")
sampleInfo<- list()
for(p in mass){
samples<- readLines(paste0("../lists/final_172_",p,".txt"))
sampleInfo[[p]]<- data.frame(SampleID=samples,Pop=p)
}
sampleInfo<- do.call("rbind",sampleInfo)

### reading VCF data ####
deg.vcf <- read.vcfR("final152.frqfilt.DEG_CDS.ns.vcf")

#dim(deg.vcf)
#variants fix_cols  gt_cols 
#5073        8      153 
sampleIds <- data.frame(SampleID=colnames(deg.vcf@gt)[-1]) %>% inner_join(sampleInfo)

deg.genind <- vcfR2genind(deg.vcf)
pop(deg.genind) <- sampleIds$Pop

#### calculate Fstats ####

deg.bstat<-basic.stats(deg.genind, diploid=TRUE, digits=3)

deg.fst <- deg.bstat$perloc %>%
  as.data.frame() %>%
  filter(!is.na(Fst))

##FOR STUPID REASONS TIDYVERSE CAN'T BE INSTALLED ON OSC right now... so cannot use below in one go
#rownames_to_column(var = "snpID") %>%
deg.fst$snpID<- rownames(deg.fst)
deg.fst<- deg.fst %>%
  left_join(nonSyn.snps, by="snpID") %>%
  group_by(geneID) %>%
  mutate(meanFst = mean(Fst)) %>%
  as.data.frame()

#write.csv(deg.fst,"deg.fst.csv",sep="/"),quote=F,row.names=F)
```

# Results & Discussion

## Within-population tests.

```{r heterozygosity noeval}
#| label: heterozygosity reading
#| message: false
#| results: hide
#| echo: false
#| eval: false
hwe.degen<- read_hwe.gz(path=datadir, pops=pop3)
t<- hwe.degen[which(hwe.degen$geneID=="Scate00562" & hwe.degen$pop=="KLDR"),]
hwe<- hwe.degen %>% 
  group_by(pop,type,geneID,n_ind) %>% 
  summarise(total_hom=sum(n_Homoz_minor, n_Homoz_major),
            total_het=sum(Hobs_n),
            total_genotypes=sum(n_Homoz_minor, n_Homoz_major, Hobs_n),
            total_sites=n(), 
            total_polymorphic=sum(polymorphic),
            n_Homoz_minor=sum(n_Homoz_minor),
            total_exp_het=sum(Hexp_n)) %>%
  mutate(heterozygosity=total_het/total_genotypes,
         minor_homoz=n_Homoz_minor/total_genotypes,
         exp_heterozygosity=total_exp_het/total_genotypes,) 
#checking to make sure there aren't any genes that are universally not polymorphic.
actually_polymorphic<- hwe %>% 
  group_by(geneID) %>%
  summarise(n_polymorphic=sum(total_polymorphic),
            pc_polymorphic=sum(total_polymorphic)/total_genotypes)

#add gene categories
hwe<- gene_categories(dat=hwe,geneID)
hwe<- hwe %>% 
  mutate(immune_DEG=ifelse(immune_DEG=="" | immune_DEG=="immune","random",as.character(immune_DEG))) %>%
  mutate(immune_DEG=factor(immune_DEG,levels=c("DEG","lower DEG","upper DEG","immune DEG","lower immune DEG","upper immune DEG","random")))
hwe<- plot_categories(hwe,pop3)

hwe<- hwe %>%
  mutate(pop_site=factor(pop_site, levels=c("STER ns","STER syn","KLDR ns","KLDR syn","CEBO ns","CEBO syn"))) 
write.csv(hwe, paste0(datadir,"/FGD/degeneracy/heterozygosity_per_gene.csv"), row.names=FALSE, quote=FALSE)
```

```{r heterozygosity n s}
#| label: heterozygosity-ns
#| message: false
#| results: hide
#| echo: false

hwe<- read.csv(paste0(datadir,"/FGD/degeneracy/heterozygosity_per_gene.csv"))
hwe$immune_DEG<- ifelse(grepl("upper", hwe$immune_DEG),
                        gsub("upper ","", hwe$immune_DEG),
                        as.character(hwe$immune_DEG))
hwe$immune_DEG<- ifelse(grepl("lower", hwe$immune_DEG),
                        gsub("lower ","", hwe$immune_DEG),
                        as.character(hwe$immune_DEG))
het.summary<- hwe %>% 
  group_by(pop,geneID,immune_DEG,n_ind) %>% 
  reframe(n_het.ns=total_het[which(type=="non-synonymous")],
            n_het.syn=total_het[which(type=="synonymous")],
            heterozygosity.ns=round(heterozygosity[which(type=="non-synonymous")],5),
          inbrF.ns=1-(round(heterozygosity[which(type=="non-synonymous")],5)/round(exp_heterozygosity[which(type=="non-synonymous")],5)),
          
            heterozygosity.syn=round(heterozygosity[which(type=="synonymous")],5),
          inbrF.syn=1-(round(heterozygosity[which(type=="synonymous")],5)/round(exp_heterozygosity[which(type=="synonymous")],5)),
            minor_homoz.ns=round(minor_homoz[which(type=="non-synonymous")],5),
            minor_homoz.syn=round(minor_homoz[which(type=="synonymous")],5),
            n_genotypes.ns=total_genotypes[which(type=="non-synonymous")],
            n_genotypes.syn=total_genotypes[which(type=="synonymous")],
            n_sites.ns=total_sites[which(type=="non-synonymous")],
            n_sites.syn=total_sites[which(type=="synonymous")],
            n_polymorphic.ns=total_polymorphic[which(type=="non-synonymous")],
            n_polymorphic.syn=total_polymorphic[which(type=="synonymous")]) 

hwe <- hwe %>%
  mutate(pop=factor(pop, levels=c("STER","KLDR","CEBO")),
         pop_site=factor(pop_site,levels=c("STER ns", "STER syn",
                                   "KLDR ns", "KLDR syn",
                                   "CEBO ns", "CEBO syn"))) %>%
  filter(immune_DEG!="random")


k.test<- hwe %>% 
  select(pop_site,immune_DEG, type2, heterozygosity) %>%
  group_by(immune_DEG) %>%
  wilcox_test(heterozygosity ~ pop_site) %>% 
  adjust_pvalue(method="holm") %>%
  add_significance() %>%
  filter(word(group1,1,1)==word(group2,1,1))



hetfig_up<- ggplot(hwe,aes(x=pop_site, y=heterozygosity, color=pop_site,group=pop_site)) +
  geom_boxplot(lwd=baselwd, show.legend = FALSE,position=position_dodge(width=0.9),outlier.colour = NA) + 
  geom_point(position=position_dodge(width=0.9), alpha=0.3) + 
  facet_wrap( ~ immune_DEG, scales="free",ncol=3) + 
  labs(y="Per-Site Heterozygosity",x="") + 
  scale_x_discrete(labels=rep(c("N","S"),3))  + peri_figure  +  
  scale_color_manual(values=viridispairs[c(1,3,4,6,10,12)],name="") +
  stat_summary(fun = "mean",geom = "point", pch=4, stroke=0.5, size=3, position=position_dodge(width=0.9),show.legend = FALSE) +
  theme(legend.position="top", legend.direction = "horizontal") +
  stat_pvalue_manual(k.test, label = "p.adj.signif", tip.length = 0.005,hide.ns=TRUE, label.size=geom.text.size,bracket.size = baselwd, y.position=c(0.082,0.101,0.1))
  
  
  #stat_pvalue_manual(k.test, label = "p.adj.signif", tip.length = 0.005,hide.ns=TRUE, label.size=geom.text.size, y.position=c(0.082,0.013,0.04,0.03), bracket.size = baselwd) + 

hetfig_up
```

```{r}
#| label: homozygosity-ns
#| message: false
#| results: hide
#| echo: false
#| eval: false
homfig_up<- ggplot(hwe,aes(x=pop_site, y=minor_homoz, color=pop_site,group=pop_site)) +
  geom_boxplot(lwd=baselwd, show.legend = FALSE,position=position_dodge(width=0.9),outlier.colour = NA) + 
  geom_point(position=position_dodge(width=0.9), alpha=0.3) + 
  facet_wrap( ~ immune_DEG, scales="free",ncol=3) + 
  labs(y="Homozygosity Minor Allele",x="") + 
  scale_x_discrete(labels=rep(c("N","S"),3))  + peri_figure  +  
  scale_color_manual(values=viridispairs[c(1,3,4,6,10,12)],name="") +
  stat_summary(fun = "mean",geom = "point", pch=4, stroke=0.5, size=3, position=position_dodge(width=0.9),show.legend = FALSE) 
  #stat_pvalue_manual(k.test, label = "p.adj.signif", tip.length = 0.005,hide.ns=TRUE, labe.size=geom.text.size, y.position=c(0.082,0.013,0.04,0.03)) + 
  #theme(legend.position="top", legend.direction = "horizontal")
homfig_up
```

```{r tbl-genes}
#| label: tbl-genes
#| tbl-cap-location: top
#| tbl-cap: "Number of genes investigated in each category in this report."
#| echo: false
#| tbl-colwidths: [75,25]
#| fig-width: 8


het_sum<- hwe %>% 
  group_by(immune_DEG) %>% 
  summarise(n_genes=length(unique(geneID)))

knitr::kable(het_sum)

```

```{r nsvs-pi-configure noeval}
#| label: nsvs-pi-configure
#| echo: false
#| results: hide
#| message: false
#| eval: false
pi.dm<- read_pestPG(paste0(datadir,"/FGD/degeneracy/"),pop3)

pi.dm<- pi.dm %>% 
  mutate(immune_DEG=ifelse(immune_DEG=="" | immune_DEG=="immune","random",as.character(immune_DEG))) %>%
  mutate(immune_DEG=factor(immune_DEG,levels=c("DEG","lower DEG","upper DEG","immune DEG","lower immune DEG","upper immune DEG","random"))) %>%
  mutate(pop_site=factor(pop_site, levels=c("STER ns","STER syn","KLDR ns","KLDR syn","CEBO ns","CEBO syn"))) %>%
  filter(immune_DEG!="random")

write.table(pi.dm, paste0(datadir,"/FGD/degeneracy/pi_per_gene.csv"), row.names=FALSE, quote=FALSE, sep="\t")
```

```{r nsvs-pi}
#| label: pi-ns
#| echo: false
#| results: hide
#| message: false

pi.dm<- read.table(paste0(datadir,"/FGD/degeneracy/pi_per_gene.csv"), header=TRUE, sep="\t")
pi.dm$immune_DEG<- ifelse(grepl("upper", pi.dm$immune_DEG),
                        gsub("upper ","", pi.dm$immune_DEG),
                        as.character(pi.dm$immune_DEG))
pi.dm$immune_DEG<- ifelse(grepl("lower", pi.dm$immune_DEG),
                        gsub("lower ","", pi.dm$immune_DEG),
                        as.character(pi.dm$immune_DEG))

pi.dm <- pi.dm %>%
    mutate(pop=factor(pop, levels=c("STER","KLDR","CEBO")),
         pop_site=factor(pop_site,levels=c("STER ns", "STER syn",
                                   "KLDR ns", "KLDR syn",
                                   "CEBO ns", "CEBO syn")))
k.test<- pi.dm %>% 
  select(pop_site,immune_DEG, type2, pi) %>%
  group_by(immune_DEG) %>%
  wilcox_test(pi ~ pop_site) %>% 
  adjust_pvalue(method="holm") %>%
  add_significance() %>%
  add_y_position() %>%
  filter(word(group1,1,1)==word(group2,1,1))


pifig_up<- ggplot(pi.dm ,aes(x=pop_site, y=pi, color=pop_site)) +
  geom_boxplot(lwd=baselwd, show.legend = FALSE,position=position_dodge(width=0.9),outlier.colour = NA) +  
  geom_point(position=position_dodge(width=0.9), alpha=0.3) + 
  facet_wrap( ~ immune_DEG,scales="free",ncol=3) +
  labs(y=expression(paste("Nucleotide Diversity (",pi,")", sep="")),x="") +
  scale_x_discrete(labels=rep(c("N","S"),3))  + peri_figure  +
  theme(ggh4x.axis.nestline.x=element_blank(),legend.position = "top", legend.title=element_blank()) +
  scale_color_manual(values=viridispairs[c(1,3,4,6,10,12)],name="") + 
  stat_summary(fun = "mean",geom = "point", pch=4, stroke=0.5, size=3, position=position_dodge(width=0.9),show.legend = FALSE) +
  stat_pvalue_manual(k.test, label = "p.adj.signif", tip.length = 0.005,hide.ns=TRUE, label.size=geom.text.size, bracket.size = baselwd,y.position=rep(max(pi.dm$pi),6)) + theme(legend.position="top", legend.direction = "horizontal")
pifig_up
```

```{r nsvs-tajd}
#| label: tajimaD-ns
#| echo: false
k.test<- pi.dm %>% 
  select(pop_site,immune_DEG, type2, Tajima) %>%
  group_by(immune_DEG) %>%
  wilcox_test(Tajima ~ pop_site) %>% 
  adjust_pvalue(method="holm") %>%
  add_significance() %>%
  add_y_position() %>%
  filter(word(group1,1,1)==word(group2,1,1))


tajfig_up<- ggplot(pi.dm ,aes(x=pop_site, y=Tajima, color=pop_site)) +
  geom_boxplot(lwd=baselwd, show.legend = FALSE,position=position_dodge(width=0.9),outlier.colour = NA) +  
  geom_point(position=position_dodge(width=0.9), alpha=0.3) + 
  facet_wrap( ~ immune_DEG,scales="free",ncol=3) +
  labs(y="Tajima's D",x="") +
  scale_x_discrete(labels=rep(c("N","S"),3))  + peri_figure  +
  scale_color_manual(values=viridispairs[c(1,3,4,6,10,12)],name="") + 
  stat_summary(fun = "mean",geom = "point", pch=4, stroke=0.5, size=3, position=position_dodge(width=0.9),show.legend = FALSE) +
  stat_pvalue_manual(k.test, label = "p.adj.signif", tip.length = 0.005,hide.ns=TRUE, label.size=geom.text.size, bracket.size = baselwd) + 
  theme(legend.position="top", legend.direction = "horizontal", legend.title=element_blank())
tajfig_up
```

```{r nsvs-fud-configure}
#| label: nsvs-fud-configure
#| echo: false
#| eval: false
k.test<- pi.dm %>% 
  select(pop_site,immune_DEG, type2, fud) %>%
  group_by(immune_DEG) %>%
  wilcox_test(fud~ pop_site) %>% 
  adjust_pvalue(method="holm") %>%
  add_significance() %>%
  add_y_position() %>%
  filter(word(group1,1,1)==word(group2,1,1))


fudfig_up<- ggplot(pi.dm ,aes(x=pop_site, y=fud, color=pop_site)) +
  geom_boxplot(lwd=baselwd, show.legend = FALSE,position=position_dodge(width=0.9),outlier.colour = NA) +  
  geom_point(position=position_dodge(width=0.9), alpha=0.3) + 
  facet_wrap( ~ immune_DEG,scales="free",ncol=3) +
  labs(y="Fu and Li's D",x="") +
  scale_x_discrete(labels=rep(c("N","S"),3))  + peri_figure  +
  scale_color_manual(values=viridispairs[c(1,3,4,6,10,12)],name="") + 
  stat_summary(fun = "mean",geom = "point", pch=4, stroke=0.5, size=3, position=position_dodge(width=0.9),show.legend = FALSE) +
  stat_pvalue_manual(k.test, label = "p.adj.signif", tip.length = 0.005,hide.ns=TRUE, label.size=geom.text.size, bracket.size = baselwd) + 
  theme(legend.position="top", legend.direction = "horizontal", legend.title=element_blank())
```

```{r nsvs-fuf-configure}
#| label: nsvs-fuf-configure
#| echo: false
#| message: false
#| eval: false
k.test<- pi.dm %>% 
  select(pop_site,immune_DEG, type2, fuf) %>%
  group_by(immune_DEG) %>%
  wilcox_test(fuf~ pop_site) %>% 
  adjust_pvalue(method="holm") %>%
  add_significance() %>%
  add_y_position() %>%
  filter(word(group1,1,1)==word(group2,1,1))


fuffig_up<- ggplot(pi.dm ,aes(x=pop_site, y=fuf, color=pop_site)) +
  geom_boxplot(lwd=baselwd, show.legend = FALSE,position=position_dodge(width=0.9),outlier.colour = NA) +  
  geom_point(position=position_dodge(width=0.9), alpha=0.3) + 
  facet_wrap( ~ immune_DEG,scales="free",ncol=3) +
  labs(y="Fu and Li's F",x="") +
  scale_x_discrete(labels=rep(c("N","S"),3))  + peri_figure  +
  scale_color_manual(values=viridispairs[c(1,3,4,6,10,12)],name="") + 
  stat_summary(fun = "mean",geom = "point", pch=4, stroke=0.5, size=3, position=position_dodge(width=0.9),show.legend = FALSE) +
  stat_pvalue_manual(k.test, label = "p.adj.signif", tip.length = 0.005,hide.ns=TRUE, label.size=geom.text.size, bracket.size = baselwd) + 
  theme(legend.position="top", legend.direction = "horizontal", legend.title=element_blank())
```

```{r fig-nsvs, eval=FALSE}
#| label: fig-nsvs
#| echo: false
#| eval: false
#| fig.width: 12
#| fig.height: 21
#| fig.cap: "Comparison of synonymous and non-synonymous summary statistics with annotation of significant differences after p-value correction"

ggarrange(hetfig_up, pifig_up, tajfig_up, common.legend=TRUE, ncol=1, nrow=3,labels=LETTERS[1:3])
ggsave(filename=paste0(outdir,"/fig_nsvs.png"),plot = last_plot(),height=200, width=140, units="mm", device="png", bg="white")
```

### Gene deserts

```{r gd-het-config}
#| label: gd-het-config
#| echo: false
#| message: false
#| results: hide

#read in the 1000 overlapping GD sites. 
gd.hwe<- read.csv(paste0(datadir,"/geneDesert/angsd_variants/all_KLDR_CEBO_STER_gd_hwe.csv"))
gd.hwe<- gd.hwe %>% 
  dplyr::rename(geneID=gene)

#%>% filter(pop=="STER" | pop=="CEBO" | pop=="KLDR")

confidences.het<- gd.hwe %>% 
  group_by(pop) %>% 
  summarise(lower95=quantile(heterozygosity,0.05), 
            upper95=quantile(heterozygosity,0.95),
            log_lower95=quantile(log10(heterozygosity),0.05),
            log_upper95=quantile(log10(heterozygosity),0.95)) %>%
  mutate(pop=factor(pop,pop3))


levene_test(gd.hwe, heterozygosity ~ pop)
#combining data

relevant<- c("geneID","immune_DEG","pop","pop_site","heterozygosity","exp_heterozygosity")

mega.hwe<- rbind(gd.hwe[,relevant],hwe[hwe$type=="non-synonymous",relevant])
mega.hwe<- mega.hwe %>% 
  mutate(gene2=ifelse(immune_DEG=="desert","",geneID),
         type=ifelse(grepl("ns",pop_site),"ns","desert"),
         pop_class=paste(pop,immune_DEG),
         log10het=log10(heterozygosity)) %>%
  mutate(type=ifelse(grepl("ns",pop_site),"ns","desert"),
         pop=factor(pop, levels=pop3)) %>%  
  mutate(pop_site=factor(pop_site, levels=c("STER desert",
                                          "STER ns",
                                          "KLDR desert",
                                          "KLDR ns",
                                          "CEBO desert",
                                          "CEBO ns")),
         pop_class=factor(pop_class, levels=c("STER desert",
                                            "STER DEG",
                                            "STER immune DEG",
                                            "KLDR desert",
                                            "KLDR DEG",
                                            "KLDR immune DEG",
                                            "CEBO desert",
                                            "CEBO DEG",
                                            "CEBO immune DEG")),
        immune_DEG=factor(immune_DEG, levels=c("desert",
                                            "DEG",
                                            "immune DEG")))

  
  
  
#  mutate(pop_site=factor(pop_site, levels=c("STER desert",
#                                            "STER ns",
#                                            "KLDR desert",
#                                            "KLDR ns",
#                                            "CEBO desert",
#                                            "CEBO ns")),
#         pop_class=factor(pop_class, levels=c("STER desert",
#                                              "STER random",
#                                              "STER DEG",
#                                              "STER immune DEG",
#                                              "STER lower DEG",
#                                              "STER lower immune DEG",
#                                              "STER upper DEG", 
#                                              "STER upper immune DEG",
#                                                         "KLDR desert",
#                                              "KLDR random",
#                                              "KLDR DEG",
#                                              "KLDR immune DEG",
#                                              "KLDR lower DEG", 
#                                              "KLDR lower immune DEG",  
#                                              "KLDR upper DEG", 
#                                              "KLDR upper immune DEG",
#                                                         "CEBO desert",
#                                              "CEBO random",
#                                              "CEBO DEG",
#                                              "CEBO immune DEG",
#                                              "CEBO lower DEG", 
#                                              "CEBO lower immune DEG", 
#                                              "CEBO upper DEG", 
#                                              "CEBO upper immune DEG")),
#         immune_DEG=factor(immune_DEG, levels=c("desert",
#                                                "random",
#                                                "DEG",
#                                                "immune DEG",
#                                                "lower DEG",
#                                                "lower immune DEG",
#                                                "upper DEG",
#                                                "upper immune DEG")))
#  
mega.hwe<- droplevels(mega.hwe)
```

```{r gd-pi-config}
#| label: gd-pi-config
#| echo: false
#| message: false
#| results: hide

gd.out<- list()

for(p in pop3){
  print(p)
  gd<- read.table(paste0(datadir,"/geneDesert/angsd_variants/",p,".genedesert.thetasWindow.gz.pestPG"))
  
  colnames(gd)<- picols[-15]
  gd$pop<- p
  #for some reason this wasn't working right in the piping
  gd$winStart<- word(gd$`indexStart_indexStop-firstPosWithData_lastPosWithData-WinStart_WinStop`,3,3, sep="\\(")
  gd$winStop<- word(gd$winStart,2,2, sep=",")
  gd$winStart<- as.numeric(word(gd$winStart,1,1, sep=","))-1
  gd$winStop<- gsub("\\)","",gd$winStop)
  gd<- gd %>% 
  mutate(geneID= paste0(Chr, ":",winStart,"-",winStop),
         pi=tP/nSites,
         watterson_theta=tW/nSites,
         pop_site=paste(pop, "desert"),
         immune_DEG="desert") %>%
  mutate(log10pi=log10(pi)) %>%
  mutate(log10pi=ifelse(pi==0,NA,log10pi))
  #keep only the sites previously included.
  gd<- gd[gd$geneID %in% gd.hwe$geneID,]
  gd.out[[p]]<- gd
}

gd.pi<- do.call("rbind",gd.out)

levene_test(gd.pi[which(gd.pi$pop!="CEBO"),], pi ~ pop)
levene_test(gd.pi[which(gd.pi$pop!="CEBO"),], Tajima ~ pop)
confidences.pi<- gd.pi %>% 
  group_by(pop) %>% 
  summarise(pi_lower95=quantile(pi,0.05, na.rm=TRUE),
            pi_upper95=quantile(pi,0.95, na.rm=TRUE),
            log10pi_upper95=quantile(log10pi,0.95, na.rm=TRUE),
            log10pi_lower95=quantile(log10pi,0.05,na.rm=TRUE),
            D_lower95=quantile(Tajima,0.05, na.rm=TRUE),
            D_upper95=quantile(Tajima,0.95, na.rm=TRUE)) %>%
  mutate(pop=factor(pop,pop3))

#combining data

relevant<- c("geneID","pi","log10pi","immune_DEG","pop","pop_site","Tajima")


mega.pi<- rbind(gd.pi[,relevant],pi.dm[pi.dm$type=="non-synonymous",relevant])
mega.pi<- mega.pi %>% 
  mutate(gene2=ifelse(immune_DEG=="desert","",geneID),
         type=ifelse(grepl("ns",pop_site),"ns","desert"),
         pop_class=paste(pop,immune_DEG)) %>%
  mutate(type=ifelse(grepl("ns",pop_site),"ns","desert"),
         pop=factor(pop, levels=pop3)) %>%  
  mutate(pop_site=factor(pop_site, levels=c("STER desert",
                                            "STER ns",
                                            "KLDR desert",
                                            "KLDR ns",
                                            "CEBO desert",
                                            "CEBO ns")),
         pop_class=factor(pop_class, levels=c("STER desert",
                                              "STER DEG",
                                              "STER immune DEG",
                                              "KLDR desert",
                                              "KLDR DEG",
                                              "KLDR immune DEG",
                                              "CEBO desert",
                                              "CEBO DEG",
                                              "CEBO immune DEG")),
         immune_DEG=factor(immune_DEG, levels=c("desert",
                                                "DEG",
                                                "immune DEG")))
  
mega.pi<- droplevels(mega.pi)


```

```{r fig-gd-FuF}
#| label: fig-gd-FuF
#| echo: false
#| message: false
#| eval: false
#| fig.cap: "Fu and Li's F in 1000 gene desert regions of 25k windows and non-synonymous sites in our focal genes"
k.test<- mega.pi %>% 
  select(pop_class,pop, immune_DEG, type, fuf) %>%
  group_by(pop) %>%
  wilcox_test(fuf ~ pop_class) %>% 
  add_y_position() %>%
  filter(word(group1,1,1)==word(group2,1,1)) %>%
  filter(grepl("desert",group1)) %>%
  adjust_pvalue(method="holm") %>%
  add_significance() 

#singletons vs avg number variants
megafuf.all<- ggplot(mega.pi,aes(x=pop_class, y=fuf, color=pop, group=pop_class)) +
  geom_boxplot(lwd=baselwd, show.legend = FALSE,position=position_dodge(width=0.9),outlier.colour = NA) + 
  geom_point(position=position_dodge(width=0.9), alpha=0.3) +
  facet_grid(~pop, scales="free") +
  labs(x="",y="Fu and Li's F") + peri_figure +
  theme(legend.title = element_blank(), 
        axis.text.x = element_text(angle=90, hjust=1, vjust=0.5),
        legend.position="top", 
        legend.direction = "horizontal")  + 
  scale_y_continuous(breaks=c(-3,-2,-1,0,1,2,3,4,5)) +
  scale_color_manual(values=viridispairs[c(3,6,12)],name="") + 
  stat_summary(fun = "mean",geom = "point", pch=4, stroke=0.5, size=3, position=position_dodge(width=0.9),show.legend = FALSE) + 
  geom_hline(dat=confidences.pi, aes(yintercept=fuF_upper95, colour=pop,), show.legend = FALSE, linetype="dashed",lwd=baselwd) + 
  geom_hline(dat=confidences.pi, aes(yintercept=fuF_lower95, colour=pop), show.legend = FALSE, linetype="dashed",lwd=baselwd) + 
  stat_pvalue_manual(k.test, label = "p.adj.signif", tip.length = 0.005,hide.ns=TRUE, label.size=geom.text.size, bracket.size = baselwd, y.position=k.test$y.position[which(k.test$p.adj.signif!="ns")]-0.5)

megafuf.all
```

```{r fig-gd-FuD}
#| label: fig-gd-FuD
#| echo: false
#| message: false
#| eval: false
#| fig.cap: "Fu and Li's D in 1000 gene desert regions of 25k windows and non-synonymous sites in our focal genes"

k.test<- mega.pi %>% 
  select(pop_class,pop, immune_DEG, type, fud) %>%
  group_by(pop) %>%
  wilcox_test(fud ~ pop_class) %>% 
  add_y_position() %>%
  filter(word(group1,1,1)==word(group2,1,1)) %>%
  filter(grepl("desert",group1)) %>%
  adjust_pvalue(method="holm") %>%
  add_significance() 

#singletons vs total number variants
megafud.all<- ggplot(mega.pi,aes(x=pop_class, y=fud, color=pop, group=pop_class)) +
  geom_boxplot(lwd=baselwd, show.legend = FALSE,position=position_dodge(width=0.9),outlier.colour = NA) + 
  geom_point(position=position_dodge(width=0.9), alpha=0.3) +
  facet_grid(~pop, scales="free") +
  labs(x="",y="Fu and Li's D") + peri_figure +
  theme(legend.title = element_blank(), 
        axis.text.x = element_text(angle=90, hjust=1, vjust=0.5),
        legend.position="top", 
        legend.direction = "horizontal")  + 
  scale_y_continuous(breaks=c(-3,-2,-1,0,1,2,3,4,5)) +
  scale_color_manual(values=viridispairs[c(3,6,12)],name="") + 
  stat_summary(fun = "mean",geom = "point", pch=4, stroke=0.5, size=3, position=position_dodge(width=0.9),show.legend = FALSE) + 
  geom_hline(dat=confidences.pi, aes(yintercept=fuD_upper95, colour=pop,), show.legend = FALSE, linetype="dashed",lwd=baselwd) + 
  geom_hline(dat=confidences.pi, aes(yintercept=fuD_lower95, colour=pop), show.legend = FALSE, linetype="dashed",lwd=baselwd) + 
  stat_pvalue_manual(k.test, label = "p.adj.signif", tip.length = 0.005,hide.ns=TRUE, label.size=geom.text.size, bracket.size = baselwd, y.position=k.test$y.position[which(k.test$p.adj.signif!="ns")]-0.2)

megafud.all
```

```{r summarise pi}
#| label: summarise
#| echo: false
pi.summary<- pi.dm %>% 
  group_by(pop,geneID,immune_DEG) %>% 
  reframe(pi.ns=round(pi[which(type=="non-synonymous")],5),
            pi.syn=round(pi[which(type=="synonymous")],5),
            watterson.ns=round(watterson_theta[which(type=="non-synonymous")],5),
            watterson.syn=round(watterson_theta[which(type=="synonymous")],5),
            tajima.ns=round(Tajima[which(type=="non-synonymous")],2), 
            tajima.syn=round(Tajima[which(type=="synonymous")],2)) 
#%>%
 # mutate(pnps.pi=round(pi.ns/pi.syn,3))
           #watterson.ns=round(watterson_theta[which(type=="non-synonymous")],5),
           #watterson.syn=round(watterson_theta[which(type=="synonymous")],5),
           #fuF.ns=round(fuf[which(type=="non-synonymous")],2),
           #fuF.syn=round(fuf[which(type=="synonymous")],2),
           #fuD.ns=round(fud[which(type=="non-synonymous")],2),
           #fuD.syn=round(fud[which(type=="synonymous")],2))

scate.summary<- merge(het.summary, pi.summary, by=c("pop","geneID","immune_DEG"), all.y=TRUE)
#scate.summary$pNpS<- round(scate.summary$n_polymorphic.ns/scate.summary$n_polymorphic.syn,3)
#p="STER"
out<- list()
for(p in pop3){
  sub<- scate.summary[which(scate.summary$pop==p),]
  ##heterozygosity
  conf_sub<- confidences.het[which(confidences.het$pop==p),]
  sub$het.ns.outlier<- ifelse(sub$heterozygosity.ns<conf_sub$lower95,-1,ifelse( sub$heterozygosity.ns>conf_sub$upper95,1,0))
  sub$het.syn.outlier<- ifelse(sub$heterozygosity.syn<conf_sub$lower95 | sub$heterozygosity.syn>conf_sub$upper95,1,0)
  
  conf_sub<- confidences.pi[which(confidences.pi$pop==p),]
  sub$pi.ns.outlier<- ifelse(sub$pi.ns<conf_sub$pi_lower95,-1,ifelse( sub$pi.ns>conf_sub$pi_upper95,1,0))
  sub$pi.syn.outlier<- ifelse(sub$pi.syn<conf_sub$pi_lower95 | sub$pi.syn>conf_sub$pi_upper95,1,0)
  
  sub$tajima.ns.outlier<- ifelse(sub$tajima.ns<conf_sub$D_lower95,-1,ifelse(sub$tajima.ns>conf_sub$D_upper95,1,0)) 
### noting if above or below the min/max cutoff.   

  sub$tajima.syn.outlier<- ifelse(sub$tajima.syn<conf_sub$D_lower95 | sub$tajima.syn>conf_sub$D_upper95,1,0)
  
  #sub$fuF.ns.outlier<- ifelse(sub$fuF.ns<conf_sub$fuF_lower95 | sub$fuF.ns>conf_sub$fuF_upper95,1,0)
  #sub$fuF.syn.outlier<- ifelse(sub$fuF.syn<conf_sub$fuF_lower95 | sub$fuF.syn>conf_sub$fuF_upper95,1,0)
  
  #sub$fuD.ns.outlier<- ifelse(sub$fuD.ns<conf_sub$fuD_lower95 | sub$fuD.ns>conf_sub$fuD_upper95,1,0)
  #sub$fuD.syn.outlier<- ifelse(sub$fuD.syn<conf_sub$fuD_lower95 | sub$fuD.syn>conf_sub$fuD_upper95,1,0)

  out[[p]]<- sub
}

scate.summary.out<- do.call("rbind",out)

outlier.summary<- scate.summary.out %>% 
  group_by(pop,immune_DEG) %>%
  summarise(n_genes=length(unique(geneID)),
            n_out_het.ns=sum(abs(het.ns.outlier)),
            n_out_het.syn=sum(abs(het.syn.outlier)),
            n_out_pi.ns=sum(abs(pi.ns.outlier)),
            n_out_pi.syn=sum(pi.syn.outlier),
            n_out_tajima.ns=sum(abs(tajima.ns.outlier)), 
            n_out_tajima.syn=sum(tajima.syn.outlier),
            #n_out_fuF.ns=sum(fuF.ns.outlier,na.rm=TRUE), 
            #n_out_fuF.syn=sum(fuF.syn.outlier,na.rm=TRUE),
            #n_out_fuD.ns=sum(fuD.ns.outlier, na.rm=TRUE),
            #n_out_fuD.syn=sum(fuD.syn.outlier, na.rm=TRUE)
            )

scate.summary.out2<- scate.summary.out %>%
  filter(immune_DEG!="random")
  

gene_annot<- read.xlsx(paste0(datadir,"/GO_annot/Scate_DEGs_annot.xlsx"),sheet=1)
gene_annot<- gene_annot %>% 
  filter(is.na(ortholog_exclude))
gene_annot$Scate<- gsub("-RA","", gene_annot$Scate)
gene_annot<- gene_annot[,c("Scate", "geneName_parsed","geneID")]
colnames(gene_annot)[colnames(gene_annot)=="geneID"]<- "Cviridis_geneID"
colnames(gene_annot)[colnames(gene_annot)=="Scate"]<- "geneID"
scate.summary.out2<- merge(scate.summary.out2, gene_annot, by="geneID", all.x=TRUE)
#scate.summary.out2$gene_pop<- paste0(scate.summary.out2$pop, scate.summary.out2$geneID)
#note: the above creates a dataframe with more rows because there are sometimes multiple Cvi orthologs for a single gene. Using the file below as a guide, I am manually inspecting these annotations and removing the one that agrees least with the Scate gene annotation (or picking one at random if they are the same). I then exclude those sites on re-exporting the annotations above. 
#scate.summary.out.dups<- scate.summary.out2[duplicated(scate.summary.out2$gene_pop),]

scate.summary.out2<- scate.summary.out2 %>%
  mutate(gene_pop=paste0(geneID,pop)) %>%
  filter(!duplicated(gene_pop))
write.csv(scate.summary.out2, file = paste0(outdir,"/report/result_pi_het.csv"), row.names=FALSE)
write.csv(outlier.summary, file = paste0(outdir,"/report/result_outlier_summary.csv"), row.names=FALSE)

```

```{r}
outlier.summary2<- scate.summary.out %>% 
  group_by(pop,immune_DEG) %>%
  summarise(n_genes=length(unique(geneID)),
            tajima.ns_pos=sum(tajima.ns.outlier[tajima.ns.outlier==1]), 
            tajima.ns_neg=abs(sum(tajima.ns.outlier[tajima.ns.outlier==-1])),
            pi.ns_pos=sum(pi.ns.outlier[pi.ns.outlier==1]), 
            pi.ns_neg=abs(sum(pi.ns.outlier[pi.ns.outlier==-1])),
            het.ns_pos=sum(het.ns.outlier[het.ns.outlier==1]),
            het.ns_neg=abs(sum(het.ns.outlier[het.ns.outlier==-1])))

write.csv(outlier.summary2, file = paste0(outdir,"/report/result_outlier_summary_updown.csv"), row.names=FALSE)

```

```{r fig-summary}
#| label: fig-summary
#| fig.cap: "Number of interest genes that lie outside 95% of the gene desert statistics. Panels arranged by statistic (y) and gene category (x), where 'all DEGs' is the sum of the non-random categories on the left."
#| echo: false
#| message: false
#| fig.height: 9
#| fig.width: 11
#outlier.summary.long<- melt(outlier.summary, id.vars=c("pop","immune_DEG","n_genes"),value.name = "number")
outlier.summary.long<- melt(outlier.summary2, id.vars=c("pop","immune_DEG","n_genes"),value.name = "number")

#outlier.summary.long<- outlier.summary.long %>%
#  mutate(stat=word(variable,3,3, sep="_")) %>%
#  mutate(stat=word(stat,1,1,sep="\\.")) %>%
#  mutate(type=word(variable,2,2,sep="\\."),
#         pc=(number/n_genes)*100) %>%
#  select(!variable) 

outlier.summary.long<- outlier.summary.long %>%
  mutate(stat=word(variable,1,1, sep="\\.")) %>%
  mutate(type=word(variable,2,2,sep="\\."),
         pc=(number/n_genes)*100) %>%
  select(!variable) %>%
  filter(number>0)

out.all<- outlier.summary.long %>% 
  group_by(pop,stat,type) %>%
  reframe(immune_DEG="all DEG",
          n_genes=sum(n_genes),
          number=sum(number)) %>%
  mutate(pc=(number/n_genes)*100)

outlier.summary.long<- rbind(outlier.summary.long, out.all)
#outlier.summary.long<- outlier.summary.long %>%
#  filter(stat=="pi" | stat=="tajima") %>%
#  mutate(stat=factor(stat, levels=c("pi","tajima"))) %>% 
#  mutate(immune_DEG=factor(immune_DEG, levels=c("all DEG", "random", "DEG","immune DEG","lower #DEG","lower immune DEG", "upper DEG", "upper immune DEG"))) %>%
#  filter(type=="ns") %>%
#  mutate(pop=factor(pop, levels=c("STER","KLDR","CEBO")))
         
outlier.summary.long<- outlier.summary.long %>%
  mutate(immune_DEG=factor(immune_DEG, levels=c("all DEG","DEG","immune DEG","lower DEG","lower immune DEG", "upper DEG", "upper immune DEG"))) %>%
  mutate(pop=factor(pop, levels=c("STER","KLDR","CEBO"))) %>%
  mutate(pc=ifelse(grepl("neg",type),pc*-1,pc)) %>%
  mutate(stat=factor(stat, levels=c("het","pi","tajima"))) %>%
  mutate(pop_class=paste(pop, immune_DEG)) %>%
  mutate(fig=paste0(number,"\n(",round(abs(pc),0),"%)"))


#stat_names<- c("fuD"="Fu and Li's D",
#                "fuF"="Fu and Li's F",
#                "het"="heterozygosity",
#                "pi"="\u03c0",
#                "tajima"="Tajima's D")

stat_names<- c("het"="Heterozygosity",
               "pi"="\u03c0",
               "tajima"="Tajima's D")
a<- ggplot(outlier.summary.long[which(outlier.summary.long$stat!="het"),], aes(x=immune_DEG,y=pc, fill=pop)) + 
  geom_bar(stat="identity", position=position_dodge(width=0.9)) + 
  facet_grid( ~ stat, labeller=labeller(stat=stat_names)) + 
  scale_fill_manual(values=viridispairs[c(3,6,12)],name="") +
  peri_figure + scale_y_continuous(labels=function(x) paste0(x, "%")) + 
  labs(y="Percent category genes outlying 95% of gene desert values", x="") + 
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5)) + 
  geom_hline(yintercept = 0, lwd=baselwd, linetype="dashed")


tajd_1<- list("STER"=scate.summary.out2$geneID[which(abs(scate.summary.out2$tajima.ns.outlier)==1 & scate.summary.out2$pop=="STER")],
                "KLDR"=scate.summary.out2$geneID[which(abs(scate.summary.out2$tajima.ns.outlier)==1 & scate.summary.out2$pop=="KLDR")],
              "CEBO"=scate.summary.out2$geneID[which(abs(scate.summary.out2$tajima.ns.outlier)==1 & scate.summary.out2$pop=="CEBO")])

tajd_overall_fig<- ggvenn(tajd_1, 
           set_name_size = 0, 
           text_size=3, 
           stroke_size=baselwd, 
           fill_color = viridispairs[c(3,6,12)], 
           show_percentage = FALSE) + 
  coord_cartesian(clip="off") + 
  coord_equal() +
  labs(title="Tajima's D outlier") 


tajd_neg_outlier<- list("STER"=scate.summary.out2$geneID[which(scate.summary.out2$tajima.ns.outlier==-1 & scate.summary.out2$pop=="STER")],
                "KLDR"=scate.summary.out2$geneID[which(scate.summary.out2$tajima.ns.outlier==-1 & scate.summary.out2$pop=="KLDR")],
              "CEBO"=scate.summary.out2$geneID[which(scate.summary.out2$tajima.ns.outlier==-1 & scate.summary.out2$pop=="CEBO")])

tajd_neg_fig<- ggvenn(tajd_neg_outlier, 
           set_name_size = 0, 
           text_size=3, 
           stroke_size=baselwd, 
           fill_color = viridispairs[c(3,6,12)], show_percentage = FALSE) + 
  coord_cartesian(clip="off") + 
  coord_equal() +
  labs(title="Tajima's D < 0 outliers") + 
  theme(plot.title = element_text(hjust = 0.5, size=basesize))

tajd_pos_outlier<- list("STER"=scate.summary.out2$geneID[which(scate.summary.out2$tajima.ns.outlier==1 & scate.summary.out2$pop=="STER")],
                "KLDR"=scate.summary.out2$geneID[which(scate.summary.out2$tajima.ns.outlier==1 & scate.summary.out2$pop=="KLDR")],
              "CEBO"=scate.summary.out2$geneID[which(scate.summary.out2$tajima.ns.outlier==1 & scate.summary.out2$pop=="CEBO")])

tajd_pos_fig<- ggvenn(tajd_pos_outlier, 
           set_name_size = 0, 
           text_size=3, 
           stroke_size=baselwd, 
           fill_color = viridispairs[c(3,6,12)], show_percentage = FALSE) + 
  coord_cartesian(clip="off") + 
  coord_equal() +
  labs(title="Tajima's D > 0 outliers") + 
  theme(plot.title = element_text(hjust = 0.5, size=basesize))

pi_venn_upper<- list("STER"=scate.summary.out2$geneID[which(scate.summary.out2$pi.ns.outlier==1 & scate.summary.out2$pop=="STER")],
                "KLDR"=scate.summary.out2$geneID[which(scate.summary.out2$pi.ns.outlier==1 & scate.summary.out2$pop=="KLDR")],
              "CEBO"=scate.summary.out2$geneID[which(scate.summary.out2$pi.ns.outlier==1 & scate.summary.out2$pop=="CEBO")])


pi_venn_upper_fig<- ggvenn(pi_venn_upper, 
                           set_name_size = 0, 
                           text_size=3, 
                           stroke_size=0.3, 
                           fill_color = viridispairs[c(3,6,12)], 
                           show_percentage = FALSE) + 
  coord_cartesian(clip="off") + 
  coord_equal() +
  labs(title=expression(paste("high ",pi," outliers"))) + 
  theme(plot.title = element_text(hjust = 0.5, size=basesize))


pi_venn_lower<- list("STER"=scate.summary.out2$geneID[which(scate.summary.out2$pi.ns.outlier==-1 & scate.summary.out2$pop=="STER")],
                "KLDR"=scate.summary.out2$geneID[which(scate.summary.out2$pi.ns.outlier==-1 & scate.summary.out2$pop=="KLDR")],
              "CEBO"=scate.summary.out2$geneID[which(scate.summary.out2$pi.ns.outlier==-1 & scate.summary.out2$pop=="CEBO")])

pi_venn_lower_fig<- ggvenn(pi_venn_lower, 
                           set_name_size = 0, 
                           text_size=3, 
                           stroke_size=0.3, 
                           fill_color = viridispairs[c(3,6,12)], 
                           show_percentage = FALSE) + 
  coord_cartesian(clip="off") + 
  coord_equal() +
  labs(title=expression(paste("low ",pi," outliers"))) + 
  theme(plot.title = element_text(hjust = 0.5, size=basesize))





ggarrange(a,ggarrange(pi_venn_upper_fig,pi_venn_lower_fig,tajd_pos_fig,tajd_neg_fig, ncol=4, labels=c("B","C","D","E"), nrow=1), nrow=2, heights=c(2.5,1),labels=c("A",""))
```

```{r fig-gd-het}
#| label: fig-gd-het
#| fig-cap: "Heterozygosity in 1000 gene desert regions of 25k windows and non-synonymous sites in our focal genes"
k.test<- mega.hwe %>% 
  select(pop_class,pop, immune_DEG, type, heterozygosity) %>%
  group_by(pop) %>%
  wilcox_test(heterozygosity ~ pop_class) %>% 
  adjust_pvalue(method="holm") %>%
  add_significance() %>%
  add_y_position() %>%
  filter(word(group1,1,1)==word(group2,1,1)) %>%
  filter(grepl("desert",group1))

megahet.all<- ggplot(mega.hwe,aes(x=pop_class, y=heterozygosity, color=pop, group=pop_class)) +
  geom_boxplot(lwd=baselwd, show.legend = FALSE,position=position_dodge(width=0.9),outlier.colour = NA) + 
  geom_point(position=position_dodge(width=0.9), alpha=0.3) +
  facet_grid(~pop, scales="free") +
  labs(x="",y="Per-site Heterozygosity")  + peri_figure +
  theme(legend.title = element_blank(), 
        axis.text.x = element_text(angle=90, hjust=1, vjust=0.5),
        legend.position="top", 
        legend.direction = "horizontal")  + 
  scale_color_manual(values=viridispairs[c(3,6,12)],name="") + 
  stat_summary(fun = "mean",geom = "point", pch=4, stroke=0.5, size=3, position=position_dodge(width=0.9),show.legend = FALSE) + 
  geom_hline(dat=confidences.het, aes(yintercept=upper95, colour=pop,), show.legend = FALSE, linetype="dashed",lwd=baselwd) + 
  geom_hline(dat=confidences.het, aes(yintercept=lower95, colour=pop), show.legend = FALSE, linetype="dashed",lwd=baselwd) + 
  stat_pvalue_manual(k.test, label = "p.adj.signif", tip.length = 0.005,hide.ns=TRUE, label.size=geom.text.size, bracket.size = baselwd)

#, y.position=c(0.08,0.09,0.10,0.11,0.12,0.08,0.09,0.10,0.08,0.09,0.10)) 

megahet.all<- ggplot(mega.hwe,aes(x=pop_class, 
                        y=heterozygosity, 
                        color=pop, 
                        group=pop_class)) +
  geom_boxplot(lwd=baselwd, 
               show.legend = FALSE,
               position=position_dodge(width=0.9),
               outlier.colour = NA) + 
  geom_point(position=position_dodge(width=0.9), 
             alpha=0.3) +
  facet_grid(~pop, scales="free") +
  labs(x="",
       y="Per-site Observed Heterozygosity") + 
  peri_figure +
  theme(legend.title = element_blank(), 
        axis.text.x = element_text(angle=90, 
                                   hjust=1,
                                   vjust=0.5),
        legend.position="top", 
        legend.direction = "horizontal")  + 
  scale_color_manual(values=viridispairs[c(3,6,12)],
                     name="") +
  stat_summary(fun = "mean",
               geom = "point", 
               pch=4, 
               stroke=0.5, 
               size=3, 
               position=position_dodge(width=0.9),
               show.legend = FALSE) + 
  geom_hline(dat=confidences.het, 
             aes(yintercept=upper95,
                 colour=pop,), 
             show.legend = FALSE, 
             linetype="dashed",
             lwd=baselwd) + 
  geom_hline(dat=confidences.het, 
             aes(yintercept=lower95, 
                 colour=pop), 
             show.legend = FALSE, 
             linetype="dashed",
             lwd=baselwd) + 
  stat_pvalue_manual(k.test, 
                     label = "p.adj.signif", 
                     tip.length = 0.005,
                     hide.ns=TRUE, 
                     label.size=geom.text.size, 
                     bracket.size = baselwd,
                     y.position=rep(c(0.11,0.12),3)) +
  geom_text(data=outlier.summary.long[which(outlier.summary.long$stat=="het" & 
                                              outlier.summary.long$type=="ns_neg" & 
                                              outlier.summary.long$immune_DEG!="all DEG"),],
            aes(x=pop_class,
                y=-0.01,
                label = fig),
            size=geom.text.size,
            show.legend=FALSE) + 
   geom_text(data=outlier.summary.long[which(outlier.summary.long$stat=="het" & 
                                               outlier.summary.long$type=="ns_pos" & 
                                               outlier.summary.long$immune_DEG!="all DEG"),],
            aes(x=pop_class,
                y=0.132,
                label = fig),
            size=geom.text.size,
            show.legend=FALSE) +
  scale_x_discrete(labels=rep(c("desert","DEG", "immune DEG"),3))


                    # y.position=rep(c(0.1,0.11,0.12),3)
megahet.all

ggsave(filename=paste0(outdir,"/fig_het.png"),plot = megahet.all,height=130, width=130, units="mm", device="png", bg="white")
```

```{r fig-gd-pi}
#| label: fig-gd-pi
#| echo: false
#| message: false
#| fig-cap: "Nucleotide diversity (π) in 1000 gene desert regions of 25k windows and non-synonymous sites in our focal genes"
k.test<- mega.pi %>% 
  select(pop_class,pop, immune_DEG, type, log10pi) %>%
  group_by(pop) %>%
  wilcox_test(log10pi ~ pop_class) %>% 
  add_y_position() %>%
  filter(word(group1,1,1)==word(group2,1,1)) %>%
  filter(grepl("desert",group1)) %>%
  adjust_pvalue(method="holm") %>%
  add_significance() 


megapi.all<- ggplot(mega.pi,
                    aes(x=pop_class, 
                        y=log10pi, 
                        color=pop, 
                        group=pop_class)) +
  geom_boxplot(lwd=baselwd, 
               show.legend = FALSE,
               position=position_dodge(width=0.9),
               outlier.colour = NA) + 
  geom_point(position=position_dodge(width=0.9), 
             alpha=0.3) +
  facet_grid(~pop, scales="free") +
  labs(x="",
       y=expression(paste("Nucleotide Diversity ",
                          log[10],
                          "(",pi,")",
                          sep=""))) + 
  peri_figure +
  theme(legend.title = element_blank(), 
        axis.text.x = element_text(angle=90, 
                                   hjust=1,
                                   vjust=0.5),
        legend.position="none", 
        legend.direction = "horizontal",
        plot.margin = margin(t = 5,  # Top margin
                                          r = 5,  # Right margin
                                          b = 0,  # Bottom margin
                                          l = 5, unit="mm"))  + 
  scale_color_manual(values=viridispairs[c(3,6,12)],
                     name="") +
  stat_summary(fun = "mean",
               geom = "point", 
               pch=4, 
               stroke=0.5, 
               size=3, 
               position=position_dodge(width=0.9),
               show.legend = FALSE) + 
  geom_hline(dat=confidences.pi, 
             aes(yintercept=log10pi_upper95, colour=pop,), 
             show.legend = FALSE, 
             linetype="dashed",
             lwd=baselwd) + 
  geom_hline(dat=confidences.pi, 
             aes(yintercept=log10pi_lower95, colour=pop), 
             show.legend = FALSE, 
             linetype="dashed",
             lwd=baselwd) + 
  scale_y_continuous(n.breaks=8) +
  stat_pvalue_manual(k.test, 
                     label = "p.adj.signif", 
                     tip.length = 0.005,
                     hide.ns=TRUE, 
                     label.size=geom.text.size, 
                     bracket.size = baselwd,
                     y.position=rep(c(-1.25,-0.75),3)) +
  geom_text(data=outlier.summary.long[which(outlier.summary.long$stat=="pi" & 
                                              outlier.summary.long$type=="ns_neg" & 
                                              outlier.summary.long$immune_DEG!="all DEG"),],
            aes(x=pop_class,
                y=-9.5,
                label = fig),
            size=geom.text.size,
            show.legend=FALSE) + 
   geom_text(data=outlier.summary.long[which(outlier.summary.long$stat=="pi" & 
                                               outlier.summary.long$type=="ns_pos" & 
                                               outlier.summary.long$immune_DEG!="all DEG"),],
            aes(x=pop_class,
                y=0,
                label = fig),
            size=geom.text.size,
            show.legend=FALSE) +
  scale_x_discrete(labels=rep(c("desert","DEG", "immune DEG"),3))


    #y.position=rep(c(-1.25,-1,-0.75),3)
megapi.all

ggsave(filename=paste0(outdir,"/fig_pi.png"),plot = megapi.all,height=130, width=130, units="mm", device="png", bg="white")
```

```{r fig-gd-tajD}
#| label: fig-gd-tajD
#| echo: false
#| message: false
#| fig.cap: "Tajima's D in 1000 gene desert regions of 25k windows and non-synonymous sites in our focal genes"
k.test<- mega.pi %>% 
  select(pop_class,pop, immune_DEG, type, Tajima) %>%
  group_by(pop) %>%
  wilcox_test(Tajima ~ pop_class) %>% 
  add_y_position() %>%
  filter(word(group1,1,1)==word(group2,1,1)) %>%
  filter(grepl("desert",group1)) %>%
  adjust_pvalue(method="holm") %>%
  add_significance() 

megataj.all<- ggplot(mega.pi,
                     aes(x=pop_class, 
                         y=Tajima, 
                         color=pop, 
                         group=pop_class)) +
  geom_boxplot(lwd=baselwd, 
               show.legend = FALSE,
               position=position_dodge(width=0.9),
               outlier.colour = NA) + 
  geom_point(position=position_dodge(width=0.9), 
             alpha=0.3) +
  facet_grid(~pop, scales="free") +
  labs(x="",y="Tajima's D") + 
  peri_figure +
  theme(legend.title = element_blank(), 
        axis.text.x = element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5),
        legend.position="none", 
        legend.direction = "horizontal",
        plot.margin = margin(t = 0,  # Top margin
                                          r = 5,  # Right margin
                                          b = 5,  # Bottom margin
                                          l = 5, unit="mm"))  + 
  scale_y_continuous(breaks=c(-3,-2,-1,0,1,2,3,4,5)) +
  scale_color_manual(values=viridispairs[c(3,6,12)],name="") + 
  stat_summary(fun = "mean",
               geom = "point", 
               pch=4, 
               stroke=0.5, 
               size=3, 
               position=position_dodge(width=0.9),
               show.legend = FALSE) + 
  geom_hline(dat=confidences.pi, 
             aes(yintercept=D_upper95, colour=pop,), 
             show.legend = FALSE, 
             linetype="dashed",
             lwd=baselwd) + 
  geom_hline(dat=confidences.pi, 
             aes(yintercept=D_lower95, colour=pop), 
             show.legend = FALSE, 
             linetype="dashed",
             lwd=baselwd) + 
  stat_pvalue_manual(k.test, 
                     label = "p.adj.signif", 
                     tip.length = 0.005,
                     hide.ns=TRUE, 
                     label.size=geom.text.size, 
                     bracket.size = baselwd) + 
  geom_text(data=outlier.summary.long[which(outlier.summary.long$stat=="tajima" & 
                                               outlier.summary.long$type=="ns_neg" & 
                                               outlier.summary.long$immune_DEG!="all DEG"),],
            aes(x=pop_class,
                y=-2.5,
                label = fig),
            size=geom.text.size,
            show.legend=FALSE) + 
   geom_text(data=outlier.summary.long[which(outlier.summary.long$stat=="tajima" & 
                                               outlier.summary.long$type=="ns_pos" & 
                                               outlier.summary.long$immune_DEG!="all DEG"),],
            aes(x=pop_class,
                y=4.5,
                label = fig),
            size=geom.text.size,
            show.legend=FALSE) +
  scale_x_discrete(labels=rep(c("desert","DEG", "immune DEG"),3))


#y.position=c(2.6,2.8,3,
#                                 3.6,3.8,4.0,
#                                 3,3.2,3.4)    
megataj.all

ggsave(filename=paste0(outdir,"/fig_taj.png"),plot = megataj.all,height=130, width=130, units="mm", device="png", bg="white")

                     
                     
                
```

```{r fig-tajD-names}
#| label: fig-tajD-names
#| eval: false
#| fig.cap: "synonymous vs non-synonymous Tajima's D. Extreme values are annotated (GLYATL3, MRM3), as well as outlier immune genes"

d<- ggplot(scate.summary.out2, aes(x=tajima.syn,y=tajima.ns, colour=pop, alpha=n_polymorphic.ns, shape=as.factor(tajima.ns.outlier))) +
  geom_point() + 
   peri_figure +
  labs(x="Synonymous Sites Tajima's D",y="Non-Synonymous Sites Tajima's D ") +
  scale_color_manual(values=viridispairs[c(3,6,12)],name="population") + 
  scale_alpha_continuous(name="No. polymorphic NS") +
  scale_shape_discrete(name="NS Tajima's D outlier") + 
  geom_text_repel(aes(label=interesting), alpha=0.9, show.legend = FALSE, size=geom.text.size, max.overlaps = 24) +
  scale_x_continuous(breaks=seq(-2,4,1)) + 
  scale_y_continuous(breaks=seq(-2,4,1)) + coord_equal()

d
```

```{r degs}
#| echo: false
liverTreat<- read.csv("https://raw.githubusercontent.com/samarth8392/SFDTranscriptomics/main/DataFiles/Liver/genelist/Liver_DGE_TreatOnly_withAnn.csv")
liverTreat$GeneID<- trimws(liverTreat$GeneID) #making sure mismatches don't occure due to random white-space at ends
liverTreat$Treatment<- "liverTreat"

liverInter<- read.csv("https://raw.githubusercontent.com/samarth8392/SFDTranscriptomics/main/DataFiles/Liver/genelist/Liver_DGE_Inter_withAnn.csv")
liverInter$GeneID<- trimws(liverInter$GeneID)
liverInter$Treatment<- "liverInter"

kidneyInter<- read.csv("https://raw.githubusercontent.com/samarth8392/SFDTranscriptomics/main/DataFiles/Kidney/genelist/Kidney_DGE_Inter_withAnn.csv")
kidneyInter$GeneID<- trimws(kidneyInter$GeneID)
kidneyInter$Treatment<- "kidneyInter"

kidneyTreat<- read.csv("https://raw.githubusercontent.com/samarth8392/SFDTranscriptomics/main/DataFiles/Kidney/genelist/Kidney_DGE_TreatOnly_withAnn.csv")
kidneyTreat$GeneID<- trimws(kidneyTreat$GeneID)
kidneyTreat$Treatment<- "kidneyTreat"
skinTreat<- read.csv("https://raw.githubusercontent.com/samarth8392/SFDTranscriptomics/main/DataFiles/Skin/genelist/Skin_DGE_TreatOnly_withAnn.csv")
skinTreat$GeneID<- trimws(skinTreat$GeneID)
skinTreat$Treatment<- "skinTreat"

#create a list of unique gene IDs for possible disease genes
total<- rbind(liverTreat, liverInter, kidneyTreat, kidneyInter, skinTreat)
total<- total %>%
  dplyr::rename(Cviridis_geneID=GeneID) %>%
  select(Cviridis_geneID,Treatment,log2FoldChange,padj) %>%
  right_join(scate.summary.out2,by="Cviridis_geneID") %>%
  mutate(gene_pop=paste0(geneID,pop)) %>%
  filter(!duplicated(gene_pop)) %>%
  filter(tajima.ns.outlier!=0) %>%
  mutate(reg=ifelse(log2FoldChange>0,"Up","Down"))


ggplot(total,aes(x=log2FoldChange, y=tajima.ns, colour=pop, shape=as.factor(tajima.ns.outlier))) +
  geom_point() + facet_wrap(~Treatment)

total2<- total %>% 
  group_by(pop,reg) %>%
  summarise(tajima.pos=sum(tajima.ns.outlier[tajima.ns.outlier==1]), 
            tajima.neg=sum(abs(tajima.ns.outlier[tajima.ns.outlier==-1]))) %>%
  gather(key="tajima_direction",value="number_genes",-pop,-reg) %>%
  mutate(pop=factor(pop, levels=c("STER","KLDR","CEBO")),
         tajima_direction=factor(tajima_direction,levels=c("tajima.neg","tajima.pos")))


degs1<- ggplot(total2,aes(x=reg,y=number_genes, fill=tajima_direction)) + 
  geom_bar(stat="identity", position=dodge) + 
  facet_wrap(~pop) +
  peri_figure +
  scale_y_continuous(expand=c(0,0), limits=c(0,max(total2$number_genes)+17), breaks=seq(0,max(total2$number_genes)+17,by=20)) + 
  labs(x="Expression Direction",
       y="Number Tajima's D\noutlier genes",
       fill="Tajima's D") + 
  scale_fill_manual(labels=c("< 0","> 0"), values=c("deepskyblue3","red3")) + 
  geom_text(data=total2,aes(x=reg,y=number_genes+4, colour=tajima_direction, label=number_genes), position=dodge, show.legend = FALSE, size=geom.text.size) +
   scale_colour_manual(values=c("deepskyblue3","red3"))

degs_ster<- total %>%
  filter(tajima.ns.outlier!=0,pop=="STER")
degs_kldr<- total %>%
  filter(tajima.ns.outlier!=0, pop=="KLDR") %>%
  filter(geneID %in% degs_ster$geneID)
degs_ster <- degs_ster %>%
  filter(geneID %in% degs_kldr$geneID)
degs_sis<- rbind(degs_ster,degs_kldr)
degs_sis<- degs_sis %>% 
  group_by(pop,reg) %>%
  summarise(tajima.pos=sum(tajima.ns.outlier[tajima.ns.outlier==1]), 
            tajima.neg=sum(abs(tajima.ns.outlier[tajima.ns.outlier==-1]))) %>%
  gather(key="tajima_direction",value="number_genes",-pop,-reg) %>%
  mutate(pop=factor(pop, levels=c("STER","KLDR","CEBO")),
         tajima_direction=factor(tajima_direction,levels=c("tajima.neg","tajima.pos")))



degs2<- ggplot(degs_sis,aes(x=reg,y=number_genes, fill=tajima_direction)) + 
  geom_bar(stat="identity", position=dodge) + 
  facet_wrap(~pop) +
  peri_figure +
  scale_y_continuous(expand=c(0,0), limits=c(0,max(degs_sis$number_genes)+4), breaks=seq(0,max(degs_sis$number_genes)+7,by=5)) + 
  labs(x="Expression Direction",
       y="Number Tajima's D\noutlier genes",
       fill="Tajima's D") + 
  scale_fill_manual(labels=c("< 0","> 0"),  values=c("deepskyblue3","red3")) + 
  geom_text(data=degs_sis,aes(x=reg,y=number_genes+1, colour=tajima_direction, label=number_genes), position=dodge, show.legend = FALSE, size=geom.text.size) + 
  scale_colour_manual(values=c("deepskyblue3","red3"))

degs_plot<- ggarrange(degs1,degs2,labels=c("A","B"), nrow=2)
ggsave(filename=paste0(outdir,"/fig_deg_direction.png"),plot = degs_plot,height=150, width=150, units="mm", device="png", bg="white")


total3<- total %>% 
  group_by(pop,reg,Treatment) %>%
  summarise(tajima.pos=sum(tajima.ns.outlier[tajima.ns.outlier==1]), 
            tajima.neg=sum(abs(tajima.ns.outlier[tajima.ns.outlier==-1]))) %>%
  gather(key="tajima_direction",value="number_genes",-pop,-reg,-Treatment) %>%
  mutate(pop=factor(pop, levels=c("STER","KLDR","CEBO")),
         tajima_direction=factor(tajima_direction,levels=c("tajima.neg","tajima.pos")))

ggplot(total3,aes(x=reg,y=number_genes, fill=tajima_direction)) + 
  geom_bar(stat="identity", position=position_dodge(width=0.9)) + 
  facet_wrap(Treatment~pop, ncol=3, scales="free_y") +
  peri_figure +
  scale_y_continuous(breaks = ~round(unique(pretty(.)))) + 
  labs(x="Expression Direction",
       y="Number Tajima's D\noutlier genes",
       fill="Tajima's D") + 
  scale_fill_manual(labels=c("< 0","> 0"), values=c("deepskyblue3","red3"))

```

```{r fig-pNpS, eval=FALSE}
#| label: fig-pNpS
#| fig.cap: "A) number of polymorphic non-synonymous sites/number polymorphic synonymous sites B) Pn/Ps is not related to Tajima's D C) Genes with positive DoS (+1) do not appear to show strong signals of positive selection with Tajima's D D) Relationship between DoS and Pn/Ps"
#| echo: false
#| warning: false
#| eval: false
a<- ggplot(scate.summary[which(!is.na(scate.summary$pNpS) & !is.infinite(scate.summary$pNpS)),], aes(x=immune_DEG,y=pNpS, colour=pop)) +
  geom_boxplot(lwd=baselwd, show.legend = FALSE,position=position_dodge(width=0.9),outlier.colour = NA) + 
  geom_point(position=position_dodge(width=0.9), alpha=0.3) + 
  geom_hline(yintercept=1, lwd=baselwd+0.3, colour="salmon", linetype="dashed") +
  peri_figure +
  labs(x="",y=expression(over(p[N],p[S]))) +
  theme(axis.title.y = element_text(angle=0, hjust=1,vjust=0.5),
        legend.title = element_blank())+
  scale_color_manual(values=viridispairs[c(3,6,12)],name="") + 
  stat_summary(fun = "mean",geom = "point", pch=4, stroke=0.5, size=3, position=position_dodge(width=0.9),show.legend = FALSE) + 
  scale_y_continuous(breaks=seq(0,15,1))


#b<- ggplot(scate.summary[which(!is.na(scate.summary$pnps.pi) & !is.infinite(scate.summary$pnps.pi)),], aes(x=immune_DEG,y=pnps.pi, colour=pop)) +
#  geom_boxplot(lwd=baselwd, show.legend = FALSE,position=position_dodge(width=0.9),outlier.colour = NA) + 
#  geom_point(position=position_dodge(width=0.9), alpha=0.3) + 
#  geom_hline(yintercept=1, lwd=baselwd+0.3, colour="salmon", linetype="dashed") +
#  peri_figure +
#  labs(x="",y=expression(over(pi[N],pi[S]))) +
#  theme(axis.title.y = element_text(angle=0, hjust=1,vjust=0.5),
#        legend.title = element_blank())+
#  scale_color_manual(values=viridispairs[c(3,6,12)],name="") + 
#  stat_summary(fun = "mean",geom = "point", pch=4, stroke=0.5, size=3, position=position_dodge(width=0.9),show.legend = FALSE)  + 
#  scale_y_continuous(breaks=seq(0,18,2))

b<- ggplot(scate.summary.out[which(!is.na(scate.summary.out$pNpS) & !is.infinite(scate.summary.out$pNpS)),], aes(x=tajima.ns,y=pNpS, colour=pop, shape=as.factor(tajima.ns.outlier))) +
  geom_point(alpha=0.5) + 
   peri_figure +
  labs(x="Non-synonymous Tajima's D",y=expression(over(p[N],p[S]))) +
  theme(axis.title.y = element_text(angle=0, hjust=1,vjust=0.5),
        legend.title = element_blank())+
  scale_color_manual(values=viridispairs[c(3,6,12)],name="")


dos<- read.csv(paste0(datadir,"/Mathur_2023_PNAS/DoS.KLDRig.csv"))
dos<- dos %>%
  right_join(scate.summary.out, by="geneID")

c<-ggplot(dos, aes(x=dos1,y=tajima.ns,colour = pop,shape=as.factor(tajima.ns.outlier))) + 
  geom_point(alpha=0.5) + peri_figure +
  labs(x="Direction of Selection (DoS)",y="Non-synonymous Tajima's D") + 
  theme(legend.title = element_blank()) + 
  scale_color_manual(values=viridispairs[c(3,6,12)],name="")

d<- ggplot(dos[which(!is.na(dos$pNpS) & !is.infinite(dos$pNpS)),], aes(x=dos1,y=pNpS,colour = pop,shape=as.factor(tajima.ns.outlier))) + 
  geom_point(alpha=0.5) + peri_figure + 
  labs(x="Direction of Selection (DoS)",y=expression(over(p[N],p[S]))) + 
  theme(legend.title = element_blank(),
        axis.title.y = element_text(angle=0, hjust=1,vjust=0.5)) + 
  scale_color_manual(values=viridispairs[c(3,6,12)],name="")

ggarrange(a,ggarrange(b,c,d, common.legend = TRUE, labels=c("B","C","D"), ncol=3), labels=c("A",""), nrow=2)
```

```{r fig-within-GO-STER, eval=FALSE}
#| label: fig-within-GO-STER
#| fig.cap: "Top GO terms for genes that have outlying non-synonymous Tajmia's D statistic."
#| warning: false
#read in the conversion between the parsed genenames and the ensembl IDs. I filled in missing ones based on manual search (Many of these were due to non-systematic errors in the gene name used for the original search)
conv.ac<- read.csv(paste0(datadir,"/lists/gProfiler_acarolinensis_04-08-2024_10-42-48.csv"))
colnames(conv.ac)[colnames(conv.ac)=="initial_alias"]<- "geneName_parsed"
conv.ac2<- merge(scate.summary.out2, conv.ac, by="geneName_parsed")

### STER
ster.go<- conv.ac2 %>%
  filter(pop=="STER" & converted_alias_manual_additions!="") %>%
  mutate(gene_pop=paste0(geneID,pop)) %>%
  filter(!duplicated(gene_pop) & tajima.ns.outlier==1)
go_res<- gost(ster.go$converted_alias_manual_additions, organism="acarolinensis", significant=TRUE, evcodes=TRUE)
go_res<- go_res$result
go_res<- go_res[order(go_res$p_value),]
go_res$pop<- "STER"
go_res$parents<- as.character(go_res$parents)
write.table(go_res, paste0(outdir,"/GO_STER_Tajima.tsv"), quote=FALSE, row.names = FALSE, sep="\t")

### KLDR GO
kldr.go<- conv.ac2 %>%
  filter(pop=="KLDR" & converted_alias_manual_additions!="") %>%
  mutate(gene_pop=paste0(geneID,pop)) %>%
  filter(!duplicated(gene_pop) & tajima.ns.outlier==1)
go_res<- gost(kldr.go$converted_alias_manual_additions, organism="acarolinensis", significant=TRUE, evcodes=TRUE)
go_res<- go_res$result
go_res<- go_res[order(go_res$p_value),]
go_res$pop<- "KLDR"
go_res$parents<- as.character(go_res$parents)
write.table(go_res, paste0(outdir,"/GO_KLDR_Tajima.tsv"), quote=FALSE, row.names = FALSE, sep="\t")


### CEBO GO
cebo.go<- conv.ac2 %>%
  filter(pop=="CEBO" & converted_alias_manual_additions!="") %>%
  mutate(gene_pop=paste0(geneID,pop)) %>%
  filter(!duplicated(gene_pop) & tajima.ns.outlier==1)
go_res<- gost(cebo.go$converted_alias_manual_additions, organism="acarolinensis", significant=FALSE, evcodes=TRUE)
go_res<- go_res$result
go_res<- go_res[order(go_res$p_value),]
go_res$pop<- "CEBO"
go_res$parents<- as.character(go_res$parents)
go_res<- go_res[1:10,]
write.table(go_res, paste0(outdir,"/GO_CEBO_Tajima.tsv"), quote=FALSE, row.names = FALSE, sep="\t")


rownames(go_res)<-1:nrow(go_res)

#weirdo<- go_res[go_res$intersection=="ENSACAG00000013078",] #just making sure we're getting all the genes.

#go_res<- rbind(go_res,weirdo)
go_out<- list()
for(i in 1:nrow(go_res)){
  sub<- go_res[i,]
  if(grepl(",",sub$intersection)){
    #i=1
    ensg<- str_split_1(sub$intersection,"\\,")
    sub2<- data.frame(p_value=rep(sub$p_value,length(ensg)),
                    source=rep(sub$source, length(ensg)),
                    term_name=rep(sub$term_name, length(ensg)),
                    converted_alias_manual_additions=ensg)
    go_out[[i]]<- sub2
  }else{
    sub$converted_alias_manual_additions<- sub$intersection

    go_out[[i]]<- sub[,c("p_value","source","term_name","converted_alias_manual_additions")]
  }
}
go_res2<- do.call("rbind",go_out)
go_res2$term<- paste(go_res2$source,go_res2$term_name)

go_res2<- merge(go_res2,conv.ac2, by="converted_alias_manual_additions")

go_res2$gene<- paste(go_res2$geneID, go_res2$geneName_parsed)

ggplot(go_res[1:10], aes(x=,y=term,fill=p_value)) + 
  geom_tile() + peri_tile + coord_equal() +
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5)) + 
  labs(x="",y="")

imm<- scate.summary.out2[grepl("immune",scate.summary.out2$immune_DEG),]
```

```{r, eval=FALSE}
go_res<- list()

for(p in pop3){
  go<-read.table(paste0(outdir,"/GO_",p,"_Tajima.tsv"), sep="\t", header=TRUE)
  go_res[[p]]<- go[which(!grepl("KEGG",go$term_id)),]
}
go_res<- do.call("rbind",go_res)

go_freq<- go_res %>%
  group_by(term_id) %>%
  summarise(Freq=n())
simMatrix <- calculateSimMatrix(go_freq$term_id,
                                orgdb="org.Hs.eg.db",
                                ont="BP",
                                method="Rel")
scores <- setNames(go_freq$Freq, go_freq$term_id)
reducedTerms <- reduceSimMatrix(simMatrix,
                               scores,
                                threshold=0.7,
                                orgdb="org.Hs.eg.db")

colnames(reducedTerms)[colnames(reducedTerms)=="go"]<- "term_id"

go<- go_res %>%
  left_join(reducedTerms, by="term_id")
edox2 <- enrichplot::pairwise_termsim(go)
ggplot(go_res, aes(x=pop,y=term_name, fill = intersection_size)) + geom_tile(colour="grey50") + scale_fill_viridis(option="A", na.value="white", direction=-1, name="Number genes") + peri_tile
``` 

### GO

```{r gene annot}
#reading the DEGs and their annotations in again so that I don't need to read all of the above code to arrive at this point.
scate.summary.out2<- read.csv(paste0(outdir,"/report/result_pi_het.csv"))
degs2anolis<- read.xlsx(paste0(datadir,"/GO_annot/Scate_DEGs_annot.xlsx"))
degs2anolis<- degs2anolis %>%
  mutate(Scate=gsub("-RA","",Scate),
         Anolis_GeneID_GCF.GAF=as.character(Anolis_GeneID_GCF.GAF),
         gene_composite=paste0(Anolis_GeneID_GCF.GAF,Scate)) %>%
  select(Scate,geneName_parsed,Anolis_GeneID_GCF.GAF,Anolis_entrez,gene_composite)

#the GO database linking GO IDs to terms
obo<- read.csv(paste0(datadir,"/GO_annot/go_ontology_obo_out.csv"))
obo <- obo %>% dplyr::rename(GO_ID=id)

#this links Anolis genes to gene ontology IDs
gaf<- read.table(paste0(datadir,"/GO_annot/GCF_035594765.1-RS_2024_02_gene_ontology.gaf"), comment.char="!",header=TRUE, sep="\t", quote="\"")
gaf<- gaf %>%
  left_join(obo, by="GO_ID") %>%
  filter(namespace=="biological_process") %>%
  mutate(Anolis_GeneID_GCF.GAF=as.character(GeneID)) %>%
  full_join(degs2anolis[,c("Scate","Anolis_GeneID_GCF.GAF")],by="Anolis_GeneID_GCF.GAF") %>%
  mutate(gene_GO=ifelse(!is.na(Scate),
                        paste0(Scate,GeneID,GO_ID),
                        paste0("",GeneID,GO_ID))) %>%
  filter(!duplicated(gene_GO)) %>%
  select(Scate,GeneID,GO_ID,name,gene_GO) 



#go2gene_bp<- gaf %>%
#  filter(namespace=="biological_process") %>%
#  select(GO_ID,GeneID)
go2name_bp<- obo %>%
  filter(namespace=="biological_process") %>%
  select(GO_ID,name)

#reading in the candidate immune genes 
immune_terms<- read.table(paste0(datadir,"/candidate genes/v2/Cvi_immune_annot.txt"), header=TRUE, sep="\t", quote="\"") 

immune_terms <- immune_terms %>%
  mutate(Scate=gsub("-RA","",Scate)) %>%
  filter(Scate!="") %>%
  filter(Anolis_entrez!="100559978")
### TLR5 is a scate annotation.
tlr5<- read.csv(paste0(datadir,"/candidate genes/v2/sca_TLR5.csv"))
immune_terms<- rbind(immune_terms,tlr5)

immune_GO<- immune_terms %>%
  filter(GO_ID!="") %>%
  select(Scate,Anolis_entrez,Anolis_Genename,Anolis_entrezName,GO_ID) %>%
  inner_join(go2name_bp, by="GO_ID") %>% #add in GO name info, only the biological process terms that exist in the OBO
  mutate(gene_GO=paste0(Scate,GO_ID)) %>%
  filter(!duplicated(gene_GO)) %>%
  full_join(degs2anolis, by="Scate")%>% #add the gene name conversion information including non-immune genes. 
  mutate(gene_GO=paste0(Scate,GO_ID))  %>%
  filter(!duplicated(gene_GO)) 


  

go2gene_bp<- immune_GO %>%
  dplyr::rename(GeneID=Anolis_GeneID_GCF.GAF) %>%
  select(Scate,GeneID,GO_ID,name) %>%
  mutate(gene_GO=paste0(Scate,GeneID,GO_ID)) %>%
  mutate(gene_GO=gsub("^NA","",gene_GO)) %>%
  full_join(gaf,by="gene_GO") %>%
  mutate(Scate=ifelse(is.na(Scate.x),Scate.y,Scate.x),
         GeneID=ifelse(is.na(GeneID.x),GeneID.y,GeneID.x),
         GO_ID=ifelse(is.na(GO_ID.x),GO_ID.y,GO_ID.x)) %>%
  mutate(GeneID=replace_na(GeneID,""),
         Scate=replace_na(Scate,"")) %>%
  mutate(gene_composite=paste0(GeneID,Scate)) %>%
  select(GO_ID,gene_composite)




  
immune_KEGG<- immune_terms %>%
  filter(!is.na(Anolis_pathway_ID) | !is.na(Python_pathway_ID) | !is.na(Thamnophis_pathway_ID)) %>%
  select(!GO_ID) %>% 
  mutate(No_Anolis=ifelse(
    is.na(Anolis_pathway_ID),TRUE,FALSE),
    Pathway_combined=ifelse(
    !is.na(Anolis_pathway_name),
    as.character(Anolis_pathway_name),
    ifelse(!is.na(Thamnophis_pathway_name),
           as.character(Thamnophis_pathway_name),
           as.character(Python_pathway_name)
    ))) 
```
 
#### Tajima's D GO
```{r Tajima STER}
foreground<-  scate.summary.out2 %>%
  filter(pop=="STER") %>%
  filter(tajima.ns.outlier==1 | tajima.ns.outlier==-1) %>%
  dplyr::rename(Scate=geneID) %>%
  left_join(degs2anolis, by="Scate") %>%
  mutate(gene_composite=paste0(Anolis_GeneID_GCF.GAF,Scate)) %>%
  select(gene_composite)

#foreground<- as.character(degs2anolis$Anolis_GeneID_GCF.GAF[which(degs2anolis$Scate %in% foreground$geneID)])
#foreground<- foreground[!is.na(foreground)]


MHC_ster<- scate.summary.out2 %>%
  filter(pop=="STER") %>%
  filter(tajima.ns.outlier==1 | tajima.ns.outlier==-1) %>%
  filter(geneID %in% MHC_paralogs)

go_results<- enricher(foreground$gene_composite, 
                           universe=go2gene_bp$gene_composite, 
                           TERM2GENE = go2gene_bp, 
                           TERM2NAME = go2name_bp)

go_results_ster<- go_results@result
write.table(go_results_ster, paste0(outdir,"/GO_STER_Tajima.tsv"), quote=FALSE, row.names = FALSE, sep="\t")

simMatrix <- calculateSimMatrix(go_results_ster$ID,
                                orgdb="org.Hs.eg.db",
                                ont="BP",
                                method="Rel")
scores <- setNames(go_results_ster$pvalue, go_results_ster$ID)
reducedTerms_ster <- reduceSimMatrix(simMatrix,
                               scores=scores,
                                threshold=0.65,
                                orgdb="org.Hs.eg.db")


reducedTerms_ster <- reducedTerms_ster %>%
  dplyr::rename("ID"="go") %>%
  right_join(go_results_ster,by="ID")
write.table(reducedTerms_ster, paste0(outdir,"/GO_STER_Tajima_rrvgo.tsv"), quote=FALSE, row.names = FALSE, sep="\t")


```

```{r Tajima KLDR}
## KLDR ###
#foreground<-  scate.summary.out2 %>%
#  filter(pop=="KLDR") %>%
#  filter(tajima.ns.outlier==1 | tajima.ns.outlier==-1)
#foreground<- as.character(degs2anolis$Anolis_GeneID_GCF.GAF[#which(degs2anolis$Scate %in% foreground$geneID)])
#foreground<- foreground[!is.na(foreground)]
foreground<-  scate.summary.out2 %>%
  filter(pop=="KLDR") %>%
  filter(tajima.ns.outlier==1 | tajima.ns.outlier==-1) %>%
  dplyr::rename(Scate=geneID) %>%
  left_join(degs2anolis, by="Scate") %>%
  mutate(gene_composite=paste0(Anolis_GeneID_GCF.GAF,Scate)) %>%
  select(gene_composite)

MHC_kldr<- scate.summary.out2 %>%
  filter(pop=="KLDR") %>%
  filter(tajima.ns.outlier==1 | tajima.ns.outlier==-1) %>%
  filter(geneID %in% MHC_paralogs)

go_results<- enricher(foreground$gene_composite, 
                           universe=go2gene_bp$gene_composite, 
                           TERM2GENE = go2gene_bp, 
                           TERM2NAME = go2name_bp)

go_results_kldr<- go_results@result
write.table(go_results_kldr, paste0(outdir,"/GO_KLDR_Tajima.tsv"), quote=FALSE, row.names = FALSE, sep="\t")

simMatrix <- calculateSimMatrix(go_results_kldr$ID,
                                orgdb="org.Hs.eg.db",
                                ont="BP",
                                method="Rel")
scores <- setNames(go_results_kldr$pvalue, go_results_kldr$ID)
reducedTerms_kldr <- reduceSimMatrix(simMatrix,
                               scores=scores,
                                threshold=0.65,
                                orgdb="org.Hs.eg.db")


reducedTerms_kldr <- reducedTerms_kldr %>%
  dplyr::rename("ID"="go") %>%
  right_join(go_results_kldr,by="ID")
write.table(reducedTerms_kldr, paste0(outdir,"/GO_kldr_Tajima_rrvgo.tsv"), quote=FALSE, row.names = FALSE, sep="\t")


```

```{r Tajima CEBO}
### CEBO
foreground<-  scate.summary.out2 %>%
  filter(pop=="CEBO") %>%
  filter(tajima.ns.outlier==1 | tajima.ns.outlier==-1) %>%
  dplyr::rename(Scate=geneID) %>%
  left_join(degs2anolis, by="Scate") %>%
  mutate(gene_composite=paste0(Anolis_GeneID_GCF.GAF,Scate)) %>%
  select(gene_composite)


MHC_cebo<- scate.summary.out2 %>%
  filter(pop=="CEBO") %>%
  filter(tajima.ns.outlier==1 | tajima.ns.outlier==-1) %>%
  filter(geneID %in% MHC_paralogs)



go_results<- enricher(foreground$gene_composite, 
                           universe=go2gene_bp$gene_composite, 
                           TERM2GENE = go2gene_bp, 
                           TERM2NAME = go2name_bp)

go_results_cebo<- go_results@result
write.table(go_results_cebo, paste0(outdir,"/GO_CEBO_Tajima.tsv"), quote=FALSE, row.names = FALSE, sep="\t")

simMatrix <- calculateSimMatrix(go_results_cebo$ID,
                                orgdb="org.Hs.eg.db",
                                ont="BP",
                                method="Rel")
scores <- setNames(go_results_cebo$pvalue, go_results_cebo$ID)
reducedTerms_cebo <- reduceSimMatrix(simMatrix,
                               scores,
                                threshold=0.65,
                                orgdb="org.Hs.eg.db")


reducedTerms_cebo <- reducedTerms_cebo %>%
  dplyr::rename("ID"="go") %>%
  right_join(go_results_cebo,by="ID")
write.table(reducedTerms_cebo, paste0(outdir,"/GO_CEBO_Tajima_rrvgo.tsv"), quote=FALSE, row.names = FALSE, sep="\t")


```

```{r summarize Tajima GO}
### sumarise
go_cebo<- read.table(paste0(outdir,"/GO_CEBO_Tajima.tsv"), sep="\t", header=TRUE)
go_kldr<- read.table(paste0(outdir,"/GO_KLDR_Tajima.tsv"),sep="\t", header=TRUE)
go_ster<- read.table(paste0(outdir,"/GO_STER_Tajima.tsv"),sep="\t", header=TRUE)
go_cebo$pop<- "CEBO"
go_kldr$pop<- "KLDR"
go_ster$pop<- "STER"
go_results<- rbind(go_cebo,go_kldr,go_ster)
go_results<- go_results %>%
  filter(p.adjust<0.05 & Count>1) %>%
  mutate(immune=ifelse(ID %in% immune_GO$GO_ID,
                       "immune",
                       "non-immune"),
        Description=ifelse(grepl("MHC",Description),
                           "antigen processing and presentation\nof peptide antigen via MHC class I",as.character(Description))) %>%
  mutate(Description=ifelse(
    grepl("calcium-dependent cell-cell",Description),
    "calcium-dependent cell-cell adhesion via\nplasma membrane cell adhesion molecules",
    Description)) %>%
  mutate(Description=ifelse(
    grepl("homophilic",Description),
    "homophilic cell adhesion via\nplasma membrane adhesion molecules",
    Description)) %>%
  mutate(Description=ifelse(
    grepl("transmembrane receptor",Description),
    "transmembrane receptor protein\ntyrosine kinase signaling pathway",
    Description))
  

ggplot(go_results, aes(x=Description, y=pop, colour=pvalue, size=Count)) + 
  geom_point() + peri_figure + 
  scale_size_continuous(breaks=seq(2,12,by=2),range = c(1,3)) +
  labs(x="",y="") +
  theme(axis.text.x = element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5),
        legend.justification="top",
        panel.spacing = unit(0.1, "lines")) +
  scale_colour_viridis(option="inferno") +
  facet_grid(~immune, scales="free")


ggsave(filename=paste0(outdir,"/fig_Tajima_GO_sig.png"), plot=last_plot(), device="png",bg="white",width=260,height=110, units="mm") 



simMatrix <- calculateSimMatrix(go_results$ID,
                                orgdb="org.Hs.eg.db",
                                ont="BP",
                                method="Rel")
scores <- setNames(go_results$pvalue, go_results$ID)
reducedTerms<- reduceSimMatrix(simMatrix,
                               "uniqueness",
                                threshold=0.65,
                                orgdb="org.Hs.eg.db")


reducedTerms <- reducedTerms %>%
  dplyr::rename("ID"="go") %>%
  right_join(go_results,by="ID")

reducedTerms2<- reducedTerms %>%
  group_by(pop,parentTerm) %>%
  summarise(n_genes=sum(Count),mean_p=mean(pvalue)) %>%
   mutate(parentTerm=ifelse(grepl("MHC",parentTerm),
                           "antigen processing and presentation\nof peptide antigen via MHC class I",as.character(parentTerm)))%>%
  mutate(parentTerm=gsub("phosphatidylinositol 3-kinase/protein kinase B signal transduction",
                         "phosphatidylinositol 3-kinase/protein\nkinase B signal transduction",
                         parentTerm))

library(scales)
ggplot(reducedTerms2, aes(x=parentTerm, y=pop, colour=mean_p, size=n_genes)) + 
  geom_point() + peri_figure + 
  scale_size_continuous(breaks=seq(2,32,by=6),range = c(1,4), name="number genes") +
  labs(x="",y="") +
  theme(axis.text.x = element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5), 
        strip.text.x = element_blank(),
        legend.justification="top",
        panel.spacing = unit(0.1, "lines")) +
  scale_colour_viridis(option="inferno",name="mean p-value")

write.table(reducedTerms, paste0(outdir,"/GO_ALL_Tajima_rrvgo.tsv"), quote=FALSE, row.names = FALSE, sep="\t")


ggsave(filename=paste0(outdir,"/fig_Tajima_GO_sig_rrvgo.png"), plot=last_plot(), device="png",bg="white",width=190,height=100, units="mm") 
```

###### Immune genes

```{r significant immune genes}
go_results<- go_results[go_results$ID %in% immune_GO$GO_ID,]

ggplot(go_results, aes(x=Description, y=pop, colour=pvalue, size=Count)) + 
  geom_point() + peri_figure + 
  scale_size_continuous(breaks=seq(2,12,by=2),range = c(1,3)) +
  labs(x="",y="") +
  theme(axis.text.x = element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5), 
        strip.text.x = element_blank(),
        legend.justification="top",
        panel.spacing = unit(0.1, "lines")) +
  scale_colour_viridis(option="inferno")


ggsave(filename=paste0(outdir,"/fig_Tajima_GO_immune_sig.png"), plot=last_plot(), device="png",bg="white",width=250,height=130, units="mm")
```


Look at immune functions among outliers irrespective of enrichment


```{r Tajima's D Outlier Immune Functions GO}
outliers_GO<- scate.summary.out2 %>%
  filter(tajima.ns.outlier!=0) %>%
  dplyr::rename(Scate=geneID) %>%
  left_join(degs2anolis, by="Scate") %>%
  inner_join(immune_GO, by="Scate") %>%
  select(Scate,pop,immune_DEG,tajima.ns,tajima.ns.outlier,geneName_parsed.y,GO_ID,name) %>%
  dplyr::rename(geneName_parsed=geneName_parsed.y) %>%
  mutate(geneName_parsed=ifelse(grepl("MHC Class I", geneName_parsed),
                                paste0("MHCIA-",Scate), as.character(geneName_parsed)),
         gene_pop=paste0(geneName_parsed,pop)) %>%
  filter(!is.na(GO_ID))


simMatrix <- calculateSimMatrix(outliers_GO$GO_ID,
                                orgdb="org.Hs.eg.db",
                                ont="BP",
                                method="Rel")
scores <- setNames(abs(outliers_GO$tajima.ns), outliers_GO$GO_ID)
reducedTerms_immuneGO <- reduceSimMatrix(simMatrix,
                               "uniqueness",
                                threshold=0.65,
                                orgdb="org.Hs.eg.db") 
outliers_GO2<- reducedTerms_immuneGO %>%
  dplyr::rename(GO_ID=go) %>%
  right_join(outliers_GO, by="GO_ID") 

outliers_GO2 <- outliers_GO2 %>%
  mutate(parentTerm=ifelse(grepl("MHC",parentTerm),
                           "antigen processing and presentation\nof peptide antigen via MHC class I",as.character(parentTerm))) %>%
  mutate(parentTerm=ifelse(grepl("CD8-positive",parentTerm),
                           "CD8-positive, alpha-beta T cell differentiation\ninvolved in immune response",as.character(parentTerm)))
#outliers_GO<- outliers_GO %>%
#  mutate(axis.colour=ifelse(geneName_parsed %in% overlaps,
#                            "grey50","black"))


ggplot(outliers_GO,
       aes(x=geneName_parsed,y=name,fill=tajima.ns)) + 
  geom_tile() + peri_tile + 
  facet_grid(~pop, scales="free", space="free") +
  theme(axis.text.x = element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5,
                                   size=basesize-1),
        legend.position = "top") +
  scale_fill_gradient2(low="deepskyblue3",
                       high="red3",
                       na.value = "white",
                       mid="azure") +
  labs(y="simplified GO term",
       x="",
       fill="Tajima's D")

ggsave(filename=paste0(outdir,"/fig_Tajima_GO_immune.png"), plot=last_plot(), device="png",bg="white",width=250,height=100, units="mm") 

```

```{r Tajima immune KEGG}
outliers_KEGG<- scate.summary.out2 %>%
  filter(tajima.ns.outlier!=0) %>%
  dplyr::rename(Scate=geneID) %>%
  left_join(degs2anolis, by="Scate") %>%
  inner_join(immune_KEGG, by="Scate") %>%
  dplyr::rename(geneName_parsed=geneName_parsed.y) %>%
  mutate(geneName_parsed=ifelse(grepl("MHC Class I", geneName_parsed),
                                paste0("MHCIA-",Scate), as.character(geneName_parsed)),
         gene_pop=paste0(geneName_parsed,pop))

ggplot(outliers_KEGG,
       aes(x=geneName_parsed,y=Pathway_combined,fill=tajima.ns)) + 
  geom_tile() + peri_tile + 
  facet_grid(~pop, scales="free", space="free") +
  theme(axis.text.x = element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5,
                                   size=basesize-1),
        legend.position = "top") +
  scale_fill_gradient2(low="deepskyblue3",
                       high="red3",
                       na.value = "white",
                       mid="azure") +
  labs(y="KEGG pathway",
       x="",
       fill="Tajima's D")



ggsave(filename=paste0(outdir,"/fig_Tajima_KEGG_immune.png"), plot=last_plot(), device="png",bg="white",width=220,height=120, units="mm") 
```

```{r summary immune terms}
outliers_KEGG<- outliers_KEGG %>%
  select(Scate,geneName_parsed,tajima.ns,tajima.ns.outlier,Pathway_combined,pop) %>%
  dplyr::rename(term=Pathway_combined) %>%
  mutate(db="KEGG",
         gene_pop_path=paste0(Scate,pop,term)) %>%
  filter(!duplicated(gene_pop_path)) %>%
  select(!gene_pop_path)

outliers_GO2<- outliers_GO2 %>%
  select(Scate,geneName_parsed,tajima.ns,tajima.ns.outlier,parentTerm,pop) %>%
  dplyr::rename(term=parentTerm) %>%
  mutate(db="GO")

outliers_tajima<- rbind(outliers_KEGG,outliers_GO2)
outliers_tajima$pop<- factor(outliers_tajima$pop, levels=c("STER","KLDR","CEBO"))
outliers_tajima_summary<-  outliers_tajima %>%
  group_by(pop,term,db) %>%
  summarise(n_positive=sum(tajima.ns.outlier[tajima.ns.outlier==1]),
            n_negative=sum(abs(tajima.ns.outlier[tajima.ns.outlier==-1]))) %>%
  pivot_longer(c(n_positive,n_negative), values_to = "count")


tajima_immune_plot<- ggplot(outliers_tajima,
       aes(x=term,y=geneName_parsed,fill=tajima.ns)) + 
  geom_tile() + peri_tile + 
  facet_grid(pop ~ db, scales="free", space="free") + 
  theme(axis.text.x = element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5,
                                   size=basesize-1),
        axis.text.y=element_text(size=basesize-1),
        legend.position = "top",
        legend.key.width=unit(5,"mm"),
        plot.margin = margin(t = 5,  # Top margin
                                          r = 5,  # Right margin
                                          b = 1,  # Bottom margin
                                          l = 5, unit="mm")) +
  scale_fill_gradient2(low="deepskyblue3",
                       high="red3",
                       na.value = "white",
                       mid="azure") +
  labs(y="",
       x="",
       fill="Tajima's D")


tajima_immune_plot
ggsave(filename=paste0( outdir,"/figure2.png"), plot=last_plot(), device="png",bg="white",width=215,height=310, units="mm") 



a<- ggplot(outliers_tajima_summary,aes(x=term,y=count,fill=name)) + 
  geom_bar(stat="identity", position=position_dodge(width=0.9)) +
  facet_grid(db~pop, scales="free") + 
  peri_figure + 
    theme(axis.text.x = element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5,
                                   size=basesize),
          axis.text.y = element_text(hjust=0.5),
        legend.position = "none",
        legend.key.width=unit(5,"mm"),
        plot.margin = margin(t = 5,  # Top margin
                                          r = 5,  # Right margin
                                          b = 5,  # Bottom margin
                                          l = 0, unit="mm")) +
  scale_fill_manual(values=c("deepskyblue3","red3")) + 
  labs(x="",
       y="") +
   coord_flip()

i=unique(outliers_tajima$term)[1]
term_overlap<- list() 
for(i in unique(outliers_tajima$term)){
  sub<- outliers_tajima[which(outliers_tajima$term==i),]
  dups<- sub$Scate[duplicated(sub$Scate)]
  sub<- sub %>%
    mutate(geneName_parsed=ifelse(!Scate %in% dups,NA,geneName_parsed),
           tajima.ns.outlier=ifelse(!Scate %in% dups,NA,tajima.ns.outlier))
  
  term_overlap[[i]]<- sub
}
term_overlap<- as.data.frame(do.call("rbind",term_overlap))
term_overlap$pop<- factor(term_overlap$pop,levels=c("STER","KLDR","CEBO"))

b<- ggplot(term_overlap,aes(x=geneName_parsed,y=term, fill=as.factor(tajima.ns.outlier))) + 
  geom_tile() + 
  facet_grid(db~pop, scales="free",space="free") +
   peri_tile +
  theme(axis.text.x = element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5,
                                   size=basesize),
        strip.background = element_blank(),
        legend.position="none",
        axis.ticks = element_line(colour = "black",
                                    linewidth = 0.1),
        strip.text.y = element_blank(),
        axis.text.y = element_blank(),
        axis.line.y = element_line(colour = "black",
                                  linewidth = baselwd), 
        axis.line.x = element_line(colour = "black",
                                   linewidth = baselwd),
        plot.margin = margin(t = 5,  # Top margin
                             r = 0,  # Right margin
                             b = 5,  # Bottom margin
                             l = 5, unit="mm")) +
  scale_fill_manual(values=c("deepskyblue3","red3"), na.value = "white")+
  scale_y_discrete(position="right") +
  labs(x="",
       y="")



  ggarrange(b,a, ncol=2, nrow=1, align="h", widths=c(1,1.2))
  
ggsave(filename=paste0( outdir,"/fig_overlaps.pdf"), plot=last_plot(), device="pdf",bg="white",width=310,height=215, units="mm") 
```


##### Positive Tajima's D

```{r Pos Tajima STER,eval=FALSE}
foreground<-  scate.summary.out2 %>%
  filter(pop=="STER") %>%
  filter(tajima.ns.outlier==1) %>%
  dplyr::rename(Scate=geneID) %>%
  left_join(degs2anolis, by="Scate") %>%
  mutate(gene_composite=paste0(Anolis_GeneID_GCF.GAF,Scate)) %>%
  select(gene_composite)



MHC_ster<- scate.summary.out2 %>%
  filter(pop=="STER") %>%
  filter(tajima.ns.outlier==1) %>%
  filter(geneID %in% MHC_paralogs)

go_results<- enricher(foreground$gene_composite, 
                           universe=go2gene_bp$gene_composite, 
                           TERM2GENE = go2gene_bp, 
                           TERM2NAME = go2name_bp)

go_results_ster<- go_results@result
write.table(go_results_ster, paste0(outdir,"/GO_STER_Tajima_pos.tsv"), quote=FALSE, row.names = FALSE, sep="\t")

simMatrix <- calculateSimMatrix(go_results_ster$ID,
                                orgdb="org.Hs.eg.db",
                                ont="BP",
                                method="Rel")
scores <- setNames(go_results_ster$pvalue, go_results_ster$ID)
reducedTerms_ster <- reduceSimMatrix(simMatrix,
                               scores=scores,
                                threshold=0.65,
                                orgdb="org.Hs.eg.db")


reducedTerms_ster <- reducedTerms_ster %>%
  dplyr::rename("ID"="go") %>%
  right_join(go_results_ster,by="ID")
write.table(reducedTerms_ster, paste0(outdir,"/GO_STER_Tajima_pos_rrvgo.tsv"), quote=FALSE, row.names = FALSE, sep="\t")


```

```{r Pos Tajima KLDR,eval=FALSE}
foreground<-  scate.summary.out2 %>%
  filter(pop=="KLDR") %>%
  filter(tajima.ns.outlier==1) %>%
  dplyr::rename(Scate=geneID) %>%
  left_join(degs2anolis, by="Scate") %>%
  mutate(gene_composite=paste0(Anolis_GeneID_GCF.GAF,Scate)) %>%
  select(gene_composite)

MHC_kldr<- scate.summary.out2 %>%
  filter(pop=="KLDR") %>%
  filter(tajima.ns.outlier==1) %>%
  filter(geneID %in% MHC_paralogs)

go_results<- enricher(foreground$gene_composite, 
                           universe=go2gene_bp$gene_composite, 
                           TERM2GENE = go2gene_bp, 
                           TERM2NAME = go2name_bp)

go_results_kldr<- go_results@result
write.table(go_results_kldr, paste0(outdir,"/GO_KLDR_Tajima_pos.tsv"), quote=FALSE, row.names = FALSE, sep="\t")

simMatrix <- calculateSimMatrix(go_results_kldr$ID,
                                orgdb="org.Hs.eg.db",
                                ont="BP",
                                method="Rel")
scores <- setNames(go_results_kldr$pvalue, go_results_kldr$ID)
reducedTerms_kldr <- reduceSimMatrix(simMatrix,
                               scores=scores,
                                threshold=0.65,
                                orgdb="org.Hs.eg.db")


reducedTerms_kldr <- reducedTerms_kldr %>%
  dplyr::rename("ID"="go") %>%
  right_join(go_results_kldr,by="ID")
write.table(reducedTerms_kldr, paste0(outdir,"/GO_kldr_Tajima_pos_rrvgo.tsv"), quote=FALSE, row.names = FALSE, sep="\t")


```

```{r Pos Tajima CEBO, eval=FALSE}
### CEBO
foreground<-  scate.summary.out2 %>%
  filter(pop=="CEBO") %>%
  filter(tajima.ns.outlier==1) %>%
  dplyr::rename(Scate=geneID) %>%
  left_join(degs2anolis, by="Scate") %>%
  mutate(gene_composite=paste0(Anolis_GeneID_GCF.GAF,Scate)) %>%
  select(gene_composite)


MHC_cebo<- scate.summary.out2 %>%
  filter(pop=="CEBO") %>%
  filter(tajima.ns.outlier==1) %>%
  filter(geneID %in% MHC_paralogs)



go_results<- enricher(foreground$gene_composite, 
                           universe=go2gene_bp$gene_composite, 
                           TERM2GENE = go2gene_bp, 
                           TERM2NAME = go2name_bp)

go_results_cebo<- go_results@result
write.table(go_results_cebo, paste0(outdir,"/GO_CEBO_Tajima_pos.tsv"), quote=FALSE, row.names = FALSE, sep="\t")

simMatrix <- calculateSimMatrix(go_results_cebo$ID,
                                orgdb="org.Hs.eg.db",
                                ont="BP",
                                method="Rel")
scores <- setNames(go_results_cebo$pvalue, go_results_cebo$ID)
reducedTerms_cebo <- reduceSimMatrix(simMatrix,
                               scores,
                                threshold=0.65,
                                orgdb="org.Hs.eg.db")


reducedTerms_cebo <- reducedTerms_cebo %>%
  dplyr::rename("ID"="go") %>%
  right_join(go_results_cebo,by="ID")
write.table(reducedTerms_cebo, paste0(outdir,"/GO_CEBO_Tajima_pos_rrvgo.tsv"), quote=FALSE, row.names = FALSE, sep="\t")


```

```{r summarize pos Tajima GO}
### sumarise
go_cebo<- read.table(paste0(outdir,"/GO_CEBO_Tajima_pos.tsv"), sep="\t", header=TRUE)
go_kldr<- read.table(paste0(outdir,"/GO_KLDR_Tajima_pos.tsv"),sep="\t", header=TRUE)
go_ster<- read.table(paste0(outdir,"/GO_STER_Tajima_pos.tsv"),sep="\t", header=TRUE)
go_cebo$pop<- "CEBO"
go_kldr$pop<- "KLDR"
go_ster$pop<- "STER"
go_results_pos<- rbind(go_cebo,go_kldr,go_ster)
go_results_pos<- go_results_pos %>%
  filter(p.adjust<0.05 & Count>1) %>%
  mutate(immune=ifelse(ID %in% immune_GO$GO_ID,
                       "immune",
                       "non-immune"),
         D="pos")
         
  

ggplot(go_results_pos, aes(x=Description, y=pop, colour=pvalue, size=Count)) + 
  geom_point() + peri_figure + 
  scale_size_continuous(breaks=seq(2,12,by=2),range = c(1,3)) +
  labs(x="",y="") +
  theme(axis.text.x = element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5),
        legend.justification="top",
        panel.spacing = unit(0.1, "lines")) +
  scale_colour_viridis(option="inferno") +
  facet_grid(~immune, scales="free")


ggsave(filename=paste0(outdir,"/fig_Tajima_pos_GO_sig.png"), plot=last_plot(), device="png",bg="white",width=260,height=110, units="mm") 
```

##### Negative Tajima's D

```{r neg Tajima STER,eval=FALSE}
foreground<-  scate.summary.out2 %>%
  filter(pop=="STER") %>%
  filter(tajima.ns.outlier==-1) %>%
  dplyr::rename(Scate=geneID) %>%
  left_join(degs2anolis, by="Scate") %>%
  mutate(gene_composite=paste0(Anolis_GeneID_GCF.GAF,Scate)) %>%
  select(gene_composite)



MHC_ster<- scate.summary.out2 %>%
  filter(pop=="STER") %>%
  filter(tajima.ns.outlier==-1) %>%
  filter(geneID %in% MHC_paralogs)

go_results<- enricher(foreground$gene_composite, 
                           universe=go2gene_bp$gene_composite, 
                           TERM2GENE = go2gene_bp, 
                           TERM2NAME = go2name_bp)

go_results_ster<- go_results@result
write.table(go_results_ster, paste0(outdir,"/GO_STER_Tajima_neg.tsv"), quote=FALSE, row.names = FALSE, sep="\t")

simMatrix <- calculateSimMatrix(go_results_ster$ID,
                                orgdb="org.Hs.eg.db",
                                ont="BP",
                                method="Rel")
scores <- setNames(go_results_ster$pvalue, go_results_ster$ID)
reducedTerms_ster <- reduceSimMatrix(simMatrix,
                               scores=scores,
                                threshold=0.65,
                                orgdb="org.Hs.eg.db")


reducedTerms_ster <- reducedTerms_ster %>%
  dplyr::rename("ID"="go") %>%
  right_join(go_results_ster,by="ID")
write.table(reducedTerms_ster, paste0(outdir,"/GO_STER_Tajima_neg_rrvgo.tsv"), quote=FALSE, row.names = FALSE, sep="\t")


```

```{r neg Tajima KLDR,eval=FALSE}
foreground<-  scate.summary.out2 %>%
  filter(pop=="KLDR") %>%
  filter(tajima.ns.outlier==-1) %>%
  dplyr::rename(Scate=geneID) %>%
  left_join(degs2anolis, by="Scate") %>%
  mutate(gene_composite=paste0(Anolis_GeneID_GCF.GAF,Scate)) %>%
  select(gene_composite)

MHC_kldr<- scate.summary.out2 %>%
  filter(pop=="KLDR") %>%
  filter(tajima.ns.outlier==-1) %>%
  filter(geneID %in% MHC_paralogs)

go_results<- enricher(foreground$gene_composite, 
                           universe=go2gene_bp$gene_composite, 
                           TERM2GENE = go2gene_bp, 
                           TERM2NAME = go2name_bp)

go_results_kldr<- go_results@result
write.table(go_results_kldr, paste0(outdir,"/GO_KLDR_Tajima_neg.tsv"), quote=FALSE, row.names = FALSE, sep="\t")

simMatrix <- calculateSimMatrix(go_results_kldr$ID,
                                orgdb="org.Hs.eg.db",
                                ont="BP",
                                method="Rel")
scores <- setNames(go_results_kldr$pvalue, go_results_kldr$ID)
reducedTerms_kldr <- reduceSimMatrix(simMatrix,
                               scores=scores,
                                threshold=0.65,
                                orgdb="org.Hs.eg.db")


reducedTerms_kldr <- reducedTerms_kldr %>%
  dplyr::rename("ID"="go") %>%
  right_join(go_results_kldr,by="ID")
write.table(reducedTerms_kldr, paste0(outdir,"/GO_kldr_Tajima_neg_rrvgo.tsv"), quote=FALSE, row.names = FALSE, sep="\t")


```

```{r neg Tajima CEBO, eval=FALSE}
### CEBO
foreground<-  scate.summary.out2 %>%
  filter(pop=="CEBO") %>%
  filter(tajima.ns.outlier==-1) %>%
  dplyr::rename(Scate=geneID) %>%
  left_join(degs2anolis, by="Scate") %>%
  mutate(gene_composite=paste0(Anolis_GeneID_GCF.GAF,Scate)) %>%
  select(gene_composite)


MHC_cebo<- scate.summary.out2 %>%
  filter(pop=="CEBO") %>%
  filter(tajima.ns.outlier==-1) %>%
  filter(geneID %in% MHC_paralogs)



go_results<- enricher(foreground$gene_composite, 
                           universe=go2gene_bp$gene_composite, 
                           TERM2GENE = go2gene_bp, 
                           TERM2NAME = go2name_bp)

go_results_cebo<- go_results@result
write.table(go_results_cebo, paste0(outdir,"/GO_CEBO_Tajima_neg.tsv"), quote=FALSE, row.names = FALSE, sep="\t")

simMatrix <- calculateSimMatrix(go_results_cebo$ID,
                                orgdb="org.Hs.eg.db",
                                ont="BP",
                                method="Rel")
scores <- setNames(go_results_cebo$pvalue, go_results_cebo$ID)
reducedTerms_cebo <- reduceSimMatrix(simMatrix,
                               scores,
                                threshold=0.65,
                                orgdb="org.Hs.eg.db")


reducedTerms_cebo <- reducedTerms_cebo %>%
  dplyr::rename("ID"="go") %>%
  right_join(go_results_cebo,by="ID")
write.table(reducedTerms_cebo, paste0(outdir,"/GO_CEBO_Tajima_neg_rrvgo.tsv"), quote=FALSE, row.names = FALSE, sep="\t")


```

```{r summarize neg Tajima GO}
### sumarise
go_cebo<- read.table(paste0(outdir,"/GO_CEBO_Tajima_neg.tsv"), sep="\t", header=TRUE)
go_kldr<- read.table(paste0(outdir,"/GO_KLDR_Tajima_neg.tsv"),sep="\t", header=TRUE)
go_ster<- read.table(paste0(outdir,"/GO_STER_Tajima_neg.tsv"),sep="\t", header=TRUE)
go_cebo$pop<- "CEBO"
go_kldr$pop<- "KLDR"
go_ster$pop<- "STER"
go_results_neg<- rbind(go_cebo,go_kldr,go_ster)
go_results_neg<- go_results_neg %>%
  filter(p.adjust<0.05 & Count>1) %>%
  mutate(immune=ifelse(ID %in% immune_GO$GO_ID,
                       "immune",
                       "non-immune"),
         D="neg")
         
  

ggplot(go_results_neg, aes(x=Description, y=pop, colour=pvalue, size=Count)) + 
  geom_point() + peri_figure + 
  scale_size_continuous(breaks=seq(2,12,by=2),range = c(1,3)) +
  labs(x="",y="") +
  theme(axis.text.x = element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5),
        legend.justification="top",
        panel.spacing = unit(0.1, "lines")) +
  scale_colour_viridis(option="inferno") +
  facet_grid(~immune, scales="free")


ggsave(filename=paste0(outdir,"/fig_Tajima_neg_GO_sig.png"), plot=last_plot(), device="png",bg="white",width=260,height=110, units="mm") 
```

```{r summarize direction}
directional_go<- rbind(go_results_pos, go_results_neg)

directional_go<- directional_go %>%
  mutate(D=ifelse(D=="pos","D > 0", "D < 0")) %>%
  mutate(immune=ifelse(ID %in% immune_GO$GO_ID,
                       "immune",
                       "non-immune"),
        Description=ifelse(grepl("MHC",Description),
                           "antigen processing and presentation\nof peptide antigen via MHC class I",as.character(Description))) %>%
  mutate(Description=ifelse(
    grepl("calcium-dependent cell-cell",Description),
    "calcium-dependent cell-cell adhesion via\nplasma membrane cell adhesion molecules",
    Description)) %>%
  mutate(Description=ifelse(
    grepl("homophilic",Description),
    "homophilic cell adhesion via\nplasma membrane adhesion molecules",
    Description)) %>%
  mutate(Description=ifelse(
    grepl("transmembrane receptor",Description),
    "transmembrane receptor protein\ntyrosine kinase signaling pathway",
    Description))

ggplot(directional_go, aes(x=Description, y=pop, colour=p.adjust, size=Count)) + 
  geom_point() + peri_figure + 
  scale_size_continuous(breaks=seq(2,12,by=2),range = c(1,3)) +
  labs(x="",y="") +
  theme(axis.text.x = element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5),
        legend.justification="top",
        panel.spacing = unit(0.1, "lines")) +
  scale_colour_viridis(option="inferno") +
  facet_grid(D~immune, scales="free")

ggsave(filename=paste0(outdir,"/fig_Tajima_directional_GO_sig.png"), plot=last_plot(), device="png",bg="white",width=270,height=120, units="mm") 
```

#### Pi GO

```{r pi STER}
foreground<-  scate.summary.out2 %>%
  filter(pop=="STER") %>%
  filter(pi.ns.outlier==1) %>%
  dplyr::rename(Scate=geneID) %>%
  left_join(degs2anolis, by="Scate") %>%
  mutate(gene_composite=paste0(Anolis_GeneID_GCF.GAF,Scate)) %>%
  select(gene_composite)


#foreground<-  scate.summary.out2 %>%
#  filter(pop=="STER") %>%
#  filter(pi.ns.outlier==1)
#foreground<- as.character(degs2anolis$Anolis_GeneID#_GCF.GAF[which(degs2anolis$Scate %in% #foreground$geneID)])
#foreground<- foreground[!is.na(foreground)]


go_results<- enricher(foreground$gene_composite, 
                           universe=go2gene_bp$gene_composite, 
                           TERM2GENE = go2gene_bp, 
                           TERM2NAME = go2name_bp)


go_results_ster<- go_results@result
write.table(go_results_ster, paste0(outdir,"/GO_STER_pi-pos.tsv"), quote=FALSE, row.names = FALSE, sep="\t")

foreground<-  scate.summary.out2 %>%
  filter(pop=="STER") %>%
  filter(pi.ns.outlier==-1) %>%
  dplyr::rename(Scate=geneID) %>%
  left_join(degs2anolis, by="Scate") %>%
  mutate(gene_composite=paste0(Anolis_GeneID_GCF.GAF,Scate)) %>%
  select(gene_composite)

go_results<- enricher(foreground$gene_composite, 
                           universe=go2gene_bp$gene_composite, 
                           TERM2GENE = go2gene_bp, 
                           TERM2NAME = go2name_bp)


go_results_ster<- go_results@result
write.table(go_results_ster, paste0(outdir,"/GO_STER_pi-neg.tsv"), quote=FALSE, row.names = FALSE, sep="\t")

MHC_ster<- scate.summary.out2 %>%
  filter(pop=="STER") %>%
  filter(pi.ns.outlier==1) %>%
  filter(geneID %in% MHC_paralogs)

```

```{r pi kldr}
foreground<-  scate.summary.out2 %>%
  filter(pop=="KLDR") %>%
  filter(pi.ns.outlier==1) %>%
  dplyr::rename(Scate=geneID) %>%
  left_join(degs2anolis, by="Scate") %>%
  mutate(gene_composite=paste0(Anolis_GeneID_GCF.GAF,Scate)) %>%
  select(gene_composite)


go_results<- enricher(foreground$gene_composite, 
                           universe=go2gene_bp$gene_composite, 
                           TERM2GENE = go2gene_bp, 
                           TERM2NAME = go2name_bp)


go_results_kldr<- go_results@result
write.table(go_results_kldr, paste0(outdir,"/GO_KLDR_pi-pos.tsv"), quote=FALSE, row.names = FALSE, sep="\t")

MHC_kldr<- scate.summary.out2 %>%
  filter(pop=="KLDR") %>%
  filter(pi.ns.outlier==1) %>%
  filter(geneID %in% MHC_paralogs)
```

```{r pi CEBO}
foreground<-  scate.summary.out2 %>%
  filter(pop=="CEBO") %>%
  filter(pi.ns.outlier==1) %>%
  dplyr::rename(Scate=geneID) %>%
  left_join(degs2anolis, by="Scate") %>%
  mutate(gene_composite=paste0(Anolis_GeneID_GCF.GAF,Scate)) %>%
  select(gene_composite)


go_results<- enricher(foreground$gene_composite, 
                           universe=go2gene_bp$gene_composite, 
                           TERM2GENE = go2gene_bp, 
                           TERM2NAME = go2name_bp)

go_results_cebo<- go_results@result
write.table(go_results_cebo, paste0(outdir,"/GO_CEBO_pi-pos.tsv"), quote=FALSE, row.names = FALSE, sep="\t")


MHC_cebo<- scate.summary.out2 %>%
  filter(pop=="CEBO") %>%
  filter(pi.ns.outlier==1) %>%
  filter(geneID %in% MHC_paralogs)
```

```{r pi_GO}
go_results_cebo$pop<- "CEBO"
go_results_kldr$pop<- "KLDR"
go_results_ster$pop<- "STER"

go_results<- rbind(go_results_cebo, go_results_kldr, go_results_ster)
go_results<- go_results[which(go_results$pvalue<0.05),]
go_results$pop<- factor(go_results$pop, levels=c("STER","KLDR","CEBO"))
go_results$immune<- ifelse(go_results$ID %in% immune_GO,"immune","other")
go_results$Description<- factor(go_results$Description, levels=unique(go_results$Description[order(go_results$immune)]))
go_results$Description=ifelse(grepl("MHC",go_results$Description),
                           "antigen processing and presentation\nof peptide antigen via MHC class I",as.character(go_results$Description))

ggplot(go_results, aes(x=Description, y=pop, colour=pvalue, size=Count)) + 
  geom_point() + peri_figure + 
  scale_size_continuous(breaks=c(1,2,3),range = c(1,3)) +
  labs(x="",y="") +
  theme(axis.text.x = element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5), 
        strip.text.x = element_blank(),
        legend.justification="top")+
  scale_colour_viridis(option="inferno")



ggsave(filename=paste0(outdir,"/fig_pi_GO_sig.png"), plot=last_plot(), device="png",bg="white",width=100,height=80, units="mm") 


```

```{r immune terms pi,eval=FALSE}

terms_MHC<- rbind(MHC_cebo, MHC_kldr, MHC_ster)
terms_MHC<- terms_MHC %>%
  group_by(pop) %>%
  summarise(n_genes=n()) %>%
  mutate(name="MHC Class I paralog",GO_ID=NA,immune="immune")

terms_MHC$pop<- factor(terms_MHC$pop,levels=c("STER","KLDR","CEBO"))
terms_MHC$total_genes<- c(20,24,20)

mhc_plot<- ggplot(terms_MHC, aes(x=pop,y=(n_genes/total_genes), fill=pop)) + 
  geom_bar(stat="identity") + 
  peri_figure + 
  labs(x="",y=expression(atop(paste("Percent positive ",pi," outlier"),paste("MHC Class I paralogs")))) + 
  scale_fill_manual(values=viridispairs[c(3,6,12)], name="") +
  theme(axis.text.x = element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5),
        legend.position="none") + 
  scale_y_continuous(expand=c(0,0),labels = scales::percent_format(accuracy = 1))


ggsave(filename=paste0(outdir,"/fig_pi_MHC.png"), plot=mhc_plot, device="png",bg="white",width=40,height=60, units="mm") 





pi_venn<- list("STER"=scate.summary.out2$geneID[which(scate.summary.out2$pi.ns.outlier==1 & scate.summary.out2$pop=="STER" & scate.summary.out2$immune_DEG=="immune DEG")],
                "KLDR"=scate.summary.out2$geneID[which(scate.summary.out2$pi.ns.outlier==1 & scate.summary.out2$pop=="KLDR" & scate.summary.out2$immune_DEG=="immune DEG")],
              "CEBO"=scate.summary.out2$geneID[which(scate.summary.out2$pi.ns.outlier==1 & scate.summary.out2$pop=="CEBO" & scate.summary.out2$immune_DEG=="immune DEG")])

ggvenn(pi_venn, set_name_size = 3, text_size=3, stroke_size=0.3, fill_color = viridispairs[c(3,6,12)], show_percentage = FALSE) + 
  coord_cartesian(clip="off") + 
  coord_equal() +
  labs(title=expression(paste(pi," positive immune outliers"))) + 
  theme(plot.title = element_text(hjust = 0.5, size=basesize+1))

ggarrange(b,c, labels=c("A","B"))
ggsave(filename=paste0(outdir,"/fig_pi_pos_immune.png"), plot=b, device="png",bg="white",width=80,height=80, units="mm") 

summary(scate.summary.out$n_polymorphic.ns[which(scate.summary.out$immune_DEG=="random")])
```

```{r Figure 1}
multipanel<- ggarrange(ggarrange(megapi.all,megataj.all,
                                 ncol=2,
                                 labels=c("A","D")),
                       ggarrange(pi_venn_upper_fig,mhc_plot,tajd_overall_fig,
                                 ncol=3,
                                 labels=c("B","C","E")), 
                       heights=c(2,1), 
                       nrow=2, 
                       font.label=list(size=basesize)) 


multipanel<- ggarrange(ggarrange(megapi.all,
                                 megataj.all,
                                 nrow=2),
                       ggarrange(NULL,
                                 ggarrange(pi_venn_lower_fig,
                                          pi_venn_upper_fig,
                                          ncol=2,
                                          widths=c(0.8,0.8)),
                                 mhc_plot,
                                 ggarrange(tajd_neg_fig,
                                           tajd_pos_fig,
                                           widths=c(0.8,0.8),
                                           ncol=2),
                                 NULL,
                                 nrow=5,
                                 ncol=1,
                                 heights = c(0.1,0.8,0.8,0.8,0.1)),
                       ncol=2, 
                       widths=c(1,0.5))
                                 
multipanel<- ggarrange(ggarrange(megapi.all,
                                 megataj.all,
                                 nrow=2,
                                 labels=LETTERS[c(1,7)]),
                       ggarrange(NULL,
                                 pi_venn_lower_fig,
                                 pi_venn_upper_fig,
                                 mhc_plot,
                                 tajd_neg_fig,
                                 tajd_pos_fig,
                                 NULL,
                                 nrow=7,
                                 ncol=1,
                                 heights=c(0.1,0.2,0.2,0.3,0.2,0.2,0.1),
                                 labels=c("",LETTERS[2:6])),
                       ncol=2, 
                       widths=c(1,0.5))
                                 
    

#ggsave(filename=paste0(outdir,"/fig_multipanel.png"),plot=multipanel, device="png", bg="white",width=220, height=205, units="mm")
ggsave(filename=paste0(outdir,"/fig_multipanel.png"),plot=multipanel, device="png", bg="white",width=150, height=250, units="mm")
#ggsave(filename=paste0(outdir,"/fig_multipanel.pdf"),plot=multipanel, device="pdf", bg="white",width=220, height=205, units="mm")
```

## F~ST~

```{r configure-fst}
#| label: configure-fst

genDes.fst <- read.csv(paste(datadir,"Fst","genDes.fst.csv",sep="/"),header = T)

deg.fst <- read.csv(paste(datadir,"Fst","deg.fst.csv",sep="/"),header = T)
max.deg.fst<- deg.fst %>%
  group_by(geneID) %>%
  summarise(maxFst=max(Fst))

deg.fst<- gene_categories(deg.fst, geneID)
deg.fst$immune_DEG<- ifelse(grepl("upper", deg.fst$immune_DEG),
                        gsub("upper ","", deg.fst$immune_DEG),
                        deg.fst$immune_DEG)
deg.fst$immune_DEG<- ifelse(grepl("lower", deg.fst$immune_DEG),
                        gsub("lower ","", deg.fst$immune_DEG),
                        deg.fst$immune_DEG)


deg.labels <- read.xlsx(paste(datadir,"GO_annot","Scate_DEGs_annot.xlsx",sep="/"))

deg.labels <- deg.labels %>% 
  select(Scate, geneName_parsed,Anolis_GeneID_GCF.GAF) %>%
  dplyr::rename(geneID=Scate) %>%
  mutate(geneID=gsub("-RA","",geneID)) %>% 
  filter(!duplicated(geneID)) %>%
  right_join(max.deg.fst, by="geneID") %>%
  mutate(geneName_parsed=ifelse(grepl("MHC Class I", geneName_parsed),
            paste0("MHCIA-",geneID),
            as.character(geneName_parsed)))

deg.labels<- gene_categories(deg.labels, geneID)
deg.labels$immune_DEG<- ifelse(grepl("upper", deg.labels$immune_DEG),
                        gsub("upper ","", deg.labels$immune_DEG),
                        deg.labels$immune_DEG)
deg.labels$immune_DEG<- ifelse(grepl("lower", deg.labels$immune_DEG),
                        gsub("lower ","", deg.labels$immune_DEG),
                        deg.labels$immune_DEG)

deg.labels.names<- deg.labels %>%
  filter(cat_immune=="immune") %>%
  filter(grepl("Uncharacterized",geneName_parsed,ignore.case=TRUE))

deg.labels99<- deg.labels %>%
  filter(maxFst>quantile(genDes.fst$Fst,0.99)) 
#conv.fst<- merge(deg.labels99, conv.ac, by="geneName_parsed", all.x=TRUE) #for the gProfiler GO

deg.labels99<- deg.labels99 %>%
  mutate(geneName_parsed=ifelse(geneName_parsed=="Uncharacterized",NA,as.character(geneName_parsed))) %>%
mutate(geneName_parsed=ifelse(immune_DEG=="DEG" & maxFst<0.90,NA,as.character(geneName_parsed)))
deg.labels95<- deg.labels %>%
  filter(maxFst>quantile(genDes.fst$Fst,0.95)) %>%
  mutate(geneName_parsed=ifelse(geneName_parsed=="Uncharacterized",NA,as.character(geneName_parsed))) %>%
  mutate(geneName_parsed=ifelse(immune_DEG=="DEG" & maxFst<0.90,NA,as.character(geneName_parsed)))

deg.labels05<- deg.labels %>%
  filter(maxFst<quantile(genDes.fst$Fst,0.05)) %>%
  mutate(geneName_parsed=ifelse(geneName_parsed=="Uncharacterized",NA,as.character(geneName_parsed)))


randNum1 <- runif(nrow(deg.labels99), min = 500, max = 6000)
randNum1 <- runif(nrow(deg.labels95), min = 500, max = 6000)
#https://support.bioconductor.org/p/9146885/
#q=number immune genes as outlier
#m=number immune genes in universe (not the genome overall but in the sampled universe of 762 genes)
#n=number of non-immune genes in universe.
#k=number of samples drawn (n>99th percentile)
phy<- phyper(q=length(unique(deg.labels95$geneID[deg.labels95$cat_immune!=""]))-1,
              m=length(unique(deg.fst$geneID[deg.fst$geneID %in% immune])),
              n=length(unique(deg.fst$geneID[!deg.fst$geneID %in% immune])),
              k=length(unique(deg.labels95$geneID)),
              lower.tail = FALSE)

length(deg.labels95$geneID[deg.labels95$geneID %in% scate.summary.out2$geneID[which(scate.summary.out2$tajima.ns.outlier!=0)]])
```

To look more directly investigate population specific divergent selection we characterized F~ST~ for `r nrow(deg.fst)` non-synonymous SNPs in `r nrow(deg.labels)` DEGs with NS SNPs and compared with gene desert SNPs. Across all *S. catenatus* populations, `r nrow(deg.labels95)` SNPs in `r length(unique(deg.labels95$geneID))` genes were outside the 95th percentile of gene desert SNP F~ST~ values, or `r round((length(unique(deg.labels95$geneID))/length(unique(deg.labels$geneID)))*100,1)`of total non-synonymous SNPs measured (@fig-fst95). Of these outliers, `r round((length(unique(deg.labels95$geneID[deg.labels95$cat_immune!=""]))/length(unique(deg.labels95$geneID))*100),1)`% of these are immune or inflammation genes, which is not a significant enrichment using a hypergeometric test (p = `r phy`).

```{r fig-fst95}
#| label: fig-fst95
#| warning: false
#| echo: false
#| fig-cap: "DEGs with highly divergent SNPs among all populations of S. catenatus. Grey bars represent the distribution of F~ST~ of gene desert SNPs. Coloured vertical bars indicate F~ST~ at each non-synonymous site in DEGS. Dark red line indicates the 99th percentile with selected outlier genes highlighted (NgeneDesert SNPs = 133,322, N non-synonymous SNPs = 5073 ,N SNPs in top 1% = 67)."
fst_plot<- ggplot(genDes.fst, 
                  aes(x=Fst)) + 
  geom_histogram(colour="gray", 
                 fill="gray", 
                 binwidth=0.005)+
  geom_vline(xintercept=quantile(genDes.fst$Fst,0.95),
             linetype=2, 
             color="darkred", 
             lwd=baselwd+0.3)+
  geom_rug(data=deg.fst, 
           aes(x = Fst, colour=immune_DEG), 
           inherit.aes = F, 
           linewidth=baselwd+0.1) +
  geom_segment(data = deg.labels95,
               aes(x = maxFst, xend = maxFst, y = 0, yend = randNum1, lwd=baselwd, colour=immune_DEG, alpha=0.5),
              linewidth = baselwd+0.1, 
              linetype=88, 
              show.legend = FALSE)+
  geom_text_repel(data = deg.labels95, 
                  aes(x = maxFst, y = randNum1, label = geneName_parsed, colour=immune_DEG),
                  hjust = -0.1, 
                  vjust = 0.5, 
                  max.overlaps = 15, 
                  box.padding = 0.5, 
                  show.legend = FALSE, 
                  size=geom.text.size, 
                  segment.size=baselwd) +
  scale_x_continuous(breaks=(seq(0,1,0.1)))+
  scale_y_continuous(breaks=(seq(0,6000,1000)))+
  labs(y=expression(paste("#",SNPs[geneDesert])),
       x=expression(paste(F[ST])),
       colour="") + 
  theme(legend.title=element_blank(),
        legend.position="top") + peri_figure

 
fst_plot

ggsave(paste0(outdir,"/fst-95.png"), fst_plot, device="png",width=150,height=110,units="mm", bg="white")
ggsave(paste(outdir,"report","figures","fst-95.png",sep="/"), fst_plot, device="pdf",width=150,height=110,units="mm", bg="white")

```

```{r}
#| warning: false
#| eval: false
#| fig.cap: "Top rated GO terms for FST outliers"

conv.fst<- conv.fst %>%
  filter(!duplicated(geneID)) %>%
  filter(converted_alias_manual_additions!="" & converted_alias_manual_additions!="None")

go_res<- gost(conv.fst$converted_alias_manual_additions, organism="acarolinensis", significant=FALSE, evcodes=TRUE)
go_res<- go_res$result
go_res<- go_res[order(go_res$p_value),]
rownames(go_res)<-1:nrow(go_res)
go_res<- go_res[1:30,]

go_out<- list()
for(i in 1:nrow(go_res)){
  sub<- go_res[i,]
  if(grepl(",",sub$intersection)){
    #i=1
    ensg<- str_split_1(sub$intersection,"\\,")
    sub2<- data.frame(p_value=rep(sub$p_value,length(ensg)),
                    source=rep(sub$source, length(ensg)),
                    term_name=rep(sub$term_name, length(ensg)),
                    converted_alias_manual_additions=ensg)
    go_out[[i]]<- sub2
  }else{
    sub$converted_alias_manual_additions<- sub$intersection

    go_out[[i]]<- sub[,c("p_value","source","term_name","converted_alias_manual_additions")]
  }
}
go_res<- do.call("rbind",go_out)
go_res$term<- paste(go_res$source,go_res$term_name)
go_res<- merge(go_res,conv.fst, by="converted_alias_manual_additions")

go_res$gene<- paste(go_res$geneID, go_res$geneName_parsed)
#remove the goterm with a long name for plotting purposes
go_res<- go_res[!grepl("acting on paired donors", go_res$term_name),]

ggplot(go_res, aes(x=gene,y=term,fill=p_value)) + 
  geom_tile() + peri_tile + coord_equal() +
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5)) + 
  labs(x="",y="")
```

```{r}
#| label: fig-fst-GO
#| warning: false
#| eval: false
#| fig.cap: "Top rated GO terms for FST outliers"

deg.labels95$gene_composite<- paste0(deg.labels95$Anolis_GeneID_GCF.GAF,deg.labels95$geneID)
go_results_fst<- enricher(unique(deg.labels95$gene_composite), 
                           universe=go2gene_bp$gene_composite, 
                           TERM2GENE = go2gene_bp, 
                           TERM2NAME = go2name_bp)
go_results_fst<- go_results_fst@result

go_results_fst$immune<- ifelse(go_results_fst$ID %in% immune_GO,"immune","other")
go_results_fst$Description<- factor(go_results_fst$Description, 
                          levels=unique(go_results_fst$Description[order(go_results_fst$immune)]))


ggplot(go_results_fst[which(go_results_fst$pvalue<0.05),], aes(y=Count, x=Description, fill=pvalue)) + 
  geom_bar(stat="identity") + peri_figure +
  labs(y="#genes",x="") + 
   theme(axis.text.x = element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5)) + 
  scale_y_continuous(expand=c(0,0))
#ggsave(filename=paste0(outdir,"/fig_FST_GO_sig.png"), plot=last_plot(), device="png",bg="white",width=460,height=130, units="mm") 

write.table(go_results_fst, paste0(outdir,"/GO_FST.tsv"), quote=FALSE, row.names = FALSE, sep="\t")

simMatrix <- calculateSimMatrix(go_results_fst$ID,
                                orgdb="org.Hs.eg.db",
                                ont="BP",
                                method="Rel")
scores <- setNames(go_results_fst$pvalue, go_results_fst$ID)
reducedTerms_fst <- reduceSimMatrix(simMatrix,
                               "uniqueness",
                                threshold=0.7,
                                orgdb="org.Hs.eg.db")

reducedTerms_fst <- reducedTerms_fst %>%
  dplyr::rename("ID"="go") %>%
  right_join(go_results_fst,by="ID")

reducedTerms_fst$immune<- ifelse(reducedTerms_fst$parent %in% immune_GO,"immune","other")
reducedTerms_fst$parentTerm<- factor(reducedTerms_fst$parentTerm, 
                          levels=unique(reducedTerms_fst$parentTerm[order(reducedTerms_fst$immune)]))

reducedTerms_fst$term<- ifelse(grepl("MHC",reducedTerms_fst$term),
                           "antigen processing and presentation\nof peptide antigen via MHC class I",
                           as.character(reducedTerms_fst$term))

fst_go<- ggplot(reducedTerms_fst[which(reducedTerms_fst$pvalue<0.05),], aes(x=term,y=Count, fill=pvalue)) + 
  geom_bar(stat="identity") + peri_figure + 
  scale_size_continuous(breaks=seq(2,12,by=2),range = c(1,3)) +
  labs(x="",y="#genes") +
  theme(axis.text.x = element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5), 
        strip.text.x = element_blank(),
        legend.justification="top",
        panel.spacing = unit(0, "lines")) + 
  facet_grid(~parentTerm, scales="free",space="free_x") +
  scale_fill_viridis(option="inferno") +
  scale_y_continuous(expand=c(0,0),
                     breaks = ~round(unique(pretty(.)))) 

ggsave(filename=paste0(outdir,"/fig_FST_GO_sig.png"), plot=fst_go, device="png",bg="white",width=130,height=120, units="mm") 
write.table(reducedTerms_fst,paste0(outdir,"/GO_FST_rrvgo.tsv"),row.names=FALSE,sep="\t")

```

```{r terms fst}
terms_fst<- gaf %>%
  filter(gaf$GeneID %in% deg.labels95$Anolis_GeneID_GCF.GAF) %>%
  filter(namespace=="biological_process") %>%
  group_by(name,GO_ID) %>%
  summarise(n_genes=n())

simMatrix <- calculateSimMatrix(terms_fst$GO_ID,
                                orgdb="org.Hs.eg.db",
                                ont="BP",
                                method="Rel")
scores <- setNames(terms_fst$n_genes, terms_fst$GO_ID)
reducedTerms_fst <- reduceSimMatrix(simMatrix,
                               "uniqueness",
                                threshold=0.7,
                                orgdb="org.Hs.eg.db")
reducedTerms_fst <- reducedTerms_fst %>%
  group_by(parentTerm,parent) %>%
  summarise(n_genes=sum(score))

reducedTerms_fst$immune<- ifelse(reducedTerms_fst$parent %in% immune_GO,"immune","other")
reducedTerms_fst$parentTerm<- factor(reducedTerms_fst$parentTerm, 
                          levels=unique(reducedTerms_fst$parentTerm[order(reducedTerms_fst$immune)]))


ggplot(reducedTerms_fst, aes(x=parentTerm,y=n_genes, fill=n_genes)) + geom_bar(stat="identity") + coord_flip() + peri_figure

```


```{r}
outliers_KEGG<- deg.labels95 %>%
  dplyr::rename(Scate=geneID) %>%
  select(Scate,maxFst,immune_DEG) %>%
  left_join(degs2anolis, by="Scate") %>%
  inner_join(immune_KEGG, by="Scate") %>%
  mutate(geneName_parsed=ifelse(grepl("MHC Class I", geneName_parsed),
                                paste0("MHCIA-",Scate), as.character(geneName_parsed)))

kegg_fst<- ggplot(outliers_KEGG,
       aes(x=geneName_parsed,y=Pathway_combined,fill=maxFst)) + 
  geom_tile() + peri_tile +
  theme(axis.text.x = element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5,
                                   size=basesize-1),
        legend.position = "top",
        legend.key.width=unit(5,"mm")) +
  scale_fill_gradient2(low="deepskyblue3",
                       high="red3",
                       na.value = "white",
                       mid="azure") +
  labs(y="KEGG pathway",
       x="",
       fill=expression(paste(F[ST])))
kegg_fst

ggsave(filename=paste0(outdir,"/fig_FST_KEGG_immune.png"), plot=last_plot(), device="png",bg="white",width=220,height=120, units="mm") 

#ggarrange(fst_plot,kegg_fst, nrow=2, labels=c("A","B"))
```


```{r FST immune GO}
outliers_GO<- deg.labels95 %>%
  dplyr::rename(Scate=geneID) %>%
  select(Scate,maxFst,immune_DEG) %>%
  left_join(degs2anolis, by="Scate") %>%
  inner_join(immune_GO, by="Scate") %>%
  dplyr::rename(geneName_parsed=geneName_parsed.x)
outliers_GO<- outliers_GO %>%
  mutate(geneName_parsed=ifelse(grepl("MHC Class I",
                                      outliers_GO$geneName_parsed),
                                paste0("MHCIA-",Scate), as.character(geneName_parsed)))


simMatrix <- calculateSimMatrix(outliers_GO$GO_ID,
                                orgdb="org.Hs.eg.db",
                                ont="BP",
                                method="Rel")
scores <- setNames(outliers_GO$maxFst, outliers_GO$GO_ID)
reducedTerms_immuneGO <- reduceSimMatrix(simMatrix,
                               "uniqueness",
                                threshold=0.65,
                                orgdb="org.Hs.eg.db") ## fiddle with the threshold till you get terms that make sense
outliers_GO<- reducedTerms_immuneGO %>%
  dplyr::rename(GO_ID=go) %>%
  right_join(outliers_GO, by="GO_ID") 

outliers_GO <- outliers_GO %>%
  mutate(parentTerm=ifelse(grepl("MHC",parentTerm),
                           "antigen processing and presentation \n of peptide antigen via MHC class I",
                           as.character(parentTerm))) 




fst_go<- ggplot(outliers_GO,
       aes(x=geneName_parsed,y=parentTerm,fill=maxFst)) + 
  geom_tile() + peri_tile +
  theme(axis.text.x = element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5,
                                   size=basesize-1),
        legend.position = "top",
        legend.key.width=unit(9,"mm")) +
  scale_fill_gradient2(low="deepskyblue3",
                       high="red3",
                       na.value = "white",
                       mid="azure") +
  labs(y="simplified GO term",
       x="",
       fill=expression(paste(F[ST])))

ggsave(filename=paste0(outdir,"/fig_fst_GO_immune.png"), plot=fst_go, device="png",bg="white",width=250,height=100, units="mm") 
```


```{r}
outliers_KEGG<- outliers_KEGG %>%
  select(Scate,geneName_parsed,maxFst,Pathway_combined) %>%
  dplyr::rename(term=Pathway_combined) %>%
  mutate(db="KEGG")

outliers_GO<- outliers_GO %>%
  select(Scate,geneName_parsed,maxFst,parentTerm) %>%
  dplyr::rename(term=parentTerm) %>%
  mutate(db="GO")

outliers_fst<- rbind(outliers_KEGG,outliers_GO)

fst_immune_plot<- ggplot(outliers_fst,
       aes(x=geneName_parsed,y=term,fill=maxFst)) + 
  geom_tile() + peri_tile + 
  facet_grid(db~., scales="free", space="free") + 
  theme(axis.text.x = element_text(angle=90, 
                                   hjust=1, 
                                   vjust=0.5,
                                   size=basesize-1),
        legend.position = "top",
        legend.key.width=unit(9,"mm")) +
  scale_fill_gradient2(low="deepskyblue3",
                       high="red3",
                       na.value = "white",
                       mid="azure") +
  labs(y="",
       x="",
       fill=expression(paste(F[ST])))


ggarrange(fst_plot,fst_immune_plot,
          nrow=2, 
          heights=c(0.8,1), 
          widths=c(0.9,1),
          labels=c("A","B")
          )

ggsave(filename=paste0(outdir,"/figure3.png"), plot=last_plot(), device="png",bg="white",width=220,height=230, units="mm") 

```

# Ideas for next time

-   Use phased SNPS to measure haplotype length (package REHH, with XP-EHH test for measuring population specific sweeps among multiple pops). This will allow more for linked selection and selection on non-coding regions. This wasn't really possible here because of issues with the way the VCF was filtered seemed to exclude a bunch of sites from *S. tergeminus* and made their diversity similar. However, this was based on looking at the filtering for the [@mathur2023], but only AN\<323 was applied with this dataset... which is essentially a missingness filter. So I think it means that there are a lot of unique sites that are variable in STER.

-   re-run Fu-Li's stats with *S. miliarus* as outgroup for *S. tergeminus* so they both can be directly.

-   simply characterize geographic genetic variation and selection in DEGs **and immune** genes across Ohio. Could do genome sliding windows of a few tests and then see if our genes overlap with regions outside of the chromosomal/genome wide averages. This would allow for linkage of synonymous and non-synonymous sites, as well as regulatory regions.


# All results summarized
```{r all outliers}

#| echo: false



imm<- scate.summary.out2 %>%
  full_join(deg.labels95,by=c("geneID")) %>%
  mutate(pop=ifelse(is.na(pop),"FST",as.character(pop))) %>%
  filter(pi.ns.outlier==1 | tajima.ns.outlier!=0 | !is.na(maxFst) | het.ns.outlier==1) %>%
  dplyr::rename(geneName_parsed=geneName_parsed.x) %>%
  select(geneID,geneName_parsed,pop,maxFst,n_polymorphic.ns, n_polymorphic.syn,heterozygosity.ns,minor_homoz.ns,pi.ns,tajima.ns,pi.ns.outlier, tajima.ns.outlier, het.ns.outlier) %>%
  mutate(immune=ifelse(geneID %in% immune_terms$Scate,"immune","non-immune"))

outlier_pi<- imm[which(imm$pi.ns.outlier==1),]
k.test<-  outlier_pi %>% 
  wilcox_test(pi.ns ~ pop) %>% 
  add_y_position() %>%
  adjust_pvalue(method="holm") %>%
  add_significance() 
ggplot(outlier_pi,aes(x=pop,y=pi.ns, colour=pop)) + geom_boxplot() + 
  stat_pvalue_manual(k.test, label = "p.adj.signif", tip.length = 0.005,hide.ns=TRUE, label.size=geom.text.size, bracket.size = baselwd)

outliers_list<- list(
  Tajima=imm$geneID[which(imm$tajima.ns.outlier!=0)],
  pi=imm$geneID[which(imm$pi.ns.outlier==1)],
  Fst=imm$geneID[!is.na(imm$maxFst)],
  Het=imm$geneID[which(imm$het.ns.outlier!=0)])

venn_all_outliers<- ggvenn(outliers_list,
                           set_name_size = 3, 
           text_size=3, 
           stroke_size=baselwd, 
           show_percentage = FALSE)



```



```{r outliers in both KLDR and STER}
kldrster<- imm %>%
  filter(pop!="CEBO") %>%
  mutate(maxFst=ifelse(!is.na(maxFst),1,0),
         pi.ns.outlier=ifelse(pi.ns.outlier==-1,0,pi.ns.outlier)) %>%
  pivot_longer(c(maxFst,tajima.ns.outlier,pi.ns.outlier),
               names_to="analysis",
               values_to="value") %>%
  group_by(geneID,analysis) %>%
  summarise(n_pops=sum(abs(value))) %>%
  filter(n_pops>1) %>%
  mutate(immune=ifelse(geneID %in% immune_terms$Scate,"immune","non-immune"))


kldrster_fst<- kldrster %>%
  filter(analysis=="maxFst") %>%
  filter(immune=="non-immune")
kldrster_tajima<- kldrster %>%
  filter(analysis=="tajima.ns.outlier")%>%
  filter(immune=="non-immune")

kldrster_pi<- kldrster %>%
  filter(analysis=="pi.ns.outlier")%>%
  filter(immune=="non-immune")

venn_kldrster_nonimmune<- ggvenn(list(Tajima=kldrster_tajima$geneID,
            pi=kldrster_pi$geneID,
            Fst=kldrster_fst$geneID),
            set_name_size = 3, 
           text_size=3, 
           stroke_size=baselwd, 
           show_percentage = FALSE) +
  labs(title="KLDR+STER+FST non-immune") +
  theme(plot.title = element_text(hjust = 0.5, size=basesize+1))

kldrster_fst<- kldrster %>%
  filter(analysis=="maxFst") %>%
  filter(immune=="immune")
kldrster_tajima<- kldrster %>%
  filter(analysis=="tajima.ns.outlier")%>%
  filter(immune=="immune")

kldrster_pi<- kldrster %>%
  filter(analysis=="pi.ns.outlier")%>%
  filter(immune=="immune")

venn_kldrster_immune<- ggvenn(list(Tajima=kldrster_tajima$geneID,
            pi=kldrster_pi$geneID,
            Fst=kldrster_fst$geneID),
            set_name_size = 3, 
           text_size=3, 
           stroke_size=baselwd, 
           show_percentage = FALSE)+
  labs(title="KLDR+STER+FST immune") +
  theme(plot.title = element_text(hjust = 0.5, size=basesize+1))
```
```{r all 3 pops agreement}
allpops<- imm %>%
  mutate(maxFst=ifelse(!is.na(maxFst),1,0),
         pi.ns.outlier=ifelse(pi.ns.outlier==-1,0,pi.ns.outlier)) %>%
  pivot_longer(c(maxFst,tajima.ns.outlier,pi.ns.outlier),
               names_to="analysis",
               values_to="value") %>%
  group_by(geneID,analysis) %>%
  summarise(n_pops=sum(abs(value))) %>%
  filter(n_pops>2) %>%
  mutate(immune=ifelse(geneID %in% immune_terms$Scate,"immune","non-immune"))

allpops_tajima<- allpops %>%
  filter(analysis=="tajima.ns.outlier")%>%
  filter(immune=="non-immune")

allpops_pi<- allpops %>%
  filter(analysis=="pi.ns.outlier")%>%
  filter(immune=="non-immune")

venn_3popoverlap_nonimmune<- ggvenn(list(Tajima=allpops_tajima$geneID,
            pi=allpops_pi$geneID,
            Fst=kldrster_fst$geneID),
            set_name_size = 3, 
           text_size=3, 
           stroke_size=baselwd, 
           show_percentage = FALSE)+
  labs(title="KLDR+STER+CEBO+FST non-immune") +
  theme(plot.title = element_text(hjust = 0.5, size=basesize+1))


allpops_tajima<- allpops %>%
  filter(analysis=="tajima.ns.outlier")%>%
  filter(immune=="immune")

allpops_pi<- allpops %>%
  filter(analysis=="pi.ns.outlier")%>%
  filter(immune=="immune")

venn_3popoverlap_immune<- ggvenn(list(Tajima=allpops_tajima$geneID,
            pi=allpops_pi$geneID,
            Fst=kldrster_fst$geneID),
            set_name_size = 3, 
           text_size=3, 
           stroke_size=baselwd, 
           show_percentage = FALSE)+
  labs(title="KLDR+STER+CEBO+FST immune") +
  theme(plot.title = element_text(hjust = 0.5, size=basesize+1))

ggarrange(venn_all_outliers,ggarrange(venn_3popoverlap_immune,
                                      venn_3popoverlap_nonimmune,
                                      venn_kldrster_immune,
                                      venn_kldrster_nonimmune,
                                      nrow=4,
                                      labels=c("B","C","D","E")),
          ncol=2,
          labels=c("A",""),
          widths=c(0.7,1))

ggsave(filename=paste0(outdir,"/figure_overlaps_outliers_all.png"), plot=last_plot(), device="png",bg="white",width=150,height=150, units="mm") 
```

```{r write all outliers}
imm <- imm %>%
  dplyr::rename("gene name"=geneName_parsed,
                "max Fst"=maxFst,
                "No. polymorphic sites non-syn"=n_polymorphic.ns,
                "No.polymorphic sites syn"=n_polymorphic.syn,
                "heterozygosity non-syn"=heterozygosity.ns,
                "homozygosity minor non-syn"=minor_homoz.ns,
                "pi non-syn"=pi.ns,
                "Tajima's D non-syn"=tajima.ns,
                "pi outlier non-syn"=pi.ns.outlier,
                "D outlier non-syn"=tajima.ns.outlier)


write.csv(imm, file = paste0(outdir,"/report/table_s6_result_outliers.csv"), row.names=FALSE)
```

```{r, eval=FALSE}
#immune.summary<- scate.summary.out2 %>%
#  filter(grepl("immune", immune_DEG)) %>%
#  filter(tajima.ns.outlier!=0) %>%
#  select(geneID,geneName_parsed,pop,maxFst,n_polymorphic.ns, n_polymorphic.syn,heterozygosity.ns,minor_homoz.ns,pi.ns,tajima.ns,pNpS,pnps.pi,pi.ns.outlier, tajima.ns.outlier)


immune.summary<- scate.summary.out2[grepl("immune",scate.summary.out2$immune_DEG),]
immune.summary$tajima.ns.outlier<- ifelse(immune.summary$tajima.ns.outlier==0,NA,immune.summary$tajima.ns.outlier)
immune.summary$tajima.ns.outlier<- as.factor(immune.summary$tajima.ns.outlier)
immune.summary<- immune.summary[!is.na(immune.summary$tajima.ns.outlier),]
immune.summary$gene_id<- paste0(immune.summary$geneName_parsed, "_",immune.summary$geneID)
immune.summary$gene_id<- factor(immune.summary$gene_id, levels=sort(unique(immune.summary$gene_id)))
immune.summary<- immune.summary[order(immune.summary$gene_id),]
immune.summary$pop<- factor(immune.summary$pop, levels=c("STER","KLDR","CEBO"))

excl<- c(CFH_1="Scate06776",CFH_2="Scate06774",S100A9_1="Scate15945",S100A9_2="Scate15944",EDNRB="Scate13869",FBXO32="Scate05992",RPS6KA2="Scate13201",TOP2A="Scate08785",PIK3R1="Scate03437",IL23R="Scate96945",LDLR="Scate02071",JCHAIN="Scate12145",MHCI_1="Scate05866",MHCI_2="Scate05865",MHCI_3="Scate05859",MHCI_4="Scate05854",MHCI_5="Scate05845",MHCI_6="Scate05841",MHCI_7="Scate05837",MHCI_8="Scate05863",MHCI_9="Scate05840",AHSG="Scate10425",APLNR="Scate15445",BIRC2="Scate10189",BMI1="Scate09285",CHD2="Scate03554",CXADR="Scate09940",CYP27B1="Scate04444",DOCK6="Scate04069",SLC1A2="Scate12392",SLC1A4="Scate01312",SLC1A5="Scate12959",SLC46A2="Scate03814",POMGNT1="Scate07070",GPRC6A="Scate00937",MAPK="Scate04428",PDE12="Scate05440",MFSD6="Scate01544",NKAIN="Scate12656",RPSK="Scate00745",ST14="Scate13393",SPPL3="Scate15244",NFATC3="Scate14109",MCOLN3="Scate06894",RUNX1T1="Scate14315",ACP5="Scate04085",CYBB="Scate09792",DLL4="Scate02451",EDNRB="Scate0956",FBXO="Scate01148",FGF2="Scate12364",FLT1="Scate10101",GNAO1="Scate14000",IMPDH="Scate12052",IMPDH2="Scate05590",LGALS4="Scate03093",LGALS="Scate00104", NOX1="Scate13824",PIKR3="Scate07072",RHOC="Scate07324", SLC12A1="Scate14582",TOPB2="Scate09159",VDAC="Scate05098")

immune.summary<- immune.summary[which(!immune.summary$geneID %in% excl),]
ggplot(immune.summary, aes(x=pop, y = gene_id, fill=tajima.ns.outlier)) + geom_tile()  + 
  peri_tile + 
  labs(x="",y="") +
  scale_fill_manual(values=c("#9bb6dc","#db9c9c")) +
  theme(legend.position="none",
        axis.text.x=element_text(angle=90,hjust=1, vjust=0.5))

ggsave(filename=paste0(outdir,"/poster/immune_tajima.pdf"),device="pdf",bg="transparent",height=18,width=7, units="in")
writeLines(unique(immune.summary$geneName_parsed),paste0(outdir,"/poster/immune_tajima.txt"))
geom_boxplot(data=scate.summary.out2[which(grepl("immune",scate.summary.out2$immune_DEG)),],aes(x=pop,y=log10(heterozygosity.ns)))
ggplot(data=scate.summary.out2[which(grepl("immune",scate.summary.out2$immune_DEG)),],aes(x=pop,y=log10(heterozygosity.ns))) + geom_boxplot()

#+ geom_line(data=scate.summary.out2[which(grepl("immune",scate.summary.out2$immune_DEG)),], aes(x=pop, y=log10(heterozygosity.ns), group=geneID))
```

# References
